### 2018-01-14 report/article.tex

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[2]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
This template helps you to create a properly formatted \LaTeX\ manuscript.
\end{abstract}

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{The Elsevier article class}

\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.

\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript. Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class. In addition, it features commands and options to format the
\begin{itemize}
\item document style
\item baselineskip
\item front matter
\item keywords and MSC codes
\item theorems, definitions and proofs
\item lables of enumerations
\item citation style and labeling.
\end{itemize}

\section{Front matter}

The author names and affiliations could be formatted in two ways:
\begin{enumerate}[(1)]
\item Group the authors per affiliation.
\item Use footnotes to indicate the affiliations.
\end{enumerate}
See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.

\section{Bibliography styles}

There are various bibliography styles available. You can select the style of your choice in the preamble of this document. These styles are Elsevier styles based on standard styles like Harvard and Vancouver. Please use Bib\TeX\ to generate your bibliography and include DOIs whenever available.

Here are two sample references: \cite{Feynman1963118,Dirac1953888}.

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2018-01-16 code/processing/L8_processing.py
updating readme to include how to download raw data and updating code folder structure

@@ -8,6 +8,7 @@ Input:
 Output:
         land surface temperature (oC)
         albedo
+        ndvi
 
 Notes:
         It is essential that the two raster files have the same:
@@ -22,11 +23,8 @@ Notes:
                 until they have the exact same cell size.
 '''
 
-
-
-
 from PIL import Image
-import numpy as np 
+import numpy as np
 from osgeo import gdal
 from osgeo import gdal_array
 from osgeo import osr
@@ -39,7 +37,7 @@ os.chdir('F:/UrbanDataProject/land_surface_temperature/data/cities/losangeles/20
 city = 'la'
 file_code = 'LC80410362016174LGN00'
 T_0 = 302.039 #max temp. observed in city (https://www.wunderground.com/history/?MR=1)
-M_L = 3.342e-4 # Band-10 specific multiplicative rescaling factor from the metadata 
+M_L = 3.342e-4 # Band-10 specific multiplicative rescaling factor from the metadata
 A_L = 0.10000 # Band-10 specific additive rescaling factor from the metadata
 K_1 = 774.8853 # Band-specific thermal conversion constant from the metadata
 K_2 =  1321.0789 # Band-specific thermal conversion constant from the metadata
@@ -79,7 +77,7 @@ def create_heatmap(filename,out_filename):
 
     #Next Step 3 in ppt
     T_sensor = step3(L_lambda)
-    
+
     temp_surface = step4(T_sensor, T_0,lc)
     temp_surface = celsius(temp_surface)
     # write to  tif
@@ -137,7 +135,7 @@ def reflectance(dn):
     mult_band = 2.0e-5
     add_band = -0.1
     refl = mult_band * dn + add_band
-    return refl 
+    return refl
 
 
 def albedo(set1_5):
@@ -148,8 +146,8 @@ def albedo(set1_5):
         reflect_band[i] = reflectance(ds.ReadAsArray())
         i += 1
 
-    alpha = ((0.356*reflect_band[0]) + (0.130*reflect_band[1]) + 
-            (0.373*reflect_band[2]) + (0.085*reflect_band[3]) + 
+    alpha = ((0.356*reflect_band[0]) + (0.130*reflect_band[1]) +
+            (0.373*reflect_band[2]) + (0.085*reflect_band[3]) +
             (0.072*reflect_band[4]) - 0.018) / 1.016
 
     array_to_raster(alpha, out_filename['albedo'], ds)

### 2018-01-16 report/article.tex

@@
-\documentclass[review]{elsarticle}
+\documentclass[]{elsarticle}
+%\documentclass[review]{elsarticle}
 
@@
 
-\journal{Journal of \LaTeX\ Templates}
+\journal{Remote Sensing of Environment}
 
@@
 
-\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
-\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
-
-%% Group authors per affiliation:
-\author{Elsevier\fnref{myfootnote}}
-\address{Radarweg 29, Amsterdam}
-\fntext[myfootnote]{Since 1880.}
+\title{Drivers of urban land surface temperature: diurnal and nocturnal}
 
 %% or include affiliations in footnotes:
-\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
-\ead[url]{www.elsevier.com}
+\author[1]{T.M. Logan\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead[url]{www.tomlogan.co.nz}
+\ead{tomlogan@umich.edu}
 
-\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
-\ead{support@elsevier.com}
+\author[2]{B. Zaitchik}
+\author[1]{S. Guikema}
 
-\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
-\address[mysecondaryaddress]{360 Park Avenue South, New York}
+\address[1]{Industrial and Operations Engineering, University of Michigan, Ann Arbor, MI}
+\address[2]{Earth and Planetary Sciences, Johns Hopkins University, Baltimore, MD}
 
 \begin{abstract}
-This template helps you to create a properly formatted \LaTeX\ manuscript.
+Understanding the drivers of urban land surface temperature can assist in ameliorating the severity of the urban heat island and risk from heat waves.
+Of all the natural events, heat waves are among the deadliest and are likely to become longer and more frequent.
+These changes, coupled with demographic shifts towards urban living, put significant impetus on reducing community exposure to heat waves.
+An understudied aspect of urban land surface temperature is factors related to nocturnal surface temperature.
+Nocturnal land surface temperature is important in heat wave mitigation given the urban heat island effect is most apparent during the night and the minimum nocturnal temperature is linked with heat stress and mortality.
+In this study, we examine the day and night urban land surface temperature (with Landsat) in seven cities across the United States.
+We test a series of hypotheses regarding the effect of greenspace, water bodies, and impervious surfaces on land surface temperature using advanced statistical methods.
+The robustness of these results is tested using different cities and satellite imagery and the results of diurnal and nocturnal analysis are compared to determine if the expected relationships hold throughout the day and night.
+Understanding whether the factors related to high urban temperatures are consistent across US cities is important for climate adaptation planning and capable predictive models suggest the potential for analysing the effect of potential biophysical changes might have on land surface temperature.
 \end{abstract}
@@
 \begin{keyword}
-\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-\MSC[2010] 00-01\sep  99-00
+
 \end{keyword}

### 2018-01-17 code/processing/L8_processing.py
adding automatic read of metadata txt

@@ -29,7 +29,11 @@ from osgeo import gdal
 from osgeo import gdal_array
 from osgeo import osr
 import os
-import logging
+
+# init logging
+import sys
+sys.path.append("..")
+from logger_config import *
 logger = logging.getLogger(__name__)
 
 os.chdir('F:/UrbanDataProject/land_surface_temperature/data/cities/losangeles/2016-06-22 LC80410362016174LGN00')
@@ -51,11 +55,60 @@ b4 = file_code + '_B4.tif'
 b5 = file_code + '_B5.tif'
 set1_5 = [b1, b2, b3, b4, b5]
 
+def main():
+    '''
+        1. Reads in metadata
+        2. Creates map of land surface temperature
+        3. Creates map of NVDI
+        4. Creates map of albedo
+    '''
 
+    # Create map of land surface temperature
+    create_heatmap(filename, out_filename)
+
+    # create nvdi map
+    calc_NDVI(set1_5)
+
+    # create map of albedo
+    albedo(set1_5)
+
+def read_image_list():
+    '''
+        Loop through the satellite image list
+        Return:
+            Satellite id
+            Temperature (unit?)
+    '''
+
+def read_metadata():
+    '''
+        For the image read the metadata txt file
+        Return
+            Dictionary of metadata
+    '''
+    # list of variables needed from metadata
+    meta_variables = set(['RADIANCE_MULT_BAND_10','RADIANCE_ADD_BAND_10'])
+
+    # init dictionary
+    meta_dict = dict.fromkeys(meta_variables)
+    # open the metadata file
+    metadata_fn = 'LC08_L1TP_015033_20170907_20170926_01_T1_MTL.txt'
+    with open(metadata_fn,'r') as fid:
+        for line in fid:
+            elements = line.split()
+            var = meta_variables.intersection(elements)
+            if var:
+                meta_dict[list(var)[0]] = float(elements[-1])
+
+        if
+
+    return meta_dict
 
 def create_heatmap(filename,out_filename):
-    ''' returns the land surface temperatue based on satellite imagery
     '''
+        Returns the land surface temperatue based on satellite imagery
+    '''
+    logger.info('Starting LST calculations')
 
     print("You're operating in " + os.getcwd())
     # open the satellite imagery
@@ -63,11 +116,11 @@ def create_heatmap(filename,out_filename):
[…]

### 2018-01-17 code/processing/L8_processing.py
commenting the metadata read script

@@ -94,13 +94,17 @@ def read_metadata():
     # open the metadata file
     metadata_fn = 'LC08_L1TP_015033_20170907_20170926_01_T1_MTL.txt'
     with open(metadata_fn,'r') as fid:
+        # loop lines in text file
         for line in fid:
+            # divide line into words, split by space
             elements = line.split()
+            # intersect with metavariables to return set of variable name if it is on line
             var = meta_variables.intersection(elements)
-            if var:
-                meta_dict[list(var)[0]] = float(elements[-1])
-
-        if
+            if var: # if not empty set
+                # extract the variable
+                var = list(var)[0]
+                # add to dictionary
+                meta_dict[var] = float(elements[-1])
 
     return meta_dict
 

### 2018-01-17 code/processing/L8_processing.py
updating L8 processing

@@ -108,9 +108,9 @@ def read_metadata():
 
     return meta_dict
 
-def create_heatmap(filename,out_filename):
+def calc_LST(filename,out_filename):
     '''
-        Returns the land surface temperatue based on satellite imagery
+        Returns the land surface temperature (LST) based on satellite imagery
     '''
     logger.info('Starting LST calculations')
 
@@ -123,12 +123,8 @@ def create_heatmap(filename,out_filename):
     # Conversion to TOA Radiance
     TOA = calc_TOA(dn)
 
-    # Step 2: Emissivity Correction
-    #returns an emissivity map from land cover
-    land_cover = gdal.Open(lc_filename)
-    lc = land_cover.ReadAsArray()
-    print("LC tif size: " + str(np.shape(lc)))
-    emissiv = get_emissivity(lc)
+    # Emissivity Correction
+    emissivity = determine_emissivity(lc)
 
     L_lambda = TOA/emissiv
 
@@ -148,25 +144,38 @@ def calc_TOA(dn):
         number (DN) data (also refered to as the Q_cal - quantized and
         calibrated standard product pixel value)
     '''
+    logger.info('Calculating TOAr')
 
     TOAr = meta_dict['RADIANCE_MULT_BAND_10'] * dn + meta_dict['RADIANCE_ADD_BAND_10']
 
     return(TOAr)
 
 
-def get_emissivity(land_cover):
+def determine_emissivity():
+    '''
+        Emissivity is determined by the land cover
+        Return
+            Emissivity array from land cover map
+    '''
+    logger.info('Determining emissivity map')
+    # import land cover
+    land_cover = gdal.Open(lc_filename)
+    # convert to array
+    land_cover = land_cover.ReadAsArray()
+    logger.info("LC tif size: " + str(np.shape(land_cover)))
     land_cover = land_cover.astype(float)
+    emissivity = land_cover.copy()
 
     #convert land_cover to emissivity array
-    land_cover[land_cover > 90] = 0.957 # Wetlands is assumed to be Sparse Vegetation
-    land_cover[(1 < land_cover) & (land_cover < 20)] = 0.989 # Water
-    land_cover[(1 < land_cover) & (land_cover < 30)] = 0.912 # Urban
-    land_cover[(1 < land_cover) & (land_cover < 40)]  = 0.896 # Barren
[…]

### 2018-01-18 code/processing/L8_processing.py
updated the calc_LST and sub functions

@@ -87,7 +87,7 @@ def read_metadata():
             Dictionary of metadata
     '''
     # list of variables needed from metadata
-    meta_variables = set(['RADIANCE_MULT_BAND_10','RADIANCE_ADD_BAND_10'])
+    meta_variables = set(['RADIANCE_MULT_BAND_10','RADIANCE_ADD_BAND_10','K1_CONSTANT_BAND_10','K2_CONSTANT_BAND_10'])
 
     # init dictionary
     meta_dict = dict.fromkeys(meta_variables)
@@ -123,22 +123,17 @@ def calc_LST(filename,out_filename):
     # Conversion to TOA Radiance
     TOA = calc_TOA(dn)
 
-    # Emissivity Correction
-    emissivity = determine_emissivity(lc)
+    # Calculate the At-Satellite Brightness Temperature
+    temp_satellite = calc_satellite_temperature(TOA)
 
-    L_lambda = TOA/emissiv
+    # Atmospheric correction
+    temp_surface = atmos_correction(temp_satellite, T_0,lc)
 
-    #Next Step 3 in ppt
-    T_sensor = step3(L_lambda)
-
-    temp_surface = step4(T_sensor, T_0,lc)
-    temp_surface = celsius(temp_surface)
-    # write to  tif
-    output = temp_surface
-    array_to_raster(output, out_filename['lst'], land_cover)
+    # write to tif
+    array_to_raster(temp_surface, out_filename['lst'], land_cover)
 
 
-def calc_TOA(dn):
+def calc_TOA(dn, meta_dict):
     '''
         Calculate the Top Atmosphere Spectral Radiance (TOAr) from Band 10 digital
         number (DN) data (also refered to as the Q_cal - quantized and
@@ -178,13 +173,34 @@ def determine_emissivity():
     return(emissivity)
 
 
-def step3(L_lambda):
+def calc_satellite_temperature(TOA, meta_dict):
+    '''
+        Calculate the At-Satellite Brightness temperature
+        First, need to calculate the emissivity from the land use
+        Then convert
+        Returns temperature in Kelvin
+    '''
+    logger.info('calculating satellite temperature')
+    # Emissivity Correction
+    emissivity = determine_emissivity(lc)
 
-    T = K_2/(np.log((K_1/L_lambda) + 1))
-    return T
+    # spectral radiance
+    L_lambda = TOA/emissiv
+
[…]

### 2018-01-18 code/processing/L8_processing.py
calculating albedo and updating meta_data function

@@ -57,6 +57,14 @@ set1_5 = [b1, b2, b3, b4, b5]
 
 def main():
     '''
+        loops through satellite images and processes them
+    '''
+    #
+
+
+
+def process_image():
+    '''
         1. Reads in metadata
         2. Creates map of land surface temperature
         3. Creates map of NVDI
@@ -64,7 +72,7 @@ def main():
     '''
 
     # Create map of land surface temperature
-    create_heatmap(filename, out_filename)
+    calc_LST(filename, out_filename)
 
     # create nvdi map
     calc_NDVI(set1_5)
@@ -87,7 +95,12 @@ def read_metadata():
             Dictionary of metadata
     '''
     # list of variables needed from metadata
-    meta_variables = set(['RADIANCE_MULT_BAND_10','RADIANCE_ADD_BAND_10','K1_CONSTANT_BAND_10','K2_CONSTANT_BAND_10'])
+    meta_variables = set(['K1_CONSTANT_BAND_10','K2_CONSTANT_BAND_10'])
+    bands = [1,2,3,4,5,10]
+    # include radiance re-scaling factors
+    for rad in ['MULT', 'ADD']:
+        for b in bands:
+            meta_variables.add('RADIANCE_{}_BAND_{}'.format(rad, b))
 
     # init dictionary
     meta_dict = dict.fromkeys(meta_variables)
@@ -115,13 +128,13 @@ def calc_LST(filename,out_filename):
     logger.info('Starting LST calculations')
 
     print("You're operating in " + os.getcwd())
-    # open the satellite imagery
+    # read in band 10 data
     ds = gdal.Open(filename)
     dn = ds.ReadAsArray()
     print("L8 tif size: " + str(np.shape(dn)))
 
     # Conversion to TOA Radiance
-    TOA = calc_TOA(dn)
+    TOA = calc_TOA(dn, meta_dict, 10)
 
     # Calculate the At-Satellite Brightness Temperature
     temp_satellite = calc_satellite_temperature(TOA)
@@ -133,7 +146,7 @@ def calc_LST(filename,out_filename):
     array_to_raster(temp_surface, out_filename['lst'], land_cover)
 
 
-def calc_TOA(dn, meta_dict):
+def calc_TOA(dn, meta_dict,band_number):
     '''
         Calculate the Top Atmosphere Spectral Radiance (TOAr) from Band 10 digital
[…]

### 2018-01-18 code/processing/L8_processing.py
adding code that calculates NVDI

@@ -262,18 +262,22 @@ def calc_albedo(set1_5):
 
 
 def calc_NDVI(set1_5):
-    ''' calculate the NDVI '''
+    '''
+        calculate the NDVI
+        For landsat8 it's (B5 – B4) / (B5 + B4) - see Ben's email dated 7/26/16
+    '''
     # import bands
     landsat_band = dict()
-    i = 0
+    band_number = 1
     for name in set1_5:
         ds = gdal.Open(name)
-        landsat_band[i] = ds.ReadAsArray()
-        i += 1
+        landsat_band[band_number] = ds.ReadAsArray()
+        band_number += 1
 
-    NDVI = (landsat_band[4] - landsat_band[3])/(landsat_band[4] + landsat_band[3])
+    # calculate the NDVI
+    ndvi = (landsat_band[5] - landsat_band[4])/(landsat_band[5] + landsat_band[4])
 
-    array_to_raster(NDVI, out_filename['nvdi'], ds)
+    array_to_raster(ndvi, out_filename['nvdi'], ds)
 
 
 def array_to_raster(output, out_filename, ds):

### 2018-04-22 report/article.tex
Updates from ShareLaTeX

@@
 
-\title{Drivers of urban land surface temperature: diurnal and nocturnal}
+\title{Understanding the drivers of diurnal and nocturnal urban land surface temperature}
 
@@
 \begin{abstract}
-Understanding the drivers of urban land surface temperature can assist in ameliorating the severity of the urban heat island and risk from heat waves.
-Of all the natural events, heat waves are among the deadliest and are likely to become longer and more frequent.
-These changes, coupled with demographic shifts towards urban living, put significant impetus on reducing community exposure to heat waves.
+Understanding the drivers of urban land thermal radiance, a key factor of land surface temperature, can assist in ameliorating the severity of the urban heat island and risk from heat waves.
+Of all the natural events, heat waves are among the deadliest and are likely to become longer and more frequent with climatic changes.
+Coupled with demographic shifts towards urban living, climate change puts significant impetus on reducing community exposure to heat waves.
 An understudied aspect of urban land surface temperature is factors related to nocturnal surface temperature.
@@
 
-\section{The Elsevier article class}
+\section{Introduction}
 
-\paragraph{Installation} If the document class \emph{elsarticle} is not available on your computer, you can download and install the system package \emph{texlive-publishers} (Linux) or install the \LaTeX\ package \emph{elsarticle} using the package manager of your \TeX\ installation, which is typically \TeX\ Live or Mik\TeX.
+\section{Data and Methods}
 
-\paragraph{Usage} Once the package is properly installed, you can use the document class \emph{elsarticle} to create a manuscript.
-Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal.
-It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.
+\section{Results and Discussion}
 
-\paragraph{Functionality} The Elsevier article class is based on the standard article class and supports almost all of the functionality of that class.
-In addition, it features commands and options to format the
-\begin{itemize}
-\item document style
-\item baselineskip
-\item front matter
-\item keywords and MSC codes
-\item theorems, definitions and proofs
-\item lables of enumerations
-\item citation style and labeling.
-\end{itemize}
-
-\section{Front matter}
-
-The author names and affiliations could be formatted in two ways:
-\begin{enumerate}[(1)]
-\item Group the authors per affiliation.
-\item Use footnotes to indicate the affiliations.
-\end{enumerate}
-See the front matter of this document for examples. You are recommended to conform your choice to the journal you are submitting to.
-
-\section{Bibliography styles}
+\section{Conclusion}
 
@@
 
+\appendix
+\section{2013 Land Surface Temperature Image}
+
 \end{document}

### 2018-04-23 report/article.tex
Updates from ShareLaTeX

@@
 
+\subsection{Variables related to Land Surface Temperature}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/holdout_results_r2.pdf}
+\caption{ADD Caption}
+\label{fig:cityholdout_errors}
+\end{center}
+\end{figure}
+
+\subsection{Predicting Land Surface Temperature Between Cities}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/cities_holdout.pdf}
+\caption{ADD Caption}
+\label{fig:cityholdout_errors}
+\end{center}
+\end{figure}
+
 \section{Conclusion}
@@
 
+\section{Additional model results}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/holdout_results_mae.pdf}
+\caption{ADD Caption}
+\label{fig:cityholdout_errors}
+\end{center}
+\end{figure}
+
 \end{document}

### 2018-04-26 report/article.tex

@@
 \includegraphics[width=\textwidth]{fig/report/holdout_results_r2.pdf}
-\caption{ADD Caption}
+\caption{The out-of-bag R$^2$ of the models when fitted on each city and used to predict a random sample from the same city.
+Out-of-bag R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
+This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.}
 \label{fig:cityholdout_errors}
@@
 \includegraphics[width=\textwidth]{fig/report/cities_holdout.pdf}
-\caption{ADD Caption}
+\caption{The out-of-bag R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
+Out-of-bag R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
+Good models have MAE near 0.
+The gradient boosted regression forest generally outperforms the other models, except when predicting Detroit.
+Note that the R$^2$ axis is truncated at -1, although the multiple linear regression for diurnal land surface temperature tested on Phoenix has an OOB R$^2$ of -9.
+This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.}
 \label{fig:cityholdout_errors}
@@
 \includegraphics[width=\textwidth]{fig/report/holdout_results_mae.pdf}
-\caption{ADD Caption}
+\caption{The Mean Absolute Error (MAE) in $^oC$ of the models when fitted on each city and used to predict a random sample from the same city.
+Better models have lower MAE.
+This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.
+}
 \label{fig:cityholdout_errors}

### 2018-04-26 code/analysis/analysis.py
running forward variable selection for all cities and day/night

@@ -1,0 +1,708 @@
+'''
+Here are a collection of functions for analyzing the land surface temperature and
+thermal radiance from LandSat images of cities
+'''
+
+# import libraries
+import matplotlib as mpl
+mpl.use('Agg')
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+import brewer2mpl
+import seaborn as sns
+from sklearn.model_selection import train_test_split
+from sklearn import preprocessing
+import pickle
+
+# regression libraries
+from sklearn.ensemble import RandomForestRegressor
+from sklearn.ensemble import GradientBoostingRegressor
+from sklearn.linear_model import LinearRegression
+from sklearn.ensemble.partial_dependence import plot_partial_dependence
+from sklearn.ensemble.partial_dependence import partial_dependence
+from sklearn.metrics import mean_squared_error, r2_score
+from xgboost import XGBRegressor
+
+# init logging
+import sys
+# sys.path.append("code")
+# from logger_config import *
+# logger = logging.getLogger(__name__)
+
+
+RANDOM_SEED = 3201
+
+def main():
+    '''
+    '''
+    # loop cities and all
+    cities = ['bal', 'por', 'det', 'phx']
+    # import data
+    df = import_data(cities)
+
+    # data transformations
+    df = transform_data(df)
+
+    # present the data - density plot
+    # plot_density(df, cities)
+
+    # regression
+    ## train on three cities, test on one
+    # loss = regression_cityholdouts(df, cities)
+    # plot the points
+    # plot_holdout_points(loss)
+
+    ## for each city, train and test
+    # sim_num = 100
+    # regressions(df, cities, sim_num)
+    # plot_holdouts()
+
+    # variable importance and partial dependence
+    # reg_gbm = full_gbm_regression(df, cities)
+    # variable importance
+
+    # variable selection
+    from datetime import datetime
+    vars_forward = {}
+    vars_forward['day'] = {}
+    vars_forward['night'] = {}
+    for city in cities+['all']:
+        for period in ['day','night']:
+            print('{}: Starting {}, {}'.format(datetime.now().strftime('%Y-%m-%d %H:%M:%S'), city, period))
+            vars_forward[period][city] = feature_selection(25, city, df, period)
[…]

### 2018-04-26 code/analysis/analysis.py
run variable selection

@@ -246,6 +246,10 @@ def prepare_lst_prediction(df):
     vars_lcov = [i for i in df.columns.values if 'lcov' in i]
     df = df.drop(vars_lcov, axis=1)
 
+    # drop impervious variables because they are 1:1 correlated with tree canopy
+    vars_imp = [i for i in df.columns.values if 'imp' in i]
+    df = df.drop(vars_imp, axis=1)
+
     return(df, lst_mean)
 
 def scale_X(X_train, X_test):
@@ -477,7 +481,8 @@ def feature_selection(holdout_num, city, df, period):
     # add variables based on which provide the best improvement to lowering MAE
     vars_inc = []
     vars_mae = []
-    while len(vars_inc)<len(variables):
+    num_vars = len(variables)
+    while len(vars_inc)<num_vars:
         # loop through the Variables
         variables = [var for var in variables if var not in vars_inc]
         variable_mae = pd.DataFrame(index=variables, columns=['mae'])

### 2018-05-23 code/analysis/analysis.py
updating elev scaling

@@ -67,7 +67,8 @@ def main():
     # based on the results of the variable selection, rerun the regression and
     # create the variable importance plots
     #vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'elev_min_sl', 'elev_max', 'tree_max_sl']
-    vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'alb_mean_min', 'elev_min_sl', 'ndvi_mean_min']
+    # vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'alb_mean_min', 'elev_min_sl', 'ndvi_mean_min']
+    vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'elev_min', 'alb_mean_min_sl', 'elev_max']
     reg_gbm, X_train = full_gbm_regression(df, cities, vars_selected)
     #
     # # plot the variable importance
@@ -139,6 +140,21 @@ def transform_data(df):
     # make tree values a percentage
     vars_tree = [i for i in vars_all if 'tree' in i]
     df.loc[:,vars_tree] = df[vars_tree]/100
+
+    ###
+    # elevation transform and scale
+    ###
+    vars_elev = [i for i in vars_all if 'elev' in i]
+    cities = np.unique(df['city'])
+    # subtract city median elevation
+    medians = df.groupby(df.city)[vars_elev].median().median(axis=1)
+    for city in cities:
+        df.loc[df['city']==city,vars_elev] = df.loc[df['city']==city,vars_elev] - medians[city]
+    # scale the cities
+    elev_max = np.max(df[vars_elev].values.max())
+    elev_min = np.min(df[vars_elev].values.min())
+    df.loc[:,vars_elev] = (df.loc[:,vars_elev] - elev_min)/(elev_max-elev_min)
+
     return(df)
 
 def regressions(df, cities, sim_num):
@@ -276,33 +292,33 @@ def scale_X(X_train, X_test):
     # X_scaled = scaler.transform(X_train)
     # X_train = pd.DataFrame(data = X_scaled, columns = X_train.columns.values)
     # X_test = pd.DataFrame(data = scaler.transform(X_test), columns = X_test.columns.values)
-    vars_elev = [i for i in vars_all if 'elev' in i]
-    if len(vars_elev)>0:
-        # print(cities)
-        # print(len(cities))
-        if len(cities) > 1:
-            # normalize elevation by subtracting median of city from the city
[…]

### 2018-05-25 report/article.tex

@@
 \includegraphics[width=\textwidth]{fig/report/cities_holdout.pdf}
-\caption{The out-of-bag R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
-Out-of-bag R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
+\caption{The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
+OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
 Good models have MAE near 0.

### 2018-06-28 report/article.tex

@@
 This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.}
-\label{fig:cityholdout_errors}
+\label{fig:intracity}
 \end{center}
@@
 \subsection{Predicting Land Surface Temperature Between Cities}
+
+See figure \ref{fig:intercity}.
 
@@
 This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.}
-\label{fig:cityholdout_errors}
+\label{fig:intercity}
 \end{center}
@@
 
-\section{Conclusion}
+\section{Drivers of LST}
 
-There are various bibliography styles available.
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/variable_importance_selected.pdf}
+\caption{}
+\label{fig:importance}
+\end{center}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/partial_dependence.pdf}
+\caption{}
+\label{fig:importance}
+\end{center}
+\end{figure}
+
+% \section{Conclusion}
+
+% There are various bibliography styles available.
 You can select the style of your choice in the preamble of this document.
@@
 
-Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
+% Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
 
-\section*{References}
+% % \section*{References}
 
-\bibliography{mybibfile}
+% \bibliography{mybibfile}
 

### 2018-06-28 code/analysis/analysis.py
keep water

@@ -120,7 +120,7 @@ def transform_data(df):
         df[var_lcov] = df[var_lcov]/area_max
 
     # drop rows with water more than 20% of area
-    df = df.loc[df['lcov_11'] < 0.2]
+    # df = df.loc[df['lcov_11'] < 0.2]
 
     # Drop the 2013 thermal radiance measure (this is in bal dataset for validation)
     tr_2013 = [s for s in df.columns.values if 'tr_2013' in s]
@@ -272,6 +272,8 @@ def prepare_lst_prediction(df):
 
     # drop lcov variables
     vars_lcov = [i for i in df.columns.values if 'lcov' in i]
+    # keep water (lcov_11)
+    vars_lcov = [i for i in vars_lcov if '11' not in i]
     df = df.drop(vars_lcov, axis=1)
 
     # drop impervious variables because they are 1:1 correlated with tree canopy

### 2018-06-28 code/analysis/analysis.py
running results with water

@@ -45,37 +45,37 @@ def main():
     df = transform_data(df)
 
     # present the data - density plot
-    # plot_density(df, cities)
-    #
-    # # regression
-    # ## train on three cities, test on one
-    # loss = regression_cityholdouts(df, cities)
-    # # plot the points
-    # plot_holdout_points(loss)
-    #
-    # ## for each city, train and test
-    # sim_num = 100
-    # regressions(df, cities, sim_num)
-    # plot_holdouts()
-    #
-    # # variable importance and partial dependence
-    # reg_gbm = full_gbm_regression(df, cities)
-    #
-    # # variable selection
-    # loop_variable_selection(df, cities)
+    plot_density(df, cities)
+
+    # regression
+    ## train on three cities, test on one
+    loss = regression_cityholdouts(df, cities)
+    # plot the points
+    plot_holdout_points(loss)
+
+    ## for each city, train and test
+    sim_num = 100 # number of holdouts
+    regressions(df, cities, sim_num)
+    plot_holdouts()
+
+    # variable importance and partial dependence
+    reg_gbm = full_gbm_regression(df, cities)
+
+    # variable selection
+    loop_variable_selection(df, cities)
 
     # based on the results of the variable selection, rerun the regression and
     # create the variable importance plots
     #vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'elev_min_sl', 'elev_max', 'tree_max_sl']
     # vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'alb_mean_min', 'elev_min_sl', 'ndvi_mean_min']
-    vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'elev_min', 'alb_mean_min_sl', 'elev_max']
-    reg_gbm, X_train = full_gbm_regression(df, cities, vars_selected)
-    #
-    # # plot the variable importance
-    importance_order = plot_importance(reg_gbm, cities)
-    #
-    # # plot the partial dependence
-    plot_dependence(importance_order, reg_gbm, cities, X_train, vars_selected, show_plot=False)
+    # vars_selected = ['tree_mean', 'ndvi_mean_mean', 'alb_mean_mean', 'elev_min', 'alb_mean_min_sl', 'elev_max']
+    # reg_gbm, X_train = full_gbm_regression(df, cities, vars_selected)
+    # #
+    # # # plot the variable importance
[…]

### 2018-06-30 code/analysis/analysis.py
consider only land cells with water < 0.2

@@ -120,7 +120,7 @@ def transform_data(df):
         df[var_lcov] = df[var_lcov]/area_max
 
     # drop rows with water more than 20% of area
-    # df = df.loc[df['lcov_11'] < 0.2]
+    df = df.loc[df['lcov_11'] < 0.2]
 
     # Drop the 2013 thermal radiance measure (this is in bal dataset for validation)
     tr_2013 = [s for s in df.columns.values if 'tr_2013' in s]

### 2018-07-08 report/article.tex

@@
 \section{Introduction}
+In a warming world, understanding the drivers of surface temperatures during both the day and night will aid in adapting cities to mitigate urban heat for the health and well being of communities.
 
@@
 \section{Drivers of LST}
+See figure \ref{fig:importance} and \ref{fig:pdp}.
+
+Regarding Fig \ref{fig:pdp}, I think I will remove the ``all" line, because in cases such as $alb_mean_min_sl$ (which is the minimum albedo in the surrounding cells (sl = spatial lag)) the values between the cities don't overlap so I think it becomes a proxy for other differences between the city when regressed together.
 
@@
 \includegraphics[width=\textwidth]{fig/report/variable_importance_selected.pdf}
-\caption{}
+\caption{The variable importance for the regression during the night (to the left) and the day (to the right).
+Variable importance indicates the amount that each variable improves the performance of the regression.
+In this case there is generally little difference between the importance of day and night variables.}
 \label{fig:importance}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{fig/report/partial_dependence.pdf}
-\caption{}
-\label{fig:importance}
+\includegraphics[height=\textheight]{fig/report/partial_dependence.pdf}
+\caption{Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.}
+\label{fig:pdp}
 \end{center}

### 2018-07-15 report/article.tex
working on readme and intro for report

@@
 \begin{abstract}
-Understanding the drivers of urban land thermal radiance, a key factor of land surface temperature, can assist in ameliorating the severity of the urban heat island and risk from heat waves.
-Of all the natural events, heat waves are among the deadliest and are likely to become longer and more frequent with climatic changes.
-Coupled with demographic shifts towards urban living, climate change puts significant impetus on reducing community exposure to heat waves.
-An understudied aspect of urban land surface temperature is factors related to nocturnal surface temperature.
-Nocturnal land surface temperature is important in heat wave mitigation given the urban heat island effect is most apparent during the night and the minimum nocturnal temperature is linked with heat stress and mortality.
-In this study, we examine the day and night urban land surface temperature (with Landsat) in seven cities across the United States.
-We test a series of hypotheses regarding the effect of greenspace, water bodies, and impervious surfaces on land surface temperature using advanced statistical methods.
-The robustness of these results is tested using different cities and satellite imagery and the results of diurnal and nocturnal analysis are compared to determine if the expected relationships hold throughout the day and night.
-Understanding whether the factors related to high urban temperatures are consistent across US cities is important for climate adaptation planning and capable predictive models suggest the potential for analysing the effect of potential biophysical changes might have on land surface temperature.
+Although heat waves and the urban heat island are nocturnal phenomenon, the drivers of land surface temperature during the night are poorly understood.
+Given that heat waves are the deadliest natural hazard and are expected to increase in frequency and severity, determining mitigation strategies is important.
+However, only recently has nocturnal imagery of cities become available by LandSat allowing nighttime land surface temperature to be analyzed.
+Furthermore, it remains unclear from existing studies the independent effects of potential drivers, and their relative importance.
+Here, we seek to answer the question: what are the effects and importance of drivers of land surface temperature during both day and night?
+To do this, we analyze the urban land surface temperature in four cities across the United States.
+We include factors related to vegetation, water, the built-environment, and topography and control them using nonlinear statistical methods which allow for their independent effects to be determined.
+The results of diurnal and nocturnal analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
+Understanding whether the factors related to high urban land surface temperatures are consistent across US cities is important for climate adaptation planning and during heat wave mitigation strategies.
 \end{abstract}
@@
 In a warming world, understanding the drivers of surface temperatures during both the day and night will aid in adapting cities to mitigate urban heat for the health and well being of communities.
+Of all the natural events, heat waves are among the deadliest and are likely to become longer and more frequent with climatic changes.
+Coupled with demographic shifts towards urban living, climate change puts significant impetus on reducing community exposure to heat waves.
+An understudied aspect of urban land surface temperature are factors related to nocturnal surface temperature.
+Nocturnal land surface temperature is important in heat wave mitigation given the urban heat island effect is most apparent during the night and the minimum nocturnal temperature is linked with heat stress and mortality.
 

### 2018-07-15 code/analysis/analysis.py
update city legend in varimp

@@ -763,7 +763,7 @@ def plot_importance(reg_gbm, cities, show_plot=False):
     plt.ylabel('Variables')
     # legend
     handles, labels = ax.get_legend_handles_labels()
-    l = plt.legend(handles[0:5], labels[0:5], loc='lower right')
+    l = plt.legend(handles[0:len(cities)], labels[0:len(cities)], loc='lower right')
     # zero line
     [plt.axvline(_x, linewidth=0.5, color='k', linestyle='--') for _x in np.arange(-0.3, 0.4, 0.1)]
     plt.axvline(x=0, color='k', linestyle='-', linewidth = 2)

### 2019-04-10 code/analysis/analysis.py
updating figures

@@ -47,8 +47,13 @@ feature_names = {'lcov_11' : '% water','tree_mean':'% tree canopy','ndvi_mean':'
                 'svf_mean':'sky view factor','dsm_mean':'digital surface model',
                 'alb_mean':'albedo','dsm_sd':'dsm stand. dev.','nbdi_max':'max nbdi',
                 'tree_max':'max % tree can.','bldg':'% building area','pdens_mean':'pop. density',
-                'tree_min':'min % tree can.','svf_max':'max sky view factor'
+                'tree_min':'min % tree can.','svf_max':'max sky view factor',
+                'tree_sd': '% tree can. stand. dev.', 'nbdi_sd_sl':'nbdi surrounding stand. dev.',
+                'tree_sd_sl':'% tree can. surrounding stand. dev.',
+                'ndvi_sd': 'ndvi stand. dev'
                 }
+
+CORES_NUM = min(50,int(os.cpu_count()*3/4))
 
 def main():
     '''
@@ -112,7 +117,6 @@ def regressions(df, cities, sim_num, grid_size, do_par = False):
     df_city, response = prepare_lst_prediction(df_city)
     # conduct the holdout
     if do_par:
-        CORES_NUM = min(50,int(os.cpu_count()))
         Parallel(n_jobs=CORES_NUM)(delayed(single_regression)(df_city, response, grid_size, predict_quant, i) for i in range(sim_num))
     else:
         for i in range(sim_num):
@@ -335,7 +339,6 @@ def bootstrap_main(df, grid_size, boot_num, do_par = False):
     loop the bootstraps to calculate the partial_dependence
     '''
     if do_par:
-        CORES_NUM = min(50,int(os.cpu_count())-4)
         Parallel(n_jobs=CORES_NUM)(delayed(boot_pd)(df, grid_size, boot_index) for boot_index in range(boot_num))
     else:
         for boot_index in range(boot_num):
@@ -895,7 +898,7 @@ def plot_importance(results_swing, grid_size):
     # fig = plt.gcf()
     # fig.set_size_inches(15,20)
 
-    plt.savefig('fig/working/regression/variableImportance_{}.pdf'.format(grid_size), format='pdf', dpi=500, transparent=True)
+    plt.savefig('fig/report/variableImportance_{}.pdf'.format(grid_size), format='pdf', dpi=500, transparent=True)
     plt.show()
     plt.clf()
     return(feature_order)
[…]

### 2019-04-10 report/article.tex
about to add new figures into report

@@
 \journal{Remote Sensing of Environment}
+
+%% Packages
+\usepackage{tabu}
+\usepackage{breakurl}
+\usepackage{float}
 
@@
 %% Numbered
-%\bibliographystyle{model1-num-names}
+% \bibliographystyle{model1-num-names}
 
@@
 %% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
+% \usepackage{numcompress}\bibliographystyle{model3-num-names}
 
 %% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
+% \usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
 
@@
 %% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
+\usepackage{numcompress}\bibliographystyle{model6-num-names}
 
 %% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
+% \bibliographystyle{elsarticle-num}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 \cortext[mycorrespondingauthor]{Corresponding author}
-\ead[url]{www.tomlogan.co.nz}
-\ead{tomlogan@umich.edu}
+% \ead[url]{www.tomlogan.co.nz}
+\ead{tom.logan@canterbury.ac.nz}
 
@@
 \begin{abstract}
-Although heat waves and the urban heat island are nocturnal phenomenon, the drivers of land surface temperature during the night are poorly understood.
-Given that heat waves are the deadliest natural hazard and are expected to increase in frequency and severity, determining mitigation strategies is important.
-However, only recently has nocturnal imagery of cities become available by LandSat allowing nighttime land surface temperature to be analyzed.
-Furthermore, it remains unclear from existing studies the independent effects of potential drivers, and their relative importance.
+Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
+Determining mitigation strategies to reduce land surface temperature is essential given that heat waves, the deadliest natural hazard, are expected to increase in frequency and severity.
+% However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
+However, it remains unclear from existing studies the independent effects of potential drivers, and their relative importance.
 Here, we seek to answer the question: what are the effects and importance of drivers of land surface temperature during both day and night?
@@
 The results of diurnal and nocturnal analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
-Understanding whether the factors related to high urban land surface temperatures are consistent across US cities is important for climate adaptation planning and during heat wave mitigation strategies.
+Understanding whether the factors related to high urban land surface temperatures are consistent across US cities is important for climate adaptation planning and heat wave mitigation strategies.
 \end{abstract}
@@
 \section{Introduction}
-In a warming world, understanding the drivers of surface temperatures during both the day and night will aid in adapting cities to mitigate urban heat for the health and well being of communities.
-Of all the natural events, heat waves are among the deadliest and are likely to become longer and more frequent with climatic changes.
-Coupled with demographic shifts towards urban living, climate change puts significant impetus on reducing community exposure to heat waves.
-An understudied aspect of urban land surface temperature are factors related to nocturnal surface temperature.
-Nocturnal land surface temperature is important in heat wave mitigation given the urban heat island effect is most apparent during the night and the minimum nocturnal temperature is linked with heat stress and mortality.
+In a warming world, understanding the drivers of LST and the UHI will aid in adapting cities to mitigate urban heat for the health and well being of communities.
+The 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine 2008} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera 2016 in Wicki} }, is expected to become an annual occurrence by 2080 under current climate projections \cite{klinenberg2015heat}.
+Heat waves are exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
+The UHI results from heat captured during the day being released during the night increasing nocturnal temperatures \cite{Oke 1987, Landsberg1981, Rotach2005(bubble)}.
+Among other negative effects on a city's environmental quality, the UHI reduces the ability of people to cool off during the night and causes increased mortality \cite{Stone2006}.
+The UHI is studied in two ways.
+The first considers the air temperature using atmospheric station data \cite{Scott2016-lc, etc} while the other examines the land surface temperature (LST) using remote sensing \cite{imhoff, Peng2012-iy, Peng2018-cp, Zhou2014-wc, etc.}.
+Here, we study the LST because the spatial coverage enables comparative studies \cite{Hung2006-qy}.
+
+%In a warming world, understanding the drivers of LST and the UHI will aid in adapting cities to mitigate urban heat for the health and well being of communities.
+Due to the availability of satellite images most studies focus on daytime LST \cite{examples looking at day LST}.
+However, the impact of urban characteristics on LST allegedly differ during the night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
+Given that UHI is a nocturnal phenomena, it is important that the drivers of nocturnal LST are understood to enhance mitigation strategies.
+
+The existing studies are yet to provide a clear understanding of the the influence and relative importance of drivers on nocturnal LST \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc }.
+There are three aspects in which these studies can be enhanced: factors considered, multiple study regions, and their statistical techniques.
+The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{}.
+LandSat8 scenes during the night have recently been made available. 
+Existing studies have also been criticized for the explanatory variables they've used.
+There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
+Recent studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
+Beyond the 2D or 3D question, five categories of important variables have been identified: Green space, Water, Landscape, Albedo, and Socio-economic \cite{Peng2018-cp}.
+However, many studies do not capture all aspects, or worse, analyze only the univariate effect of variables \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
+Considering only single variables limits the understanding of variable effect given the interdependencies between drivers.
+
+The statistical approaches used in the studies limit their ability to explain the interdependencies between variables. 
+Almost all studies continue to use linear techniques \cite{Li2017-yl, etc.}. 
+This makes quantifying the independent effects of factors on LST and their relative importance difficult \cite{Peng2018-cp, Zhou2014-wc}.
+These studies also only assess their statistical accuracy using in-sample validation techniques \cite{something about validation e.g. Schmuli}.
+These result in higher reported values of explained variance (R$^2$ or other metric) because the trained models are not tested on unseen data.
+Furthermore, the variance explained is not high to begin with \cite{Chun2017-mm}.
+
+Finally, comparative studies between cities are necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}.
+
+Therefore, in this study we determine the influence and relative importance of the various drivers of LST, during both the night and day. 
+To achieve this conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the inter dependencies and relative importance between the factors.
 
 \section{Data and Methods}
+\subsection{Cities studied}
+
+Phoenix.
+* studied by Brazel 2007, Guhathakurta 2010.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1.2\linewidth]{fig/map/cities_lst.pdf}
+\caption{The land surface temperature in $^o$C of the cities studied.}
+\label{fig:cities}
+\end{center}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/joyplot_lst.png}
+    \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied.}
+    \label{fig:joy}
+\end{figure}
+
+\subsection{Land surface temperature}
+We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery as well as land cover and an air temperature observation (data sources provided in \ref{tab:data}.
+Band-10 digital-number data is converted to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}.
+We correct for emissivity using land cover data \cite{Alipour2003-gb}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}.
+Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station.
+This follows the process is described in \cite{Scott2016-lc} and is explicitly outlined in our open source code on Github\footnote{URL redacted for review}.
+
+To ameliorate the effect of shadows and other ephemeral changes we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
+
+- include scatter of night vs day LST at 100m - also change the symbols for colorblind - remove background
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/lst_night-vs-day.png}
+\caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution (UPDATE WITH 100m)}
+\label{fig:scatter_lst}
+\end{center}
+\end{figure}
+\subsection{Independent variables}
+- present correlation matrix of independent variables
+
+\textit{Albedo}.
+Albedo is a measure of the reflectivity or brightness of a surface.
+It is normalized between 0-1 where the darker surfaces are lower values.
+Albedo is calculated using the LandSat8 images using the algorithm described in \cite{Smith2010-nw, Liang2001-jd}.
+
+\textit{Building floor area}. The building floor area within the cell. This uses data released in 2018 by Microsoft where building footprints are estimated using areal images.
+
+\textit{Building height}.
+Building height is estimated using the building footprints and available lidar data following \cite{Chun2017-mm}.
+The mean lidar elevation within each building's footprint is subtracted by the mean digital elevation (topography) within the footprint to estimate the building's height.
+
+\textit{Elevation}.
+1/3 arc-second (~10m) bare-earth elevation (topography) data is available courtesy the U.S.
+Geological Survey.
+The mean, max, and min of elevation points within the grid cell is calculated.
+
+\textit{Surface elevation}. The surface elevation is determined from the lidar data. Surface elevation captures the natural and built features.
+
+\textit{Impervious surface percentage}.
+This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}.
+To generate the impervious surface area (ISA) LandSat data, NLCD land cover, and nighttime light imagery is used.
+The stable nighttime light intensity is only used to estimate the boundary of of urban areas.
+The Landsat images were converted to top-of-atmosphere reflectance.
+The data is provided at a 30m resolution and we calculate the mean, max, and min for each grid cell.
+
+\textit{Land cover}.
+Classification of land cover is described in \cite{Homer2015-ce}.
+However, as it is used in the calculation of emissivity when calculating the LST it cannot be used as an independent variable.
+The only landcover that we use is category 11, water.
+Cells with more than 20\% area classified as water are excluded from the data set.
+
+\textit{NBDI}.
+Normalized difference built-up index indicates the intensity of imperviousness \cite{Bhatti2014-ae}.
+It is calculated from satellite images as $$NDVI=\frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}}$$ where $B_{NIR}$ and $B_{SWIR}$ are the reflectances in the near-infrared and short-wave infrared bands respectively \cite{Alhawiti2016-wv}.
+Using LandSat8 imagery, this is $$NDVI=\frac{B6-B5}{B6+B5}$$ \cite{barsi2014}.
+% *Kaplan 2018, Peng 2018
+
+\textit{NDVI}.
+The normalized difference vegetation index (NDVI) measures green vegetation.
+It is calculated from satellite images as $$NDVI=\frac{B_{NIR}-B_{red}}{B_{NIR}+B_{red}}$$ where $B_{NIR}$ and $B_{red}$ are the reflectances in the near-infrared and red bands respectively \cite{Alhawiti2016-wv}.
+Using LandSat8 imagery, this is $$NDVI=\frac{B5-B4}{B5+B4}$$ \cite{barsi2014}.
+
+\textit{Nighttime light intensity}.
+Stable nighttime light intensity is available from the Defense Meteorological Satellite Program (DMSP) Operational Linescan System (OLS).
+They prepare the data to remove clouds and ephemeral light sources.
+We use 2013 (the most recently available) data of version 4 at a spatial resolution of 30 arc-seconds (1km at the equator) \cite{ngdc2013version}.
+The nighttime light intensity was found to be only in moderate agreement with density \cite{bagan2018}, meaning lights may indicate anthropogenic energy use during the night.
+However, it's lower resolution means that can not capture minor differences.
+We resampled the raster to a 30m resolution, as per the NLDC \cite{Homer2015-ce}, and for each grid calculated the mean, min, and max.
+
+\textit{Population density}.
+Density allegedly increases the LST \cite{Li2017-yl}, although this result may be the result of confounding with other factors.
+Population density calculated from the U.S.
+census at the block level.
+The most recent census was 2010, so that data is used as an estimator of where people reside in the evening.
+This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that it's centroid is contained within.
+
+\textit{Sky view factor}.
+Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
+We calculate SVF following using xxx and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$.
+We use a spatial resolution of x-m.
+% * Yuan 2011, Unger 2004.
+
+\textit{Tree canopy cover}.
+The percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}.
+The data provided is at an xx resolution.
+The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the product does not contain information used in the LST calculation (which requires radiance).
+% * Rogan 2013, Elmes 2017
+
+\subsection{Statistical analysis}
+The data was gridded into 500m square and 100m square cells for analysis.
+For each of the variables, the mean, maximum, and minimum were calculated.
+Spatially lagging, the mean of the surrounding cells (including diagonal) for each variable, was included as an additional independent variable to capture the spatial effects which may be present within the data.
+* test / train
+
+* statistical models
 
@@
 
-\subsection{Variables related to Land Surface Temperature}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/report/holdout_results_r2.pdf}
-\caption{The out-of-bag R$^2$ of the models when fitted on each city and used to predict a random sample from the same city.
-Out-of-bag R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
-This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.}
-\label{fig:intracity}
-\end{center}
-\end{figure}
-
-\subsection{Predicting Land Surface Temperature Between Cities}
-
-See figure \ref{fig:intercity}.
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/report/cities_holdout.pdf}
-\caption{The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/holdout_results_500.pdf}
+    \caption{
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
 OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
@@
 Note that the R$^2$ axis is truncated at -1, although the multiple linear regression for diurnal land surface temperature tested on Phoenix has an OOB R$^2$ of -9.
-This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.}
-\label{fig:intercity}
-\end{center}
+This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.
+    }
+    \label{fig:holdout_results}
+\end{figure}
+
+\subsection{Variables related to Land Surface Temperature}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_night.png}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_night}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_day.png}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_day}
 \end{figure}
@@
 
-% % \section*{References}
+\section*{References}
 
-% \bibliography{mybibfile}
+\bibliography{mybibfile}
 
 \appendix
-\section{2013 Land Surface Temperature Image}
+\section{Data sources}
+
+\begin{tabu}to \textwidth{ X[l]  X[c]  X[c] X[l] X[l] }
+ \hline
+ Data Provider & Data Type & Data Date & Description & URL \\
+ \hline
+U.S. Geological Survey  & Raster  & 2013-2017 &
+    Landsat 8 day and night satellite imagery & \burl{https://earthexplorer.usgs.gov/} \\
+Microsoft  & Polygon  & 2018 & Building footprint polygons for the US &
+    \burl{https://github.com/Microsoft/USBuildingFootprints} \\
+Defense Meteorological Satellite Program  & Raster  & 2013 &
+    Stable nighttime light intensity & \burl{https://www.ngdc.noaa.gov/eog/dmsp/downloadV4composites.html} \\
+Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
+    Land cover & \burl{https://www.mrlc.gov} \\
+Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
+    Percent developed imperviousness & \burl{https://viewer.nationalmap.gov} \\
+Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
+    Percent tree canopy cover & \burl{https://viewer.nationalmap.gov} \\
+U.S. Geological Survey  & Raster  & 2015 & 1/3 arc-second elevation &
+    \burl{https://nationalmap.gov/3DEP/3dep_prodserv.html} \\
+U.S. National Oceanic and Atmospheric Administration  & Lidar  & 2014 & Point cloud of surface elevation &
+    \burl{https://coast.noaa.gov/htdata/lidar2_z/geoid12b/data/6377/} \\
+IPUMS NHGIS  & Area Level  & 2010 &
+    Block-level population from the US census & \cite{nhgis}\\
+\hline
+\label{tab:data}
+\end{tabu}
 

### 2019-04-10 code/analysis/analysis.py
updating figures

@@ -961,9 +961,9 @@ def scatter_lst(df, cities, grid_size):
         df_city = df.loc[df['city']==city]
         plt.scatter(df_city['lst_day_mean'], df_city['lst_night_mean'], label = city, alpha = 0.5)
     plt.legend(loc='lower right')
-    plt.xlabel('Day LST ($^o$C)')
-    plt.ylabel('Night LST ($^o$C)')
-    plt.text(20, 40,'Correlation = {0:.2f}'.format(df_city['lst_day_mean'].corr(df_city['lst_night_mean'])), ha='left', va='top')
+    plt.xlabel('diurnal LST ($^o$C)')
+    plt.ylabel('nocturnal LST ($^o$C)')
+    plt.text(20, 40,'correlation = {0:.2f}'.format(df_city['lst_day_mean'].corr(df_city['lst_night_mean'])), ha='left', va='top')
     plt.savefig('fig/report/lst_night-vs-day_{}.pdf'.format(grid_size), format='pdf', dpi=300, transparent=True)
     plt.show()
     plt.clf()
@@ -979,7 +979,7 @@ def joyplot_lst(df, grid_size):
     df1 = df1.dropna(axis=0, how='any')
     # with plt.style.context('fivethirtyeight'):
     fig, axes = joypy.joyplot(df1, by='city', ylim='own',legend=True)
-    plt.xlabel('Land Surface Temperature ($^{o}$C)')
+    plt.xlabel('land surface temperature ($^{o}$C)')
     plt.savefig('fig/report/joyplot_lst_{}.pdf'.format(grid_size), format='pdf', dpi=300, transparent=True)
     plt.show()
     plt.clf()

### 2019-04-10 report/article.tex
Updates from Overleaf

@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1.2\linewidth]{fig/map/cities_lst.pdf}
-\caption{The land surface temperature in $^o$C of the cities studied.}
-\label{fig:cities}
-\end{center}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/map_nocturnal_lst.pdf}
+    \caption{The gridded nocturnal land surface temperature in $^o$C of the cities studied. This is change from the mean (average) for each city.}
+    \label{fig:map}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/joyplot_lst.png}
+    \includegraphics[width=\linewidth]{fig/report/joyplot_lst.pdf}
     \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied.}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{fig/report/lst_night-vs-day.png}
-\caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution (UPDATE WITH 100m)}
+\includegraphics[width=\textwidth]{fig/report/lst_night-vs-day_500.pdf}
+\caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution}
 \label{fig:scatter_lst}

### 2019-04-16 code/analysis/analysis.py
updating plots

@@ -755,7 +755,7 @@ fig_transparency = False
 width_1col = 8.7/2.54
 width_2col = 17.8/2.54
 golden_mean = (sqrt(5)-1.0)/2.0    # Aesthetic ratio
-height_1c = width_1col/golden_mean
+height_1c = width_1col*golden_mean
 height_2c = width_2col/golden_mean
 # font size
 font_size = 11
@@ -872,7 +872,7 @@ def plot_holdouts(loss, grid_size):
             ax.set_ylabel('out-of-bag R$^2$')
             ax.set_xlabel('')
             # plt.gca().invert_yaxis()
-    plt.savefig('fig/working/regression/holdout_results_{}.pdf'.format(grid_size), format='pdf', dpi=500, transparent=True)
+    plt.savefig('fig/report/holdout_results_{}.pdf'.format(grid_size), format='pdf', dpi=500, transparent=True)
     plt.show()
     plt.clf()
 
@@ -887,6 +887,7 @@ def plot_importance(results_swing, grid_size):
     results_swing = results_swing.replace(model_names)
 
     # plot
+    plt.figure(figsize=(width_2col, height_2c))
     g = sns.factorplot(x='swing', y='independent', hue='dependent',
                         data=results_swing, kind='bar', col='model',
                         order = feature_order, hue_order=['lst_night_mean','lst_day_mean'])
@@ -956,6 +957,7 @@ def scatter_lst(df, cities, grid_size):
     # scatter plot thermal radiance against land surface, colored by city
     # bmap = brewer2mpl.get_map('Paired','Qualitative',4).mpl_colors
     # with plt.style.context('fivethirtyeight'):
+    plt.figure(figsize=(width_1col, height_1c))
     for i in range(len(cities)):
         city = cities[i]
         df_city = df.loc[df['city']==city]
@@ -964,7 +966,7 @@ def scatter_lst(df, cities, grid_size):
     plt.xlabel('diurnal LST ($^o$C)')
     plt.ylabel('nocturnal LST ($^o$C)')
     plt.text(20, 40,'correlation = {0:.2f}'.format(df_city['lst_day_mean'].corr(df_city['lst_night_mean'])), ha='left', va='top')
-    plt.savefig('fig/report/lst_night-vs-day_{}.pdf'.format(grid_size), format='pdf', dpi=300, transparent=True)
[…]

### 2019-04-16 report/article.tex
Updates from Overleaf

@@
-\documentclass[]{elsarticle}
+% \documentclass[sort&compress]{elsarticle}
+\documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
 %\documentclass[review]{elsarticle}
@@
 %% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
+% \bibliographystyle{model2-names.bst}\biboptions{authoryear}
 
@@
 %% AMA style
-\usepackage{numcompress}\bibliographystyle{model6-num-names}
+% \usepackage{numcompress}
+% \bibliographystyle{model6-num-names}
 
 %% `Elsevier LaTeX' style
-% \bibliographystyle{elsarticle-num}
+\bibliographystyle{elsarticle-num}
+
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\title{Understanding the drivers of diurnal and nocturnal urban land surface temperature}
+\title{Urban characteristics associated with diurnal and nocturnal land surface temperature}
 
@@
 Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
-Determining mitigation strategies to reduce land surface temperature is essential given that heat waves, the deadliest natural hazard, are expected to increase in frequency and severity.
+Understanding these drivers is necessary for reducing land surface temperature and mitigating heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
 % However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
-However, it remains unclear from existing studies the independent effects of potential drivers, and their relative importance.
-Here, we seek to answer the question: what are the effects and importance of drivers of land surface temperature during both day and night?
+However, the independent effects and relative importance of potential drivers remain unclear from existing studies.
+We seek to answer the question: what are the effects and importance of drivers of land surface temperature during both day and night?
 To do this, we analyze the urban land surface temperature in four cities across the United States.
-We include factors related to vegetation, water, the built-environment, and topography and control them using nonlinear statistical methods which allow for their independent effects to be determined.
-The results of diurnal and nocturnal analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
-Understanding whether the factors related to high urban land surface temperatures are consistent across US cities is important for climate adaptation planning and heat wave mitigation strategies.
+In our analysis, we include factors related to vegetation, water, the built-environment, and topography and control them using nonlinear statistical methods which allow for their independent effects to be assessed.
+The effects from the diurnal and nocturnal analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
+Understanding the relationships influencing both nocturnal and diurnal land surface temperature will improve climate adaptation planning and heat wave mitigation.
 \end{abstract}
@@
 
-\begin{figure}[H]
+\begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/map_nocturnal_lst.pdf}
+    \includegraphics[width=\textwidth]{fig/report/map_nocturnal_lst.pdf}
     \caption{The gridded nocturnal land surface temperature in $^o$C of the cities studied. This is change from the mean (average) for each city.}
     \label{fig:map}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[H]
+\begin{figure}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/joyplot_lst.pdf}
+    \includegraphics[width=\linewidth]{fig/report/joyplot_lst_500.pdf}
     \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied.}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{fig/report/lst_night-vs-day_500.pdf}
+\includegraphics[width=\linewidth]{fig/report/lst_night-vs-day_500.pdf}
 \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution}
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \label{fig:holdout_results}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{fig/report/variableImportance_500.pdf}
+\caption{The variable importance for the regression during the night (to the left) and the day (to the right).
+Variable importance indicates the amount that each variable improves the performance of the regression.
+In this case there is generally little difference between the importance of day and night variables.}
+\label{fig:importance}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_night.png}
+    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_night_500.png}
     \caption{
@@
     \label{fig:pdp_night}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_day.png}
+    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_day_500.png}
     \caption{
@@
     \label{fig:pdp_day}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_night_500.pdf}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_2dnight}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_day_500.pdf}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_2dday}
+\end{figure*}
 
@@
 Regarding Fig \ref{fig:pdp}, I think I will remove the ``all" line, because in cases such as $alb_mean_min_sl$ (which is the minimum albedo in the surrounding cells (sl = spatial lag)) the values between the cities don't overlap so I think it becomes a proxy for other differences between the city when regressed together.
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{fig/report/variable_importance_selected.pdf}
-\caption{The variable importance for the regression during the night (to the left) and the day (to the right).
-Variable importance indicates the amount that each variable improves the performance of the regression.
-In this case there is generally little difference between the importance of day and night variables.}
-\label{fig:importance}
-\end{center}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[height=\textheight]{fig/report/partial_dependence.pdf}
-\caption{Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.}
-\label{fig:pdp}
-\end{center}
-\end{figure}
 
@@
 
-\bibliography{mybibfile}
+\bibliography{mybibfile.bib}
 
+\newpage
+\onecolumn
 \appendix
+
 \section{Data sources}
@@
 \section{Additional model results}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/holdout_results_100.pdf}
+    \caption{
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
+OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1.
+Good models have MAE near 0.
+The gradient boosted regression forest generally outperforms the other models, except when predicting Detroit.
+Note that the R$^2$ axis is truncated at -1, although the multiple linear regression for diurnal land surface temperature tested on Phoenix has an OOB R$^2$ of -9.
+This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.
+    }
+    \label{fig:holdout_results}
+\end{figure}
+
 \begin{figure}[h]
 \begin{center}
-\includegraphics[width=\textwidth]{fig/report/holdout_results_mae.pdf}
-\caption{The Mean Absolute Error (MAE) in $^oC$ of the models when fitted on each city and used to predict a random sample from the same city.
-Better models have lower MAE.
-This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.
-}
-\label{fig:cityholdout_errors}
+\includegraphics[width=\textwidth]{fig/report/variableImportance_100.pdf}
+\caption{The variable importance for the regression during the night (to the left) and the day (to the right).
+Variable importance indicates the amount that each variable improves the performance of the regression.
+In this case there is generally little difference between the importance of day and night variables.}
+\label{fig:importance}
 \end{center}
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_night_100.png}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_night}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_day_100.png}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_day}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_night_100.pdf}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_2dnight}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_day_100.pdf}
+    \caption{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
+This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    }
+    \label{fig:pdp_2dday}
+\end{figure}
+
 \end{document}

### 2019-04-16 code/analysis/analysis.py
updating the figures

@@ -957,7 +957,7 @@ def scatter_lst(df, cities, grid_size):
     # scatter plot thermal radiance against land surface, colored by city
     # bmap = brewer2mpl.get_map('Paired','Qualitative',4).mpl_colors
     # with plt.style.context('fivethirtyeight'):
-    plt.figure(figsize=(width_1col, height_1c))
+    plt.figure(figsize=(width_2col, height_2c))
     for i in range(len(cities)):
         city = cities[i]
         df_city = df.loc[df['city']==city]
@@ -981,7 +981,7 @@ def joyplot_lst(df, grid_size):
     df1 = df1.dropna(axis=0, how='any')
     # with plt.style.context('fivethirtyeight'):
     fig, axes = joypy.joyplot(df1, by='city', ylim='own',legend=True,
-                            figsize=(width_1col, height_1c))
+                            figsize=(width_2col, height_2c))
     plt.xlabel('land surface temperature ($^{o}$C)')
     plt.savefig('fig/report/joyplot_lst_{}.pdf'.format(grid_size), format='pdf', dpi=300, transparent=True)
     plt.show()

### 2019-04-16 code/analysis/analysis.py
updating figures

@@ -756,7 +756,7 @@ width_1col = 8.7/2.54
 width_2col = 17.8/2.54
 golden_mean = (sqrt(5)-1.0)/2.0    # Aesthetic ratio
 height_1c = width_1col*golden_mean
-height_2c = width_2col/golden_mean
+height_2c = width_2col*golden_mean
 # font size
 font_size = 11
 dpi = 500

### 2019-04-16 code/analysis/analysis.py
updating figures

@@ -756,7 +756,7 @@ width_1col = 8.7/2.54
 width_2col = 17.8/2.54
 golden_mean = (sqrt(5)-1.0)/2.0    # Aesthetic ratio
 height_1c = width_1col*golden_mean
-height_2c = width_2col*golden_mean
+height_2c = width_2col*golden_mean/2
 # font size
 font_size = 11
 dpi = 500
@@ -860,6 +860,7 @@ def plot_holdouts(loss, grid_size):
     '''
     plot boxplots of holdouts
     '''
+    plt.figure(figsize=(width_2col, height_2c))
     g = sns.catplot(y="error", x="time_of_day", hue="model", col = "error_metric", data=loss, sharey = False, kind="box")
     g.set_titles('')
     for i, ax in enumerate(g.axes.flat): # set every-other axis for testing purposes
@@ -887,21 +888,34 @@ def plot_importance(results_swing, grid_size):
     results_swing = results_swing.replace(model_names)
 
     # plot
-    plt.figure(figsize=(width_2col, height_2c))
-    g = sns.factorplot(x='swing', y='independent', hue='dependent',
-                        data=results_swing, kind='bar', col='model',
-                        order = feature_order, hue_order=['lst_night_mean','lst_day_mean'])
-    g.set_axis_labels("variable importance (swing)", "")
-    g.set_titles("{col_name}")
-    new_labels = ['nocturnal','diurnal']
-    for t, l in zip(g._legend.texts, new_labels): t.set_text(l)
+    # font_size = 15
+    with sns.plotting_context("paper", font_scale=1.5):
+    # sns.set_context("paper", rc={"font.size":font_size,"axes.titlesize":font_size,"axes.labelsize":font_size})
+    # plt.figure(figsize=(width_2col, height_2c))
+        g = sns.factorplot(x='swing', y='independent', hue='dependent',
+                            data=results_swing, kind='bar', col='model',
+                            order = feature_order,
+                            hue_order=['lst_night_mean','lst_day_mean'],
+                            col_order=['random forest','gradient boosted trees',
+                                        'multivariate adaptive spline (mars)'],
+                                        # 'generalized additive (gam)',
+                                        # 'multivariate linear'],
+                            col_wrap = 3
+                            )
 
-    # fig = plt.gcf()
-    # fig.set_size_inches(15,20)
+        g.set_axis_labels("variable influence", "")
[…]

### 2019-04-16 code/analysis/analysis.py
updating figures

@@ -756,7 +756,7 @@ width_1col = 8.7/2.54
 width_2col = 17.8/2.54
 golden_mean = (sqrt(5)-1.0)/2.0    # Aesthetic ratio
 height_1c = width_1col*golden_mean
-height_2c = width_2col*golden_mean/2
+height_2c = width_2col*golden_mean
 # font size
 font_size = 11
 dpi = 500
@@ -861,7 +861,10 @@ def plot_holdouts(loss, grid_size):
     plot boxplots of holdouts
     '''
     plt.figure(figsize=(width_2col, height_2c))
-    g = sns.catplot(y="error", x="time_of_day", hue="model", col = "error_metric", data=loss, sharey = False, kind="box")
+    g = sns.catplot(y="error", x="time_of_day", hue="model",
+                    col = "error_metric", data=loss, sharey = False,
+                    kind="box",
+                    height = height_2c, aspect = width_2col/height_2c)
     g.set_titles('')
     for i, ax in enumerate(g.axes.flat): # set every-other axis for testing purposes
         if i%2==1:

### 2019-04-16 code/analysis/analysis.py
updating influence figures

@@ -757,6 +757,7 @@ width_2col = 17.8/2.54
 golden_mean = (sqrt(5)-1.0)/2.0    # Aesthetic ratio
 height_1c = width_1col*golden_mean
 height_2c = width_2col*golden_mean
+aspect_2c = width_2col/height_2c
 # font size
 font_size = 11
 dpi = 500
@@ -864,7 +865,7 @@ def plot_holdouts(loss, grid_size):
     g = sns.catplot(y="error", x="time_of_day", hue="model",
                     col = "error_metric", data=loss, sharey = False,
                     kind="box",
-                    height = height_2c, aspect = width_2col/height_2c)
+                    height = height_2c, aspect = aspect_2c)
     g.set_titles('')
     for i, ax in enumerate(g.axes.flat): # set every-other axis for testing purposes
         if i%2==1:
@@ -892,7 +893,8 @@ def plot_importance(results_swing, grid_size):
 
     # plot
     # font_size = 15
-    with sns.plotting_context("paper", font_scale=1.5):
+    font_scale = 2#1.75
+    with sns.plotting_context("paper", font_scale=font_scale):
     # sns.set_context("paper", rc={"font.size":font_size,"axes.titlesize":font_size,"axes.labelsize":font_size})
     # plt.figure(figsize=(width_2col, height_2c))
         g = sns.factorplot(x='swing', y='independent', hue='dependent',
@@ -900,14 +902,14 @@ def plot_importance(results_swing, grid_size):
                             order = feature_order,
                             hue_order=['lst_night_mean','lst_day_mean'],
                             col_order=['random forest','gradient boosted trees',
-                                        'multivariate adaptive spline (mars)'],
-                                        # 'generalized additive (gam)',
-                                        # 'multivariate linear'],
-                            col_wrap = 3
+                                        'multivariate adaptive spline (mars)',
+                                        'generalized additive (gam)',
+                                        'multivariate linear'],
+                            # col_wrap = 3
                             )
 
         g.set_axis_labels("variable influence", "")
-        g.set_titles("{col_name}",size=font_size)
+        g.set_titles("{col_name}",size=font_size*1.75)
         # g.tick_params(labelsize=font_size)
 
         new_labels = ['nocturnal','diurnal']

### 2019-04-16 code/analysis/analysis.py
updating figures

@@ -893,7 +893,7 @@ def plot_importance(results_swing, grid_size):
 
     # plot
     # font_size = 15
-    font_scale = 2#1.75
+    font_scale = 2.5#1.75
     with sns.plotting_context("paper", font_scale=font_scale):
     # sns.set_context("paper", rc={"font.size":font_size,"axes.titlesize":font_size,"axes.labelsize":font_size})
     # plt.figure(figsize=(width_2col, height_2c))
@@ -909,7 +909,7 @@ def plot_importance(results_swing, grid_size):
                             )
 
         g.set_axis_labels("variable influence", "")
-        g.set_titles("{col_name}",size=font_size*1.75)
+        g.set_titles("{col_name}",size=font_size*2)
         # g.tick_params(labelsize=font_size)
 
         new_labels = ['nocturnal','diurnal']

### 2019-04-16 report/article.tex
Updates from Overleaf

@@
 In a warming world, understanding the drivers of LST and the UHI will aid in adapting cities to mitigate urban heat for the health and well being of communities.
-The 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine 2008} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera 2016 in Wicki} }, is expected to become an annual occurrence by 2080 under current climate projections \cite{klinenberg2015heat}.
+The 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine 2008} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera 2016 in Wicki} }, is expectAZed to become an annual occurrence by 2080 under current climate projections \cite{klinenberg2015heat}.
 Heat waves are exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
@@
 \begin{center}
-\includegraphics[width=\linewidth]{fig/report/lst_night-vs-day_500.pdf}
+\includegraphics[width=\linewidth]{fig/report/lst_night-vs-day_500.png}
 \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution}

### 2019-04-16 code/analysis/analysis.py
updating influence figure

@@ -41,8 +41,10 @@ RANDOM_SEED = 3201
 
 # other constants
 city_names = {'bal':'baltimore','por':'portland','phx':'phoenix','det':'detroit'}
-model_names = {'rf':'random forest','mlr':'multivariate linear','gam':'generalized additive (gam)',
-                'gbrt':'gradient boosted trees','mars':'multivariate adaptive spline (mars)'}
+model_names = {'rf':'random forest','mlr':'multivariate linear',
+                'gam':'generalized additive\n(gam)',
+                'gbrt':'gradient boosted\ntrees',
+                'mars':'multivariate adaptive\nspline (mars)'}
 feature_names = {'lcov_11' : '% water','tree_mean':'% tree canopy','ndvi_mean':'ndvi',
                 'svf_mean':'sky view factor','dsm_mean':'digital surface model',
                 'alb_mean':'albedo','dsm_sd':'dsm stand. dev.','nbdi_max':'max nbdi',
@@ -901,10 +903,11 @@ def plot_importance(results_swing, grid_size):
                             data=results_swing, kind='bar', col='model',
                             order = feature_order,
                             hue_order=['lst_night_mean','lst_day_mean'],
-                            col_order=['random forest','gradient boosted trees',
-                                        'multivariate adaptive spline (mars)',
-                                        'generalized additive (gam)',
+                            col_order=['random forest','gradient boosted\ntrees',
+                                        'multivariate adaptive\nspline (mars)',
+                                        'generalized additive\n(gam)',
                                         'multivariate linear'],
+                            height = height_2c*2
                             # col_wrap = 3
                             )
 

### 2019-04-16 code/analysis/analysis.py
updating holdout figure

@@ -863,25 +863,27 @@ def plot_holdouts(loss, grid_size):
     '''
     plot boxplots of holdouts
     '''
-    plt.figure(figsize=(width_2col, height_2c))
-    g = sns.catplot(y="error", x="time_of_day", hue="model",
-                    col = "error_metric", data=loss, sharey = False,
-                    kind="box",
-                    height = height_2c, aspect = aspect_2c)
-    g.set_titles('')
-    for i, ax in enumerate(g.axes.flat): # set every-other axis for testing purposes
-        if i%2==1:
-            ax.set_ylim(0,5)
-            ax.set_ylabel('mean absolute error ($^o$C)')
-            ax.set_xlabel('')
-        elif i%2==0:
-            ax.set_ylim(0.5,1)
-            ax.set_ylabel('out-of-bag R$^2$')
-            ax.set_xlabel('')
-            # plt.gca().invert_yaxis()
-    plt.savefig('fig/report/holdout_results_{}.pdf'.format(grid_size), format='pdf', dpi=500, transparent=True)
-    plt.show()
-    plt.clf()
+    font_scale = 1.75
+    with sns.plotting_context("paper", font_scale=font_scale):
+        plt.figure(figsize=(width_2col, height_2c))
+        g = sns.catplot(y="error", x="time_of_day", hue="model",
+                        col = "error_metric", data=loss, sharey = False,
+                        kind="box",
+                        height = height_2c, aspect = aspect_2c)
+        g.set_titles('')
+        for i, ax in enumerate(g.axes.flat): # set every-other axis for testing purposes
+            if i%2==1:
+                ax.set_ylim(0,5)
+                ax.set_ylabel('mean absolute error ($^o$C)',size=font_size*1.5)
+                ax.set_xlabel('')
+            elif i%2==0:
+                ax.set_ylim(0.5,1)
+                ax.set_ylabel('out-of-bag R$^2$',size=font_size*1.5)
+                ax.set_xlabel('')
+                # plt.gca().invert_yaxis()
+        plt.savefig('fig/report/holdout_results_{}.pdf'.format(grid_size), format='pdf', dpi=500, transparent=True)
+        plt.show()
+        plt.clf()
 
 def plot_importance(results_swing, grid_size):
     '''
@@ -907,7 +909,8 @@ def plot_importance(results_swing, grid_size):
                                         'multivariate adaptive\nspline (mars)',
                                         'generalized additive\n(gam)',
                                         'multivariate linear'],
-                            height = height_2c*2
[…]

### 2019-04-16 code/analysis/analysis.py
updating var influence

@@ -897,7 +897,7 @@ def plot_importance(results_swing, grid_size):
 
     # plot
     # font_size = 15
-    font_scale = 2.5#1.75
+    font_scale = 3#1.75
     with sns.plotting_context("paper", font_scale=font_scale):
     # sns.set_context("paper", rc={"font.size":font_size,"axes.titlesize":font_size,"axes.labelsize":font_size})
     # plt.figure(figsize=(width_2col, height_2c))
@@ -915,7 +915,7 @@ def plot_importance(results_swing, grid_size):
                             )
 
         g.set_axis_labels("variable influence", "")
-        g.set_titles("{col_name}",size=font_size*2)
+        g.set_titles("{col_name}",size=font_size*2.5)
         # g.tick_params(labelsize=font_size)
 
         new_labels = ['nocturnal','diurnal']

### 2019-04-16 code/analysis/analysis.py
updating scatter figure

@@ -987,10 +987,11 @@ def scatter_lst(df, cities, grid_size):
         city = cities[i]
         df_city = df.loc[df['city']==city]
         plt.scatter(df_city['lst_day_mean'], df_city['lst_night_mean'], label = city, alpha = 0.5)
+        print(df_city['lst_day_mean'].corr(df_city['lst_night_mean']))
     plt.legend(loc='lower right')
-    plt.xlabel('diurnal LST ($^o$C)')
-    plt.ylabel('nocturnal LST ($^o$C)')
-    plt.text(20, 40,'correlation = {0:.2f}'.format(df_city['lst_day_mean'].corr(df_city['lst_night_mean'])), ha='left', va='top')
+    plt.xlabel('$\Delta$mean diurnal LST ($^o$C)')
+    plt.ylabel('$\Delta$mean nocturnal LST ($^o$C)')
+    plt.text(-20, 5,'correlation = {0:.2f}'.format(df['lst_day_mean'].corr(df['lst_night_mean'])), ha='left', va='top')
     plt.savefig('fig/report/lst_night-vs-day_{}.png'.format(grid_size), format='png', dpi=300, transparent=True)
     plt.show()
     plt.clf()

### 2019-04-19 code/analysis/analysis.py
update swing and create tree scatter

@@ -330,6 +330,7 @@ def calc_swing(results_pd, grid_size):
             swing = swing.to_frame('swing').reset_index()
             swing['model'] = m
             swing['dependent'] = h
+            swing['raw'] = model_range['mean'].values
             # save
             results_swing = results_swing.append(swing, ignore_index=True)
         # save results
@@ -1069,6 +1070,29 @@ def plot_2d_partialdependence(regressor, time_of_day, grid_size, df_x):
         axes[plot_row, left_right].set_ylabel(feature_names[feature_y])
     # save
     plt.savefig('fig/report/pdp_2d_{}_{}.pdf'.format(time_of_day,grid_size), format='pdf', dpi=500, transparent=True)
+    plt.show()
+    plt.clf()
+
+def scatter_tree_imp(df, cities, grid_size):
+    '''
+    scatter lst night vs day
+    '''
+    df = df.replace(city_names)
+    cities = [city_names[i] for i in cities]
+    # scatter plot thermal radiance against land surface, colored by city
+    # bmap = brewer2mpl.get_map('Paired','Qualitative',4).mpl_colors
+    # with plt.style.context('fivethirtyeight'):
+    plt.figure(figsize=(width_2col, height_2c))
+    for i in range(len(cities)):
+        city = cities[i]
+        df_city = df.loc[df['city']==city]
+        plt.scatter(df_city['imp_mean'], df_city['tree_mean'], label = city, alpha = 0.5)
+        print(df_city['imp_mean'].corr(df_city['tree_mean']))
+    plt.legend(loc='lower right')
+    plt.xlabel('\% impervious surface')
+    plt.ylabel('\% tree canopy cover')
+    plt.text(20, 80,'correlation = {0:.2f}'.format(df['imp_mean'].corr(df['tree_mean'])), ha='left', va='top')
+    plt.savefig('fig/report/imp_v_tree_{}.png'.format(grid_size), format='png', dpi=300, transparent=True)
     plt.show()
     plt.clf()
 

### 2019-04-19 code/analysis/analysis.py
analyzing the max as well as the mean

@@ -103,7 +103,7 @@ def import_data(grid_size, selected_vars = True):
         df = pd.read_csv('data/data_vif_{}.csv'.format(grid_size))
         df = df.drop('Unnamed: 0', axis=1)
     else:
-        df = pd.read_csv('data/data_regressions_{}_20190405.csv'.format(grid_size))
+        df = pd.read_csv('data/data_regressions_{}_20190419.csv'.format(grid_size))
         df = df.drop('Unnamed: 0', axis=1)
     return(df)
 
@@ -199,7 +199,7 @@ def prepare_lst_prediction(df):
     thermal radiance values
     '''
     # drop lst
-    lst_vars = ['lst_day_mean','lst_night_mean']
+    lst_vars = ['lst_day_mean','lst_night_mean','lst_night_max','lst_day_max']
     lst_mean = df[lst_vars].copy()
     df = df.drop(lst_vars, axis=1)
 
@@ -225,9 +225,13 @@ def define_response_lst(y_train, y_test):
     y = {}
     y['day_train'] = y_train['lst_day_mean']
     y['night_train'] = y_train['lst_night_mean']
+    y['nightmax_train'] = y_train['lst_night_max']
+    y['daymax_train'] = y_train['lst_day_max']
     # test
     y['day_test'] = y_test['lst_day_mean']
     y['night_test'] = y_test['lst_night_mean']
+    y['nightmax_test'] = y_test['lst_night_max']
+    y['daymax_test'] = y_test['lst_day_max']
     return(y)
 
 
@@ -244,7 +248,7 @@ def calculate_partial_dependence(df, grid_size, boot_index = None):
     df_reference = df.copy()
     feature_resolution = 25
     # loop day and night
-    for h in ['lst_day_mean', 'lst_night_mean']:
+    for h in ['lst_day_mean', 'lst_night_mean', 'lst_night_max','lst_day_max']:
         print(h)
         ###
         # fit models
@@ -391,28 +395,17 @@ def regression_null(y, city, predict_quant, loss):
     # predict the model
     predict_day = np.ones(len(y['day_test'])) * np.mean(y['day_train'])
     predict_night = np.ones(len(y['night_test'])) * np.mean(y['night_train'])
+    predict_nightmax = np.ones(len(y['nightmax_test'])) * np.mean(y['nightmax_train'])
+    predict_daymax = np.ones(len(y['daymax_test'])) * np.mean(y['daymax_train'])
 
     # plot predict vs actual
[…]

### 2019-04-19 code/analysis/prepare_data.py

@@ -8,11 +8,11 @@ import pandas as pd
 import numpy as np
 import time
 import code
-scale = False
+scale = True
 def main(scale=True):
 
     cities = ['bal','det','phx','por']
-    grid_size = 100
+    grid_size = 500
     # init all city dataframe
     df = pd.DataFrame()
     add_index = 0
@@ -59,6 +59,8 @@ def scaling_city(df_city, scale):
     # drop any values where the landsurface temperature is inf or nan
     df_city = df_city[np.isfinite(df_city['lst_day_mean'])]
     df_city = df_city[np.isfinite(df_city['lst_night_mean'])]
+    df_city = df_city[np.isfinite(df_city['lst_night_max'])]
+    df_city = df_city[np.isfinite(df_city['lst_day_max'])]
 
     ## for phoenix - remove large rural areas outside of lidar zone
     if df_city.city.iloc[100] == 'phx':
@@ -107,7 +109,7 @@ def scaling_city(df_city, scale):
 
     ### LST ###
     # drop the lst vars except night and day mean
-    vars_response = ['lst_day_mean','lst_night_mean']
+    vars_response = ['lst_day_mean','lst_night_mean','lst_night_max','lst_day_max']
     vars_lst = [i for i in vars_all if 'lst' in i]
     # remove vars to keep
     vars_lst_drop = [x for x in vars_lst if x not in vars_response]

### 2019-04-21 code/analysis/analysis.py
adding max temp to results

@@ -405,7 +405,7 @@ def regression_null(y, city, predict_quant, loss):
     mae_day, mae_night,r2_day,r2_night,mae_daymax,mae_nightmax,r2_daymax,r2_nightmax = calculate_errors(y, predict_day, predict_night, predict_daymax, predict_nightmax)
 
     # record results
-    record_result(loss, city, model, mae_day, mae_night, r2_night, r2_day, mae_daymax, mae_nightmax, r2_nightmax, r2_daymax)
+    loss = record_result(loss, city, model, mae_day, mae_night, r2_night, r2_day, mae_daymax, mae_nightmax, r2_nightmax, r2_daymax)
 
     return(loss)
 
@@ -439,7 +439,7 @@ def regression_gradientboost(X_train, y, X_test, city, predict_quant, loss):
 
 
     # record results
-    record_result(loss, city, model, mae_day, mae_night, r2_night, r2_day, mae_daymax, mae_nightmax, r2_nightmax, r2_daymax)
+    loss = record_result(loss, city, model, mae_day, mae_night, r2_night, r2_day, mae_daymax, mae_nightmax, r2_nightmax, r2_daymax)
 
     return(loss)
 
@@ -473,7 +473,7 @@ def regression_linear(X_train, y, X_test, city, predict_quant, loss):
     mae_day, mae_night,r2_day,r2_night,mae_daymax,mae_nightmax,r2_daymax,r2_nightmax = calculate_errors(y, predict_day, predict_night, predict_daymax, predict_nightmax)
 
     # record results
-    record_result(loss, city, model, mae_day, mae_night, r2_night, r2_day, mae_daymax, mae_nightmax, r2_nightmax, r2_daymax)
+    loss = record_result(loss, city, model, mae_day, mae_night, r2_night, r2_day, mae_daymax, mae_nightmax, r2_nightmax, r2_daymax)
 
     return(loss)
 
@@ -495,6 +495,8 @@ def regression_randomforest(X_train, y, X_test, city, predict_quant, loss):
     # predict the model
     predict_day = reg_day.predict(X_test)
     predict_night = reg_night.predict(X_test)
+    predict_daymax = reg_daymax.predict(X_test)
+    predict_nightmax = reg_nightmax.predict(X_test)
 
     # plot predict vs actual
     plot_actualVpredict(y, predict_day, predict_night, 'gbrf', city, predict_quant)
@@ -503,7 +505,7 @@ def regression_randomforest(X_train, y, X_test, city, predict_quant, loss):
[…]

### 2019-04-21 report/article.tex

@@
 
-\title{Urban characteristics associated with diurnal and nocturnal land surface temperature}
+\title{Night and day: What are the nonlinear associations between urban characteristics and land surface temperature?}
 
@@
 \begin{keyword}
-
+Urban climate $|$ LandSat $|$ Land surface temperature $|$ Urban canyon $|$ Machine learning
 \end{keyword}
@@
 \section{Introduction}
-In a warming world, understanding the drivers of LST and the UHI will aid in adapting cities to mitigate urban heat for the health and well being of communities.
-The 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine 2008} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera 2016 in Wicki} }, is expectAZed to become an annual occurrence by 2080 under current climate projections \cite{klinenberg2015heat}.
-Heat waves are exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
-The UHI results from heat captured during the day being released during the night increasing nocturnal temperatures \cite{Oke 1987, Landsberg1981, Rotach2005(bubble)}.
-Among other negative effects on a city's environmental quality, the UHI reduces the ability of people to cool off during the night and causes increased mortality \cite{Stone2006}.
-The UHI is studied in two ways.
-The first considers the air temperature using atmospheric station data \cite{Scott2016-lc, etc} while the other examines the land surface temperature (LST) using remote sensing \cite{imhoff, Peng2012-iy, Peng2018-cp, Zhou2014-wc, etc.}.
-Here, we study the LST because the spatial coverage enables comparative studies \cite{Hung2006-qy}.
+In a warming world, understanding the drivers of high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and well being of their communities.
+And mitigate they must, the 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}.
+Heat waves' affect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
+The UHI results from heat captured during the day being released during the night, thus increasing nocturnal temperatures \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
+This increase in nocturnal temperature reduces people's ability to cool off during the night and drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.
 
-%In a warming world, understanding the drivers of LST and the UHI will aid in adapting cities to mitigate urban heat for the health and well being of communities.
-Due to the availability of satellite images most studies focus on daytime LST \cite{examples looking at day LST}.
-However, the impact of urban characteristics on LST allegedly differ during the night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
-Given that UHI is a nocturnal phenomena, it is important that the drivers of nocturnal LST are understood to enhance mitigation strategies.
+Urban heat is studied in two ways.
+The first uses weather station data to analyze air temperature \cite{Scott2016-lc}(etc.) while the other examines the land surface temperature (LST) using remote sensing \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc}.
+Although there are differences between air temperature and surface temperature \cite{Good2016-yk}, the extensive spatial coverage is a major advantage and enables comparative studies \cite{Hung2006-qy}.
+For the sake of comparative analysis, we use LST in this study.
 
-The existing studies are yet to provide a clear understanding of the the influence and relative importance of drivers on nocturnal LST \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc }.
-There are three aspects in which these studies can be enhanced: factors considered, multiple study regions, and their statistical techniques.
-The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{}.
-LandSat8 scenes during the night have recently been made available. 
-Existing studies have also been criticized for the explanatory variables they've used.
+Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy},due to the lack of availability of nocturnal satellite images.
+However, the mechanisms and urban characteristics driving LST allegedly differs between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
+Given that UHI is a nocturnal phenomena and the lack of studies focused on nocturnal LST, we are yet to fully understand what drives LST during the night.
+
+The existing studies on nocturnal LST are yet to provide a clear understanding of the the influence and relative importance of urban characteristics associated with land surface temperature \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
+These studies can be enhanced in three ways: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) improving their statistical techniques.
+The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
+However, LandSat8 (L8) night scenes have recently become available. 
+Due to the wide spatial availability of LandSat imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
+
+Existing studies have also been criticized for the explanatory variables they've used. [who's criticised them?]
 There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
-Recent studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
+Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
 Beyond the 2D or 3D question, five categories of important variables have been identified: Green space, Water, Landscape, Albedo, and Socio-economic \cite{Peng2018-cp}. 
-However, many studies do not capture all aspects, or worse, analyze only the univariate effect of variables \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
+However, many studies do not capture all categories and, worse, many analyze only the univariate effect of variables \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
 Considering only single variables limits the understanding of variable effect given the interdependencies between drivers.
 
-The statistical approaches used in the studies limit their ability to explain the interdependencies between variables. 
-Almost all studies continue to use linear techniques \cite{Li2017-yl, etc.}. 
-This makes quantifying the independent effects of factors on LST and their relative importance difficult \cite{Peng2018-cp, Zhou2014-wc}.
-These studies also only assess their statistical accuracy using in-sample validation techniques \cite{something about validation e.g. Schmuli}.
-These result in higher reported values of explained variance (R$^2$ or other metric) because the trained models are not tested on unseen data.
-Furthermore, the variance explained is not high to begin with \cite{Chun2017-mm}.
+The third potential enhancement is in their statistical approaches.
+* multicollinearity problems and multiple statistical techniques might be need to enhance reliability and interpretation \cite{Zhou2014-wc}
+The current approaches used in the studies limit their ability to explain the interdependencies between variables. 
+Almost all studies continue to use linear techniques (e.g.
+\cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}, making quantifying the independent effects of factors on LST and their relative influence difficult \cite{Peng2018-cp, Zhou2014-wc}.
+These studies also only assess their accuracy using in-sample validation techniques.
+In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data; the result is potential overestimation of explained variance (e.g.
+R$^2$ or other validation metric).
 
-Finally, comparative studies between cities are necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}.
-
-Therefore, in this study we determine the influence and relative importance of the various drivers of LST, during both the night and day. 
-To achieve this conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the inter dependencies and relative importance between the factors.
+Therefore, in this study we determine the influence and relative importance of urban characteristics associated with LST, during both the night and day. 
+To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
+The objective is to assess whether the characteristics thought to influence high temperatures during the day are also influencing temperatures at night.
+This has implications for mitigating the severity of future heat waves.
 
@@
 
-Phoenix.
-* studied by Brazel 2007, Guhathakurta 2010.
+We study four cities in the contiguous United States (Figure \ref{fig:map}): Baltimore, MD; Detroit, MI; Phoenix, AZ; and Portland, OR.
+These four cities were selected as they include East and West coast cities, a mid-western city, and an arid central city.
+Phoenix, additionally, has been the subject of numerous other studies on land surface temperature.
+The cities were also selected due to lidar availability which is required for calculating the 3D variables.
+The constraint on selecting more cities was the time and computational requirements primarily for the LST and sky view factor calculations.
+
+Figure \ref{fig:map} shows the gridded LST data for each of the cities and Figure \ref{fig:joy} shows the distribution of each city's day and night land surface temperature.
 
@@
 Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station.
-This follows the process is described in \cite{Scott2016-lc} and is explicitly outlined in our open source code on Github\footnote{URL redacted for review}.
+This follows the process is described in \cite{Scott2016-lc} and is explicit in our open source code on Github\footnote{URL redacted for review}.
 
-To ameliorate the effect of shadows and other ephemeral changes we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
+To ameliorate the effect of shadows and other ephemeral changes \cite{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
 
-- include scatter of night vs day LST at 100m - also change the symbols for colorblind - remove background
-\begin{figure}[h]
+So we can conduct the comparative study between the cities, the land surface temperature for each city is calculated as the difference from the city's mean.
+
+\begin{figure}
 \begin{center}
@@
 \subsection{Independent variables}
-- present correlation matrix of independent variables
 
@@
 
+**** show scatter against tree canopy - cannot include both ****
+
 \textit{Land cover}.
@@
 
-\textit{Nighttime light intensity}.
+% \textit{Nighttime light intensity}.
 Stable nighttime light intensity is available from the Defense Meteorological Satellite Program (DMSP) Operational Linescan System (OLS).
@@
 \textit{Sky view factor}.
-Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
+Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
 We calculate SVF following using xxx and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$.
@@
 The data provided is at an xx resolution.
-The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the product does not contain information used in the LST calculation (which requires radiance).
+The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance).
 % * Rogan 2013, Elmes 2017
 
-\subsection{Statistical analysis}
-The data was gridded into 500m square and 100m square cells for analysis.
-For each of the variables, the mean, maximum, and minimum were calculated.
-Spatially lagging, the mean of the surrounding cells (including diagonal) for each variable, was included as an additional independent variable to capture the spatial effects which may be present within the data.
-* test / train
+\subsection{Data preparation and robustness}
 
-* statistical models
+To analyze the data, it needs to transformed into a spatially consistent set.
+The raw data is a variety of spatial data types from area level (the population density data), to geostatistical raster (e.g. the land surface temperature).
+We choose to resample the data into a grid of square cells.
+We conduct the resampling twice to produce a grid of 500m and 100 meter cells. 
+Having two grid resolutions allowed us to ensure our conclusions are robust.
+For each of the cells the the mean, maximum, and minimum of all variables were calculated. 
+To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable.
+The result is a data set which with we can train a variety of statistical models.
+
+Various statistical models (\S \ref{ss:models}) were trained and then, crucially, validated on unseen data using a technique known as holdout cross-validation. 
+This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
+The model resulting from the training, is then tested on this unseen test set, to get an estimate of the model accuracy.
+This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
+Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
+Overfitting occurs when a model fits to the randomness in a dataset, causing it to not be suitable for generalizations. 
+When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
+This is essential for statistical models and overlooked by the majority of existing studies (e.g. \cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
+Cross-validation helps to avoid this by evaluating the model on unseen data.
+
+A further way to avoid overfitting is to carefully consider how the test data is selected.
+This is especially important with spatial data.
+Therefore, when selecting the test and training data the grid cells were grouped into a larger 8x8 grid.
+This avoids over fitting of the model by training the model on a cell adjacent to a cell that is included in the test set.
+
+Finally, we attempt to capture uncertainty in the data and the models. 
+To represent the data uncertainty we conduct randomly sample from our data to create a new dataset.
+The models are trained on this new dataset and this is repeated.
+This approach, known as bootstrapping, means that the conclusions' sensitivity to the data can be assessed.
+If the results are vastly different for each bootstrap sample, we assume that the result is dependent on the data and is unlikely to be generalizable to other cities.
+Additionally, we capture model uncertainty by using different models and assessing how each model shows the urban characteristics influencing the land surface temperature.
+If all bootstrapped models are generally in agreement, it suggests that the conclusions are robust.
+
+\subsection{Statistical models}
+\label{ss:models}
+The relationship between urban characteristics and land surface temperature is complex.
+This complexity may not be captured by a linear regression model. 
+We fit a series of regression and data-mining models to the data and their predictive accuracy and variable association are compared.
+
+\textit{Null model: average}. The first model, to compare other models against, is the null model. 
+This is a benchmark model to ensure the models we fit are not doing worse than no model.
+The mean, average, of the observations is calculated and is used as the prediction for the test observations.
+
+\textit{Linear model}. Linear models are suitable when there is a linear relationship between the explanatory variables (urban characteristics) and the response variable (LST).
+
+\textit{Multivariate Adaptive Regression Spline (MARS)}. The MARS model extends the linear model, by using piece-wise functions to fit the data. \cite{Friedman1991-of}
+
+\textit{Generalized Additive Model (GAM)}.
+The GAM is also an extension of the linear model, although it does not require a linear relationship.
+Instead, the response variable is estimated as the sum of smoothing functions that are applied to each covariate \cite{Hastie1990-cg}.
+
+\textit{Random Forest}.
+A random forest is an ensemble model of \textit{Regression Trees}.
+A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}.
+It continues to divide the data on different covariates until the partitions are as similar as possible.
+The result is a tree-like structure, after which the model is named.
+In similar naming fashion, a random forest is a collection of regression trees.
+Each tree is \textit{grown} from a random subset of the data and the prediction from the random forest model is the average of the prediction from all of the trees \cite{Breiman2001-rt}.
+
+\textit{Gradient Boosted Regression Trees}. This is similar to the random forest, that it is a collective of regression trees. 
+However, the difference is that each tree is trained sequentially on the residuals of the previous tree.
+The result is the average of all of the regression trees \cite{Geron2017-ek}.
 
 \section{Results and Discussion}
+
+During the day, Increased impervious surfaces and decreased vegetation increase sensible heat flux and lower latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}
+
+Heat storage during the day, increases temperatures during the night \cite{Voogt2003-mm, Zhou2014-wc}
+
+Day and night mechanisms that drive temperature differ, likely because latent and sensible heat dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
+
+Vegetation and pervious surfaces caused the greatest reduction in land surface temperature, with approximately 10oC change during the day (Figure \ref{fig:pdp_2dday_500}b). 
+This result supports existing findings \cite{Zhou2014-wc, Chun2018-so}.
+it is likely due to vegetation increasing the latent heat fluxes due to transpiration \cite{Zhou2014-wc}, which explains why it has a greater effect during the day. 
+vegetation does not transpire during the night however, so the reduction in temperature due to increased tree cnaopy cover, is likely due to a reduction in impervious surfaces that store heat \cite{Voogt2003-mm}.
+Not all types of pervious surfaces are as good as others \cite{Gober2009-im}. Care must be taken when irrigating \cite{Gober2009-im}.
+urban design needs to consider interaction between surface materials \cite{Guhathakurta2010-ib}.
+- vegetation important during the day \cite{Chun2017-mm, Peng2018-cp}
+- vegetation is important \cite{Wang2019-tree}
+
+water
+- showed no effect in the 100m data, because so few data observations had water in them.
+- dramatic reductions in temperature during both night and day (y $^o$C).
+- this contrasts with findings and rationale that water releases heat at night \cite{Chun2017-mm} - although that logic applies to air temperature, not sure how strong water would be at increasing temperature of nearby land surface
+- supports strong negative findings \cite{Wang2019-water}
+
+Urbanization leads to heat storage in roads and buildings \cite{Zhou2014-wc, Voogt2003-mm}. 
+the results were surpringly inconsistent, regarding albedo - however, this is confounded as vegetation and highly impervious surfaces are both dark (low albedo).
+considering 2d partial dependence, we see that albedo has little affect at night it does show that the highest temperatures are the most built up and the least reflective (Figure \ref{fig:pdp_2dnight_500}a).
+during the day, tincreasing the whiteness, decreases the temperature.
+the higest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_500}a) and high impervious area and low albedo (Figure \ref{fig:pdp_2dday_500}b).
+- albedo, then ndbi to be most important, only consider day \cite{Peng2018-cp}
+- role of impervious surface \cite{Lu2006-ch}
+
+The canyon effect is also attributed to cause warmer temperatures. \cite{Chun2017-mm,Oke1988-re}
+while there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_night_500}d).
+This is potentially due to heat being stored within the canyons (areas with low, sky view factors).
+Conversely, during the day, areas with high sky view factor had increased temperature (Figure \ref{fig:pdp_day_500}d).
+This is may be a result of shade from the buildings \cite{Yu2019-ps}.
+Urban canyon was found to be important at night \cite{Chun2017-mm}
+
+The building area had no affect during the night (Figure \ref{fig:pdp_night_500}) and increased temperature during the day (Figure \ref{fig:pdp_day_500}).
+The population density had no discernible affect during night or day.
+
+Supports suggestion that 2D metrics outperform 3D metrics \cite{Berger2017-lx}
 
@@
     }
-    \label{fig:holdout_results}
+    \label{fig:holdout_500}
 \end{figure*}
-
-\subsection{Variables related to Land Surface Temperature}
 
@@
 In this case there is generally little difference between the importance of day and night variables.}
-\label{fig:importance}
+\label{fig:importance_500}
 \end{center}
@@
     }
-    \label{fig:pdp_night}
+    \label{fig:pdp_night_500}
 \end{figure*}
@@
     }
-    \label{fig:pdp_day}
+    \label{fig:pdp_day_500}
 \end{figure*}
@@
     }
-    \label{fig:pdp_2dnight}
+    \label{fig:pdp_2dnight_500}
 \end{figure*}
@@
     }
-    \label{fig:pdp_2dday}
+    \label{fig:pdp_2dday_500}
 \end{figure*}
 
-\section{Drivers of LST}
-See figure \ref{fig:importance} and \ref{fig:pdp}.
+\section{Conclusion}
 
-Regarding Fig \ref{fig:pdp}, I think I will remove the ``all" line, because in cases such as $alb_mean_min_sl$ (which is the minimum albedo in the surrounding cells (sl = spatial lag)) the values between the cities don't overlap so I think it becomes a proxy for other differences between the city when regressed together.
+We analyzed ...
 
-% \section{Conclusion}
+Tree canopy, impervious surface, and greeness had the largest effect on temperature during the night (change of x oC).
+During the day xyz
+
+Our results highlight
+
+This work strongly supports initiatives for increasing green infrastructure in cities \cite{Larsen2015-da, Meerow2017-xv}. 
+conducting a thorough nonlinear statistical analysis that accounts for seasonality is the next step to ensuring that interventions are effective year-round.
+
+It does not support claims that constrain floor area ratios, which would lead to reduced density \cite{Chun2017-mm}. 
+A further necessary study is to compare sprawling and dense cities, before steps to dissuade density are taken. 
+And steps such as vertical and horizontal randomness of buildings or simply green roofs may be what is required \cite{Gago2013-ta}.
+highlights the importance of urban planning strategies and studies that can quantify these would be valuable.
+
+To mitigate the effects of future heat waves, cities need to reduce their impervious surfaces and increase the amount of vegetation.
 
@@
 % Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
+\section*{Acknowledgements}
+TML - Hazard SEES, Rackham PreDoctoral
+What else?
 
@@
 
-\begin{figure}[h]
+\begin{figure}
 \begin{center}

### 2019-04-22 report/article.tex
Updates from Overleaf

@@
 \begin{keyword}
-Urban climate $|$ LandSat $|$ Land surface temperature $|$ Urban canyon $|$ Machine learning
+Urban climate; LandSat; Land surface temperature; Urban canyon; Machine learning
 \end{keyword}
@@
 Urban heat is studied in two ways.
-The first uses weather station data to analyze air temperature \cite{Scott2016-lc}(etc.) while the other examines the land surface temperature (LST) using remote sensing \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc}.
-Although there are differences between air temperature and surface temperature \cite{Good2016-yk}, the extensive spatial coverage is a major advantage and enables comparative studies \cite{Hung2006-qy}.
+The first uses weather station data to analyze air temperature \cite{Scott2016-lc}(etc.) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc}.
+Although there are differences between land temperature and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
 For the sake of comparative analysis, we use LST in this study.
 
-Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy},due to the lack of availability of nocturnal satellite images.
+Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy}.
 However, the mechanisms and urban characteristics driving LST allegedly differs between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
@@
     \begin{center}
-\includegraphics[width=\linewidth]{fig/report/lst_night-vs-day_500.png}
+    \includegraphics[width=\linewidth]{fig/report/scatter_500.png}
     \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution}
@@
 \end{figure}
+
 \subsection{Independent variables}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/holdout_results_500.pdf}
-    \caption{
-    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
+    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
+    \caption[Holdout cross-validation results at 500 meter resolution]{
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
+    The models were trained on 80\% of the data and tested on the unseen 20\%.
+    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
     OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
     Good models have MAE near 0.
-The gradient boosted regression forest generally outperforms the other models, except when predicting Detroit.
-Note that the R$^2$ axis is truncated at -1, although the multiple linear regression for diurnal land surface temperature tested on Phoenix has an OOB R$^2$ of -9.
-This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.
     }
@@
 \begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{fig/report/variableImportance_500.pdf}
-\caption{The variable importance for the regression during the night (to the left) and the day (to the right).
-Variable importance indicates the amount that each variable improves the performance of the regression.
-In this case there is generally little difference between the importance of day and night variables.}
-\label{fig:importance_500}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_night_500.png}
-    \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    \includegraphics[width=\linewidth]{fig/report/pdp_500.png}
+    \caption[Partial dependence plots for LST at 500 meter resolution]{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each urban characteristic as the other variables are held at their average (mean) value. 
+    The left hand side shows the effect each variable has on the (a) mean land surface temperature (LST) during the night, (b) maximum LST during the night, (c) mean LST during the day, (d) maximum LST during the day.
+    This is for the 500 meter resolution.
     }
@@
 \begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_day_500.png}
-    \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
-    }
-    \label{fig:pdp_day_500}
+    \begin{center}
+    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
+    \caption[Variable influence on LST at 500 meter resolution]{
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
+    \label{fig:importance_500}
+    \end{center}
 \end{figure*}
@@
 \section*{Acknowledgements}
-TML - Hazard SEES, Rackham PreDoctoral
-What else?
+Hazard SEES\\
+TML - Rackham PreDoctoral\\
+Anything else?
 
@@
 
-\section{Additional model results}
-\begin{figure}
+\newpage
+\section{City specific results}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_cities_500.png}
+    \caption[City specific partial dependence plots]{
+    The partial dependence plots for a random forest model trained on each city.
+    This is to evaluate whether the influence of urban characteristics on LST is consistent between the cities.
+    }
+    \label{fig:cities_500}
+\end{figure}
+
+\newpage
+\section{100 meter resolution results}
+\label{sec:100_meter}
+The figures presented in the main text, are replicated here based on data at a 100 meter resolution.
+
+\begin{figure}[h]
     \centering
@@
 
-\begin{figure}
+\begin{figure}[h]
 \begin{center}

### 2019-04-22 code/analysis/analysis.py
add city specific pdps

@@ -381,6 +381,84 @@ def boot_pd(df, grid_size, boot_index):
     # calculate the pd
     calculate_partial_dependence(df, grid_size, boot_index)
 
+
+def calculate_partial_dependence_city(df_full, grid_size, cities):
+    '''
+    fit the models to the entire dataset
+    loop through each feature
+    vary the feature over its range
+    predict the target variable to see how it is influenced by the feature
+    '''
+    results_partial = pd.DataFrame()
+    for city in cities:
+        print(city)
+        df = df_full[df_full.city == city]
+        df, target = prepare_lst_prediction(df)
+        df = subset_regression_data(df, df)[0]
+        df_reference = df.copy()
+        feature_resolution = 25
+        # loop day and night
+        for h in ['lst_day_mean', 'lst_night_mean', 'lst_night_max','lst_day_max']:
+            print(h)
+            ###
+            # fit models
+            ###
+            # gradient boosted tree
+            # gbm = GradientBoostingRegressor(max_depth=2, random_state=RANDOM_SEED, learning_rate=0.1, n_estimators=500, loss='ls')
+            # gbm.fit(df, target[h])
+            # random forest
+            rf = RandomForestRegressor(random_state=RANDOM_SEED, n_estimators=500, max_features=1/3)
+            rf.fit(df, target[h])
+            # mars
+            # mars = Earth(max_degree=1, penalty=1.0, endspan=5)
+            # mars.fit(df, target[h])
+            # GAM
+            # gam = LinearGAM(n_splines=10).fit(df, target[h])
+            # linear
+            # mlr = LinearRegression()
+            # mlr = mlr.fit(df, target[h])
+            ###
+            # loop through features and their ranges
+            ###
+            var_interests = {'lst_day_mean':['lcov_11', 'ndvi_mean', 'tree_mean', 'alb_mean','dsm_mean'],
+                            'lst_night_mean':['tree_mean', 'ndvi_mean', 'lcov_11', 'svf_mean', 'dsm_mean'],
+                            'lst_day_max':['tree_min', 'ndvi_mean','alb_mean','lcov_11','nbdi_max'],
+                            'lst_night_max':['tree_min', 'ndvi_mean','tree_mean','dsm_mean','svf_mean']}
+            for var_interest in var_interests[h]: #['tree_mean','density_housesarea']:
+                # loop through range of var_interest
[…]

### 2019-04-22 code/analysis/analysis.py
updating 2d pdp

@@ -1232,7 +1232,7 @@ def plot_2d_partialdependence(regressor, time_of_day, grid_size, df_x):
     Plot the 2d partial dependence
     '''
     # two way partial dependence
-    figs, axes = plt.subplots(3, 2, figsize = (width_2col, height_2c), sharey=False, sharex=False)
+    figs, axes = plt.subplots(3, 2, figsize = (width_2col, height_2c/3), sharey=False, sharex=False)
     # loop through the top n variables by nocturnal importance
     df_vars = list(df_x)
     if grid_size == 500:

### 2019-04-22 code/analysis/analysis.py
updating 2d pdp

@@ -1232,7 +1232,7 @@ def plot_2d_partialdependence(regressor, time_of_day, grid_size, df_x):
     Plot the 2d partial dependence
     '''
     # two way partial dependence
-    figs, axes = plt.subplots(3, 2, figsize = (width_2col, height_2c/3), sharey=False, sharex=False)
+    figs, axes = plt.subplots(3, 2, figsize = (width_2col, height_2c*2), sharey=False, sharex=False)
     # loop through the top n variables by nocturnal importance
     df_vars = list(df_x)
     if grid_size == 500:

### 2019-04-23 code/analysis/analysis.py
running results for 100 meter

@@ -1286,8 +1286,8 @@ def scatter_tree_imp(df, cities, grid_size):
         plt.scatter(df_city['imp_mean'], df_city['tree_mean'], label = city, alpha = 0.5)
         print(df_city['imp_mean'].corr(df_city['tree_mean']))
     plt.legend(loc='lower right')
-    plt.xlabel('\% impervious surface')
-    plt.ylabel('\% tree canopy cover')
+    plt.xlabel('% impervious surface')
+    plt.ylabel('% tree canopy cover')
     plt.text(20, 80,'correlation = {0:.2f}'.format(df['imp_mean'].corr(df['tree_mean'])), ha='left', va='top')
     plt.savefig('fig/report/imp_v_tree_{}.png'.format(grid_size), format='png', dpi=300, transparent=True)
     plt.show()

### 2019-04-23 code/analysis/prepare_data.py

@@ -12,7 +12,7 @@ scale = True
 def main(scale=True):
 
     cities = ['bal','det','phx','por']
-    grid_size = 500
+    grid_size = 100
     # init all city dataframe
     df = pd.DataFrame()
     add_index = 0

### 2019-04-30 code/analysis/analysis.py
importing the cnn results and updating 100 figures

@@ -44,7 +44,8 @@ city_names = {'bal':'baltimore','por':'portland','phx':'phoenix','det':'detroit'
 model_names = {'rf':'random forest','mlr':'multivariate linear',
                 'gam':'generalized additive\n(gam)',
                 'gbrt':'gradient boosted\ntrees',
-                'mars':'multivariate adaptive\nspline (mars)'}
+                'mars':'multivariate adaptive\nspline (mars)',
+                'cnn':'convolutional\nneural network'}
 feature_names = {'lcov_11' : '% water','tree_mean':'% tree canopy','ndvi_mean':'ndvi',
                 'svf_mean':'sky view factor','dsm_mean':'digital surface model',
                 'alb_mean':'albedo','dsm_sd':'dsm stand. dev.','nbdi_max':'max nbdi',
@@ -420,10 +421,16 @@ def calculate_partial_dependence_city(df_full, grid_size, cities):
             ###
             # loop through features and their ranges
             ###
-            var_interests = {'lst_day_mean':['lcov_11', 'ndvi_mean', 'tree_mean', 'alb_mean','dsm_mean'],
-                            'lst_night_mean':['tree_mean', 'ndvi_mean', 'lcov_11', 'svf_mean', 'dsm_mean'],
-                            'lst_day_max':['tree_min', 'ndvi_mean','alb_mean','lcov_11','nbdi_max'],
-                            'lst_night_max':['tree_min', 'ndvi_mean','tree_mean','dsm_mean','svf_mean']}
+            if grid_size == 500:
+                var_interests = {'lst_day_mean':['lcov_11', 'ndvi_mean', 'tree_mean', 'alb_mean','dsm_mean'],
+                                'lst_night_mean':['tree_mean', 'ndvi_mean', 'lcov_11', 'svf_mean', 'dsm_mean'],
+                                'lst_day_max':['tree_min', 'ndvi_mean','alb_mean','lcov_11','nbdi_max'],
+                                'lst_night_max':['tree_min', 'ndvi_mean','tree_mean','dsm_mean','svf_mean']}
+            else:
+                var_interests = {'lst_day_mean':['tree_mean', 'ndvi_mean', 'alb_mean','dsm_mean','svf_mean'],
+                                'lst_night_mean':['tree_mean', 'ndvi_mean', 'alb_mean', 'svf_mean', 'dsm_mean'],
[…]

### 2019-04-30 report/article.tex
Updates from Overleaf

@@
 \author[1]{S. Guikema}
+\author[]{A. Nisbet}
 
@@
 However, the independent effects and relative importance of potential drivers remain unclear from existing studies.
-We seek to answer the question: what are the effects and importance of drivers of land surface temperature during both day and night?
+We seek to answer the question: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?'' 
 To do this, we analyze the urban land surface temperature in four cities across the United States.
-In our analysis, we include factors related to vegetation, water, the built-environment, and topography and control them using nonlinear statistical methods which allow for their independent effects to be assessed.
-The effects from the diurnal and nocturnal analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
-Understanding the relationships influencing both nocturnal and diurnal land surface temperature will improve climate adaptation planning and heat wave mitigation.
+In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
+We control these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
+The effects from the daytime and nighttime analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
+Understanding the relationships influencing both night and day land surface temperature will improve climate adaptation planning and heat wave mitigation.
 \end{abstract}
@@
 \begin{keyword}
-Urban climate; LandSat; Land surface temperature; Urban canyon; Machine learning
+Urban climate; LandSat; Land surface temperature; Urban canyon; Machine learning; Convolutional Neural Network
 \end{keyword}
@@
 In a warming world, understanding the drivers of high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and well being of their communities.
-And mitigate they must, the 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}.
+And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}.
 Heat waves' affect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
@@
 Urban heat is studied in two ways.
-The first uses weather station data to analyze air temperature \cite{Scott2016-lc}(etc.) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc}.
-Although there are differences between land temperature and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
-For the sake of comparative analysis, we use LST in this study.
+The first uses weather station data to analyze air temperature (e.g.
+\cite{Scott2016-lc}) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
+Although there are differences between LST and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
+We analyze LST in this study.
 
@@
 However, the mechanisms and urban characteristics driving LST allegedly differs between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
-Given that UHI is a nocturnal phenomena and the lack of studies focused on nocturnal LST, we are yet to fully understand what drives LST during the night.
+Given that UHI is a nocturnal phenomena, the lack of study on nighttime LST leaves a critical gap in our understanding. 
+This can now be addressed as high resolution nighttime satellite images have become available.
 
-The existing studies on nocturnal LST are yet to provide a clear understanding of the the influence and relative importance of urban characteristics associated with land surface temperature \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
-These studies can be enhanced in three ways: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) improving their statistical techniques.
+Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
+Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) improving their statistical techniques.
 The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
-However, LandSat8 (L8) night scenes have recently become available. 
-Due to the wide spatial availability of LandSat imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
+1km resolution makes it difficult to attribute LST to urban characteristics.
+However, 30-m resolution LandSat8 (L8) night scenes have recently become available. 
+Due to the wide spatial availability of L8 imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
 
-Existing studies have also been criticized for the explanatory variables they've used. [who's criticised them?]
+Existing studies have also been criticized for the explanatory variables they've used \cite{Chun2017-mm,Peng2018-cp}.
 There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
 Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
-Beyond the 2D or 3D question, five categories of important variables have been identified: Green space, Water, Landscape, Albedo, and Socio-economic \cite{Peng2018-cp}. 
-However, many studies do not capture all categories and, worse, many analyze only the univariate effect of variables \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
-Considering only single variables limits the understanding of variable effect given the interdependencies between drivers.
+Beyond the 2D or 3D debate, five categories of important variables have been identified: Green space, water, landscape, albedo, and socio-economic \cite{Peng2018-cp}. 
+However, many studies do not capture these categories and, worse, many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
+Considering a variable in isolation, without accounting for potential conflating by other variables, limits the understanding of the interdependent effects that exist.
 
-The third potential enhancement is in their statistical approaches.
-* multicollinearity problems and multiple statistical techniques might be need to enhance reliability and interpretation \cite{Zhou2014-wc}
-The current approaches used in the studies limit their ability to explain the interdependencies between variables. 
-Almost all studies continue to use linear techniques (e.g.
-\cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}, making quantifying the independent effects of factors on LST and their relative influence difficult \cite{Peng2018-cp, Zhou2014-wc}.
-These studies also only assess their accuracy using in-sample validation techniques.
-In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data; the result is potential overestimation of explained variance (e.g.
-R$^2$ or other validation metric).
+The third potential enhancement is in the statistical approaches.
+Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
+Using linear models incur a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
+The first is that many of the urban characteristics exhibit high multicollinearity \cite{Zhou2014-wc}.
+The second is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \cite{Peng2018-cp, Zhou2014-wc}.
+Additionally, and most importantly, these studies only assess their accuracy using in-sample validation techniques.
+In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data. 
+This risks overestimating the accuracy of their models.
 
-Therefore, in this study we determine the influence and relative importance of urban characteristics associated with LST, during both the night and day. 
+The question we address is: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?''
 To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
-The objective is to assess whether the characteristics thought to influence high temperatures during the day are also influencing temperatures at night.
-This has implications for mitigating the severity of future heat waves.
+In addressing our question, we will address the, at times fundamental, limitations of existing studies.
+Most importantly we look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent effects of urban characteristics on urban heat.
+The results have implications for mitigating the severity of future heat waves.
 
@@
 \subsection{Land surface temperature}
-We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery as well as land cover and an air temperature observation (data sources provided in \ref{tab:data}.
-Band-10 digital-number data is converted to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}.
+We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}).
+Our code is available on our Github\footnote{URL redacted for review} repository, and is as follows:
+We convert Band-10 digital-number data to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. 
 We correct for emissivity using land cover data \cite{Alipour2003-gb}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
 Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
-This follows the process is described in \cite{Scott2016-lc} and is explicit in our open source code on Github\footnote{URL redacted for review}.
+This follows the process is described in \cite{Scott2016-lc}.
 
@@
     \includegraphics[width=\linewidth]{fig/report/scatter_500.png}
-    \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution}
+    \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution.}
     \label{fig:scatter_lst}
@@
 Albedo is a measure of the reflectivity or brightness of a surface.
-It is normalized between 0-1 where the darker surfaces are lower values.
+It is normalized between 0-100 where the darker surfaces are lower values.
 Albedo is calculated using the LandSat8 images using the algorithm described in \cite{Smith2010-nw, Liang2001-jd}.
@@
 
-\textit{Building height}.
+% \textit{Building height}.
 Building height is estimated using the building footprints and available lidar data following \cite{Chun2017-mm}.
@@
 
-\textit{Elevation}.
-1/3 arc-second (~10m) bare-earth elevation (topography) data is available courtesy the U.S.
-Geological Survey.
-The mean, max, and min of elevation points within the grid cell is calculated.
+\textit{Elevation}. 1/3 arc-second (~10m) bare-earth elevation (topography) data is available courtesy the U.S. Geological Survey.
 
@@
 The Landsat images were converted to top-of-atmosphere reflectance.
-The data is provided at a 30m resolution and we calculate the mean, max, and min for each grid cell.
-
-**** show scatter against tree canopy - cannot include both ****
+The data is provided at a 30m resolution.
 
@@
 The only landcover that we use is category 11, water.
-Cells with more than 20\% area classified as water are excluded from the data set.
 
@@
 \textit{Population density}. 
-Density allegedly increases the LST \cite{Li2017-yl}, although this result may be the result of confounding with other factors.
-Population density calculated from the U.S.
-census at the block level.
-The most recent census was 2010, so that data is used as an estimator of where people reside in the evening.
+Density allegedly increases the LST \cite{Li2017-yl, Peng2018-cp}, although this result may be the result of confounding with other factors. 
+Population density is calculated from the U.S. census at the block level. 
+The most recent census was 2010, so we use that data as an estimator of where people reside in the evening. 
 This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that it's centroid is contained within.
@@
 Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
-We calculate SVF following using xxx and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$.
-We use a spatial resolution of x-m.
+We calculate SVF using \cite{Van_doninck2018-ib} and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$.
+We use a spatial resolution of 6 meters.
 % * Yuan 2011, Unger 2004.
@@
 The percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}.
-The data provided is at an xx resolution.
+The data provided is at a 30m resolution. 
 The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance).
@@
 \subsection{Data preparation and robustness}
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
+    \caption[Holdout cross-validation results at 500 meter resolution]{
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
+    The models were trained on 80\% of the data and tested on the unseen 20\%.
+    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
+    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
+    Good models have MAE near 0.
+    }
+    \label{fig:holdout_500}
+\end{figure*}
+
+The objective of our study is to determine the most influential urban characteristics and understand how they relate to land surface temperature.
+To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
+Given a dataset and for a chosen characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature.
+This is repeated over a range for the feature of interest.
+To evaluate how multiple urban characteristics interact, partial dependence can also be conducted in multiple dimensions.
+The swing \cite{Shortridge2015-ub} measures the relative importance of variables by calculating the magnitude of change in temperature associated with each variable.
+To calculate partial dependence, we need a data set and a statistical model.
 
@@
 To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable.
+To address multicollinearity between the variables, that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than approximately 5.
 The result is a data set which with we can train a variety of statistical models.
@@
 This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
-The model resulting from the training, is then tested on this unseen test set, to get an estimate of the model accuracy.
+The model resulting from the training, is then tested on this unseen test set (known as out-of-bag), to get an estimate of the model accuracy.
 This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
@@
 
-\textit{Random Forest}.
-A random forest is an ensemble model of \textit{Regression Trees}.
+\textit{Random Forest}. A random forest is an ensemble model of \textit{Regression Trees}. 
 A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}.
@@
 Each tree is \textit{grown} from a random subset of the data and the prediction from the random forest model is the average of the prediction from all of the trees \cite{Breiman2001-rt}.
+Tree-based models do not assume linearity and so are generally very flexible and powerful models \cite{Breiman2001-rt, Geron2017-ek}.
 
@@
 \section{Results and Discussion}
-
-During the day, Increased impervious surfaces and decreased vegetation increase sensible heat flux and lower latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}
-
-Heat storage during the day, increases temperatures during the night \cite{Voogt2003-mm, Zhou2014-wc}
-
-Day and night mechanisms that drive temperature differ, likely because latent and sensible heat dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
-
-Vegetation and pervious surfaces caused the greatest reduction in land surface temperature, with approximately 10oC change during the day (Figure \ref{fig:pdp_2dday_500}b). 
-This result supports existing findings \cite{Zhou2014-wc, Chun2018-so}.
-it is likely due to vegetation increasing the latent heat fluxes due to transpiration \cite{Zhou2014-wc}, which explains why it has a greater effect during the day. 
-vegetation does not transpire during the night however, so the reduction in temperature due to increased tree cnaopy cover, is likely due to a reduction in impervious surfaces that store heat \cite{Voogt2003-mm}.
-Not all types of pervious surfaces are as good as others \cite{Gober2009-im}. Care must be taken when irrigating \cite{Gober2009-im}.
-urban design needs to consider interaction between surface materials \cite{Guhathakurta2010-ib}.
-- vegetation important during the day \cite{Chun2017-mm, Peng2018-cp}
-- vegetation is important \cite{Wang2019-tree}
-
-water
-- showed no effect in the 100m data, because so few data observations had water in them.
-- dramatic reductions in temperature during both night and day (y $^o$C).
-- this contrasts with findings and rationale that water releases heat at night \cite{Chun2017-mm} - although that logic applies to air temperature, not sure how strong water would be at increasing temperature of nearby land surface
-- supports strong negative findings \cite{Wang2019-water}
-
-Urbanization leads to heat storage in roads and buildings \cite{Zhou2014-wc, Voogt2003-mm}. 
-the results were surpringly inconsistent, regarding albedo - however, this is confounded as vegetation and highly impervious surfaces are both dark (low albedo).
-considering 2d partial dependence, we see that albedo has little affect at night it does show that the highest temperatures are the most built up and the least reflective (Figure \ref{fig:pdp_2dnight_500}a).
-during the day, tincreasing the whiteness, decreases the temperature.
-the higest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_500}a) and high impervious area and low albedo (Figure \ref{fig:pdp_2dday_500}b).
-- albedo, then ndbi to be most important, only consider day \cite{Peng2018-cp}
-- role of impervious surface \cite{Lu2006-ch}
-
-The canyon effect is also attributed to cause warmer temperatures. \cite{Chun2017-mm,Oke1988-re}
-while there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_night_500}d).
-This is potentially due to heat being stored within the canyons (areas with low, sky view factors).
-Conversely, during the day, areas with high sky view factor had increased temperature (Figure \ref{fig:pdp_day_500}d).
-This is may be a result of shade from the buildings \cite{Yu2019-ps}.
-Urban canyon was found to be important at night \cite{Chun2017-mm}
-
-The building area had no affect during the night (Figure \ref{fig:pdp_night_500}) and increased temperature during the day (Figure \ref{fig:pdp_day_500}).
-The population density had no discernible affect during night or day.
-
-Supports suggestion that 2D metrics outperform 3D metrics \cite{Berger2017-lx}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
-    \caption[Holdout cross-validation results at 500 meter resolution]{
-    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
-    The models were trained on 80\% of the data and tested on the unseen 20\%.
-    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
-    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
-    Good models have MAE near 0.
-    }
-    \label{fig:holdout_500}
-\end{figure*}
+In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
+To achieve this, the statistical models needed to be accurate.
+The results in Figure \ref{fig:holdout_500} indicate that both day and night land surface temperature can be predicted within 1$^o$C at night and 2$^o$C during the day using urban characteristics.
+The R$^2$ results, also based on unseen data, suggest that more than 80\% of the data variance is captured by the models for the mean LST during both night and day, while the mean \% variance explained is generally between 50 and 80\% for the maximum temperatures.
+Compared to models presented in similar studies, these are strong results, especially given that the existing studies failed to validate their models on unseen data.
+We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees.
+The weakest model is the linear regression, incidentally the one that is used in the majority of existing studies into land surface temperature.
+This strong result allows us to assess the influence and relative importance that these characteristics.
 
@@
     The left hand side shows the effect each variable has on the (a) mean land surface temperature (LST) during the night, (b) maximum LST during the night, (c) mean LST during the day, (d) maximum LST during the day.
+    Each of the models are shown and this indicates the model uncertainty in the relationships.
+    There are multiple lines for each model based on bootstrap samples of the data, which indicates the data uncertainty.
+    The histograms on the $x$-axis shown the distribution of the observed data.
     This is for the 500 meter resolution.
     }
-    \label{fig:pdp_night_500}
+    \label{fig:pdp_500}
 \end{figure*}
+
+Relative variable importance for all of the urban characteristics, calculated with \textit{swing}, is shown in Figure \ref{fig:importance_500}. 
+The characteristics, shown on the $y$ axis are ordered by the mean swing across all of the models, and the models are ordered left-to-right by their cross-validated accuracy.
+The most important characteristics during the night are tree canopy cover, greeness (nbdi), water, the sky view factor, and the elevation (digital surface model).
+This is consistent for both the mean and maximum temperature during the night.
+It is also generally consistent between the models. 
+In contrast, during the day, albedo (surface whiteness) is more important and the sky view factor is less influential.
+Albedo's small association with nighttime temperatures is surprising given that darker surfaces store heat during the day that is released during the night \cite{Voogt2003-mm, Zhou2014-wc}.
+This, and the other influences can be examined using partial dependence.
+
+\textit{Vegetation and impervious surfaces}. 
+During the day, increasing impervious surfaces and decreasing vegetation causes increased sensible heat flux and lowered latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}.
+This is thought to be less important during the night because latent and sensible heat dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
+Additionally, vegetation is expected to reduce daytime temperature due to transpiration which increases the latent heat flux \cite{Zhou2014-wc}. 
+However, because transpiration does not occur at night, vegetation's effect on night temperatures is debated. 
+In our study, percentage tree canopy and impervious surface cover must be discussed together because they are 100\% correlated (Figure \ref{fig:imp_tree}).
+To distinguish between vegetated surfaces and other pervious surfaces, we also include NDVI, a measure of a surface's greeness.
+At night, we find that \% tree canopy is the most influencial characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:pdp_500}).
+These partial dependence results indicate a decrease in up to 7$^oC$ as the percentage area of trees or other pervious surface increases.
+This reduction in temperature could be due to a reduction in the impervious surfaces that store heat.
+However, we also observe that as the NDVI increases, the land surface temperature decreases.
+The two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_500}) evaluates how LST varies with both characteristics.
+Figure \ref{fig:pdp_2dnight_500}d shows how the LST changes as both tree canopy/impervious surface and the NDVI change.
+The greener and more pervious a surface is, the cooler it is during the night. 
+The total change here is also approximately 7$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less variable of the five models shown in Figure \ref{fig:pdp_500}.
+These results contradict standing conclusions that vegetation has no effect on night time LST \cite{Peng2012-iy, Zhou2014-wc};
+rather, vegetation appears to be strongly associated with lower land surface temperatures during the night.
+
+During the day, we again see that vegetation and impervious surfaces are associated with the greatest reduction in land surface temperature, with approximately 10$^oC$ change (Figure \ref{fig:pdp_2dday_500}d).
+This daytime result is consistent with expectation \cite{Chun2017-mm, Peng2018-cp, Wang2019-tree,Zhou2014-wc, Chun2018-so}.
+Because not all pervious surface types have the same cooling potential, and because of trade-offs with irrigation of vegetated surfaces \cite{Gober2009-im}, understanding the ideal vegetated surface type for day and night temperatures is a future step.
+
+\textit{Water}.
+Water is widely expected to decrease the LST during the day \cite{Wicki2017-fv, Zhou2018-iy, Wang2019-water}, but some claim it increases LST during the night \cite{Chun2017-mm}.
+The rationale is that water releases heat during the night, resulting in elevated nighttime temperature \cite{Chun2017-mm}.
+However, while this is intuitive for air temperature, the affect on LST is not as apparent.
+Our findings show that water reduces the LST during both the day and night (Figure \ref{fig:pdp_500}).
+Although these results are not supported at the 100 meter resolution (\S \ref{ss:100_meter}), this is due to 98\% of the data at the 100 meter level having zero water.
+At the 500 meter resolution, we see that the presence of water can decrease LST between 1.5 and 8 $^oC$ during the night and by substantially more during the day (Figure \ref{fig:pdp_500}).
+
+\textit{Urbanization}.
+Urbanization can lead to heat storage in roads and buildings \cite{Zhou2014-wc, Voogt2003-mm}. 
+We discussed the role of impervious surfaces, alongside the effect of vegetation, but we also considered the influence of albedo (the whiteness of a surface), the percentage area of building, the NBDI (built-up index), and the sky view factor (a measure of the urban canyon effect).
+The results for albedo are surpringly inconsistent.
+However, this is potentially confounded as both vegetation and impervious surfaces can be dark surfaces (with low albedo).
+This stresses the importance of statistical techniques for evaluating the effect of multiple variables.
+Considering 2D partial dependence, we see that albedo has little affect at night, although the highest temperature does occur when there is high acreage of building footprint with low albedo (Figure \ref{fig:pdp_2dnight_500}a).
+During the day, albedo has a greater influence: increasing the whiteness, decreases the temperature (Figure \ref{fig:pdp_500}).
+The higest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_500}a), as well as high impervious area and low albedo (Figure \ref{fig:pdp_2dday_500}b).
+This supports existing findings that albedo decreases LST during the day, however, that we find no strong effect during the night contrasts existing reports \cite{Peng2012-iy, Zhou2014-wc}.
+
+The built-up index (NDBI), \% building area, and digital surface elevation model also had relatively minor associations with LST.
+The building area had no affect during the night (Figure \ref{fig:pdp_500}a) and increased temperature during the day (Figure \ref{fig:pdp_500}c).
+The largest affect observed is that maximum NDBI is associated with maximum daytime LST (Figure \ref{fig:pdp_500}).
+This is surprising given that NDBI has been reported as among the most important urban characteristics during the day \cite{Peng2018-cp}.
+However, the discrepancy may be due to \% impervious surface area being incorporated already.
+
+The canyon effect is also often attributed with causing warmer temperatures \cite{Chun2017-mm,Oke1988-re}.
+While there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_500}a,b).
+This is potentially due to heat being stored within the canyons (areas with low, sky view factors).
+This is supported in the two-dimensional partial dependence plot (Figure \ref{fig:pdp_2dnight_500}) where the higher temperatures are observed when there is high \% building area and low sky view factor.
+It follows that heat is being captured in the canyons, however, compared to the other urban characteristics this has a lesser affect, changing the temperature by approximately 1.5$^o$C.
+The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_500} and \ref{fig:pdp_2dday_500}).
+
+\textit{Population density}.
+Contrary to expectation (e.g. \cite{Peng2012-iy}), we found that population density had no discernible affect during night or day (Figure \ref{fig:importance_500}).
+
+\textit{Result sensitivity}.
+To assess the robustness of the results, we conduct the analysis again at the 100 meter resolution (\S \ref{ss:100_meter}).
+The results are consistent.
+Additionally, to ensure that the affects are consistent between cities, we construct the partial dependence plots for each city (\S \ref{ss:city}).
+The partial dependence is similar to Figure \ref{fig:pdp_500}.
+Therefore these results are robust.
 
@@
     \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    \textbf{Night time, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
     }
@@
     \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    \textbf{Day time, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
     }
@@
 \section{Conclusion}
+For the benefit of climate adaptation planning, we seek to determine how different urban characteristics are associated with land surface temperatures. 
+The results strongly support initiatives for increasing green infrastructure in cities \cite{Larsen2015-da, Meerow2017-xv}. 
+We found that increasing vegetation and reducing impervious surface area had the greatest effect on temperature during both night and day, with the potential to reduce the temperature up to 10$^o$C.
+We also see strong evidence for blue space and increasing the area of water, although there various limitations with doing so in cities such as Phoenix.
 
-We analyzed ...
+Our findings demonstrate that accurate prediction of land surface temperature using urban characteristics is possible.
+This result opens opportunities for further detailed analysis into potential interventions.
+Such interventions for mitigating high temperatures are naturally place-specific, and while these results have proven general to four different US cities, work is needed to understand questions such as what types of greenery are better than others \cite{Gober2009-im}.
+This study also demonstrates the importance of improved statistical analysis in these studies, given that the results observed constrasted with some existing conclusions.
+At the very least, studies need to appropriately validate their models using cross-validation in future.
+Another future step is conducting a thorough nonlinear statistical analysis that accounts for seasonality to ensuring that interventions are effective year-round.
 
-Tree canopy, impervious surface, and greeness had the largest effect on temperature during the night (change of x oC).
-During the day xyz
+Rigorous statistical analysis can continue to answer on-going questions central to land surface temperature.
+For example, our results support the suggestion that 3D metrics do not outperform 2D ones \cite{Berger2017-lx}, however, iteratively removing these variables would address the potential for conflating the importance of different characteristics \cite{Chun2017-mm}.
 
-Our results highlight
+Additionally, our results do not support claims advocating to constrain floor area ratios \cite{Chun2017-mm}.
+Such policy could lead to increased sprawl.
+If this type of action is explored, a further necessary study is to compare sprawling and dense cities, prior to taking steps to dissuade density. 
+Now that 30 meter resolution nighttime data is available, as well as the increasing availability of lidar data, studies advocating creative interventions such as increasing the vertical and horizontal randomness of buildings, increasing the prevalence of green roofs \cite{Gago2013-ta}, or other urban planning strategies can quantitatively be explored.
 
-This work strongly supports initiatives for increasing green infrastructure in cities \cite{Larsen2015-da, Meerow2017-xv}. 
-conducting a thorough nonlinear statistical analysis that accounts for seasonality is the next step to ensuring that interventions are effective year-round.
-
-It does not support claims that constrain floor area ratios, which would lead to reduced density \cite{Chun2017-mm}. 
-A further necessary study is to compare sprawling and dense cities, before steps to dissuade density are taken. 
-And steps such as vertical and horizontal randomness of buildings or simply green roofs may be what is required \cite{Gago2013-ta}.
-highlights the importance of urban planning strategies and studies that can quantify these would be valuable.
-
-To mitigate the effects of future heat waves, cities need to reduce their impervious surfaces and increase the amount of vegetation.
+This study into daytime and nighttime land surface temperature in four US cities has highlighted the importance of vegetation in our cities' mitigation options. 
+It has also demonstrated the utility of leveraging advanced statistical analysis to study land surface temperature.
+The results are robust to both data and model uncertainty, and are general across the cities studied.
+They suggest that vegetation and impervious surfaces are the most important urban characteristics associated with land surface temperature.
+Increasing and decreasing these, respectively, is necessary for reducing high urban temperatures during both night and day.
 
@@
 \section*{Acknowledgements}
-Hazard SEES\\
-TML - Rackham PreDoctoral\\
-Anything else?
+This research was conducted with the support of the University of Michigan's Rackham PreDoctoral Fellowship, and the National Science Foundation's Grant SEES-1631409.
+This support is gratefully acknowledged.
 
@@
 \section{City specific results}
+\label{ss:city}
 \begin{figure}[h]
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/pdp_cities_100.png}
+    \caption[City specific partial dependence plots]{
+    The partial dependence plots for a random forest model trained on each city.
+    This is to evaluate whether the influence of urban characteristics on LST is consistent between the cities.
+    }
+    \label{fig:cities_100}
+\end{figure}
+
 \newpage
 \section{100 meter resolution results}
-\label{sec:100_meter}
+\label{ss:100_meter}
 The figures presented in the main text, are replicated here based on data at a 100 meter resolution.
 
-\begin{figure}[h]
+Discuss the VIF results
+Discuss the results of the variable influence - why is water not recognized
+
+\begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/holdout_results_100.pdf}
-    \caption{
-    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models when fitted on three of the four cities and then used to predict the remaining city.
+    \includegraphics[width=\linewidth]{fig/report/holdout_100.png}
+    \caption[Holdout cross-validation results at 100 meter resolution]{
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
+    The models were trained on 80\% of the data and tested on the unseen 20\%.
+    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
     OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
     Good models have MAE near 0.
-The gradient boosted regression forest generally outperforms the other models, except when predicting Detroit.
-Note that the R$^2$ axis is truncated at -1, although the multiple linear regression for diurnal land surface temperature tested on Phoenix has an OOB R$^2$ of -9.
-This shows that the gradient boosted regression forest (gbm) consistently outperforms the multiple linear regression and null models.
+    This is for the 100 meter resolution.
     }
-    \label{fig:holdout_results}
-\end{figure}
+    \label{fig:holdout_100}
+\end{figure*}
 
-\begin{figure}[h]
+\begin{figure*}
     \begin{center}
-\includegraphics[width=\textwidth]{fig/report/variableImportance_100.pdf}
-\caption{The variable importance for the regression during the night (to the left) and the day (to the right).
-Variable importance indicates the amount that each variable improves the performance of the regression.
-In this case there is generally little difference between the importance of day and night variables.}
-\label{fig:importance}
+    \includegraphics[width=\linewidth]{fig/report/importance_100.png}
+    \caption[Variable influence on LST at 100 meter resolution]{
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature at 100 meter resolution.}
+    \label{fig:importance_100}
     \end{center}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_night_100.png}
-    \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
+    \includegraphics[width=\linewidth]{fig/report/pdp_100.png}
+    \caption[Partial dependence plots for LST at 100 meter resolution]{
+    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each urban characteristic as the other variables are held at their average (mean) value. 
+    The left hand side shows the effect each variable has on the (a) mean land surface temperature (LST) during the night, (b) maximum LST during the night, (c) mean LST during the day, (d) maximum LST during the day.
+    Each of the models are shown and this indicates the model uncertainty in the relationships.
+    There are multiple lines for each model based on bootstrap samples of the data, which indicates the data uncertainty.
+    The histograms on the $x$-axis shown the distribution of the observed data.
+    This is for the 100 meter resolution.
     }
-    \label{fig:pdp_night}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_uncert_day_100.png}
-    \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
-    }
-    \label{fig:pdp_day}
-\end{figure}
+    \label{fig:pdp_100}
+\end{figure*}
 
@@
 
+\newpage
+\section{Additional figures}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/imp_v_tree_500.png}
+    \caption{
+    Percentage tree canopy cover and impervious surface are 100\% correlated.
+    }
+    \label{fig:imp_tree}
+\end{figure}
+
 \end{document}

### 2019-04-30 report/article.tex

@@
 \author[]{A. Nisbet}
+% \footnote{Beginning September 2019, T Logan will be a faculty member at the University of Canterbury, NZ}
 
@@
 Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
-Understanding these drivers is necessary for reducing land surface temperature and mitigating heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
+Understanding these drivers is necessary for reducing land surface temperature and mitigating the effects of heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
 % However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
@@
 We seek to answer the question: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?'' 
-To do this, we analyze the urban land surface temperature in four cities across the United States.
+To do this, we analyze urban land surface temperature in four cities across the United States.
 In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
-We control these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
+We model the effects of these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
 The effects from the daytime and nighttime analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
-Understanding the relationships influencing both night and day land surface temperature will improve climate adaptation planning and heat wave mitigation.
+Our results suggest that vegetation and impervious surfaces are the most important urban characteristics associated with land surface temperature.
+Increasing and decreasing these, respectively, is necessary for reducing high urban temperatures during both night and day.
+Our results also demonstrate the potential for using nonlinear statistical analysis to investigate land surface temperature and its relationships with urban characteristics. 
+Improved understanding of these relationships influencing both night and day land surface temperature will assist planners undertaking climate change adaptation and heat wave mitigation.
 \end{abstract}
@@
 \begin{keyword}
-Urban climate; LandSat; Land surface temperature; Urban canyon; Machine learning; Convolutional Neural Network
+LandSat; Land surface temperature; Urban canyon; Machine learning; Convolutional Neural Network
 \end{keyword}
@@
 Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
-Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) improving their statistical techniques.
+Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relax their linear assumption, 4) rigorously testing and validing their models.
 The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
@@
 Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
-Beyond the 2D or 3D debate, five categories of important variables have been identified: Green space, water, landscape, albedo, and socio-economic \cite{Peng2018-cp}. 
+Beyond the 2D or 3D debate, important variables include green space, water bodies, albedo, and socio-economic factors \cite{Peng2018-cp}. 
 However, many studies do not capture these categories and, worse, many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
@@
 
-The third potential enhancement is in the statistical approaches.
+The third potential enhancement is in the statistical models.
 Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
@@
 The second is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \cite{Peng2018-cp, Zhou2014-wc}.
-Additionally, and most importantly, these studies only assess their accuracy using in-sample validation techniques.
+
+Finally, and perhaps most importantly, the existing studies fail to appropriately validate their models.
+The approach they use to assess their accuracy is in-sample validation.
 In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data. 
@@
 In addressing our question, we will address the, at times fundamental, limitations of existing studies.
-Most importantly we look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent effects of urban characteristics on urban heat.
+Most importantly we look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
 The results have implications for mitigating the severity of future heat waves.
@@
 
-To analyze the data, it needs to transformed into a spatially consistent set.
-The raw data is a variety of spatial data types from area level (the population density data), to geostatistical raster (e.g. the land surface temperature).
+To analyze the data, it needs to be transformed into a suitable format that is consistent between the variables.
+The raw data is a variety of spatial data types from ``area level" (the population density data), to ``geostatistical raster" (e.g. the land surface temperature).
 We choose to resample the data into a grid of square cells.
@@
 Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
-Overfitting occurs when a model fits to the randomness in a dataset, causing it to not be suitable for generalizations. 
+Over fitting occurs when a model fits to the randomness in a dataset, causing it to be unsuitable for generalizations. 
 When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
-This is essential for statistical models and overlooked by the majority of existing studies (e.g. \cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
+Avoiding over fitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies (e.g.
+\cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
 Cross-validation helps to avoid this by evaluating the model on unseen data.
 
-A further way to avoid overfitting is to carefully consider how the test data is selected.
+A further way to avoid over fitting is to carefully consider how the test data is selected.
 This is especially important with spatial data.
@@
 Finally, we attempt to capture uncertainty in the data and the models. 
-To represent the data uncertainty we conduct randomly sample from our data to create a new dataset.
+To represent the data uncertainty we draw a random sample from our data, with replacement, to simulate a new dataset.
 The models are trained on this new dataset and this is repeated.
@@
 This is a benchmark model to ensure the models we fit are not doing worse than no model.
-The mean, average, of the observations is calculated and is used as the prediction for the test observations.
+The mean of the observations in the training set is calculated and is used as the prediction for the test observations.
 
@@
 The GAM is also an extension of the linear model, although it does not require a linear relationship.
-Instead, the response variable is estimated as the sum of smoothing functions that are applied to each covariate \cite{Hastie1990-cg}.
+Instead, the response variable is estimated as the sum of smoothing functions that are applied to each covariate or set of covariates \cite{Hastie1990-cg}.
 
@@
 
+\textit{Convolutional Neural Network}. 
+<add explanation>
+
 \section{Results and Discussion}
 In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
-To achieve this, the statistical models needed to be accurate.
+To achieve this, the statistical models need to be accurate.
 The results in Figure \ref{fig:holdout_500} indicate that both day and night land surface temperature can be predicted within 1$^o$C at night and 2$^o$C during the day using urban characteristics.
@@
 This strong result allows us to assess the influence and relative importance that these characteristics.
+
+\begin{figure*}
+    \begin{center}
+    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
+    \caption[Variable influence on LST at 500 meter resolution]{
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
+    \label{fig:importance_500}
+    \end{center}
+\end{figure*}
 
@@
 The characteristics, shown on the $y$ axis are ordered by the mean swing across all of the models, and the models are ordered left-to-right by their cross-validated accuracy.
-The most important characteristics during the night are tree canopy cover, greeness (nbdi), water, the sky view factor, and the elevation (digital surface model).
+The most important characteristics during the night are tree canopy cover, greeness (NDVI), water, the sky view factor, and the elevation (digital surface model).
 This is consistent for both the mean and maximum temperature during the night.
@@
 \begin{figure*}
-    \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
-    \caption[Variable influence on LST at 500 meter resolution]{
-    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
-    \label{fig:importance_500}
-    \end{center}
-\end{figure*}
-
-\begin{figure*}
     \centering
@@
 \section{Conclusion}
-For the benefit of climate adaptation planning, we seek to determine how different urban characteristics are associated with land surface temperatures. 
+To assist planners tackling heat waves, climate change and high urban temperatures generally, we seek to determine how different urban characteristics are associated with land surface temperatures. 
 The results strongly support initiatives for increasing green infrastructure in cities \cite{Larsen2015-da, Meerow2017-xv}. 
@@
 \newpage
+\section{Statistical Models}
+
+\lbrack Here I'll provide in-depth explanation of the statistical models used \rbrack
+
+\newpage
 \section{City specific results}
@@
 
-\newpage
+\clearpage
 \section{100 meter resolution results}
@@
 The figures presented in the main text, are replicated here based on data at a 100 meter resolution.
+\\
+\lbrack Discuss the VIF results  \rbrack\\
+\lbrack Discuss the results of the variable influence - why is water not recognized  \rbrack
 
-Discuss the VIF results
-Discuss the results of the variable influence - why is water not recognized
-
-\begin{figure*}
+\begin{figure*}[h]
     \centering
@@
 
-\begin{figure*}
+\begin{figure*}[h]
     \begin{center}
@@
 
-\begin{figure*}
+\begin{figure*}[h]
     \centering
@@
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
-\newpage
+\clearpage
 \section{Additional figures}
-\begin{figure}
+\begin{figure*}[h]
     \centering
@@
     \label{fig:imp_tree}
-\end{figure}
+\end{figure*}
 

### 2019-05-21 report/article.tex
updating with katie's edits and andrew's appendix

@@
 % \documentclass[sort&compress]{elsarticle}
+% \documentclass[final,3p,times,onecolumn,sort&compress]{elsarticle}
 \documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
@@
 \usepackage{float}
+\usepackage{amsmath}
 
@@
 
-\title{Night and day: What are the nonlinear associations between urban characteristics and land surface temperature?}
+\title{Night and day: What is the influence and relative importance of urban characteristics on land surface temperature?}
 
@@
 \begin{abstract}
-Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
-Understanding these drivers is necessary for reducing land surface temperature and mitigating the effects of heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
-% However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
-However, the independent effects and relative importance of potential drivers remain unclear from existing studies.
-We seek to answer the question: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?'' 
-To do this, we analyze urban land surface temperature in four cities across the United States.
+Heat waves are among the deadliest natural hazards and are expected to increase in frequency and severity under climate change. 
+Their impacts in cities can be exacerbated by the urban heat island (UHI), but the mechanisms that underlie severity and timing of the UHI and its interactions with heat waves are poorly understood.
+Understanding these mechanisms is necessary to design strategies that reduce land surface temperature and mitigate the effect of heat waves.
+Making use of recently available high-resolution day and night thermal satellite imagery and employing advanced nonlinear statistical models, we seek to answer the question \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
+To answer this question, we analyze urban land surface temperature in four cities across the United States.
 In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
@@
 \section{Introduction}
-In a warming world, understanding the drivers of high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and well being of their communities.
+In a warming world, understanding the factors contributing to high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and wellbeing of their communities.
 And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}.
-Heat waves' affect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
-The UHI results from heat captured during the day being released during the night, thus increasing nocturnal temperatures \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
-This increase in nocturnal temperature reduces people's ability to cool off during the night and drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.
+Heat waves' effect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
+The UHI is understood to be a product of multiple factors including: enhanced absorption of solar radiation, geometric effects that limit radiative cooling and ventilation, anthropogenic heat from vehicles and buildings, air pollution that traps outgoing radiation, high ratios between sensible and latent heat flux due to low vegetation cover and high impervious land cover, and the radiative characteristics of building materials.
+% \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
+The result is increased diurnal and nocturnal temperatures that reduce people's ability to cool off, especially during the night, which drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.
 
-Urban heat is studied in two ways.
-The first uses weather station data to analyze air temperature (e.g.
-\cite{Scott2016-lc}) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
-Although there are differences between LST and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
+Studies of urban heat include analyses of air temperature, which is a conventional meteorological variable that is often associated with health outcomes(e.g.
+\cite{Scott2016-lc}), and studies that use satellite-derived land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
+LST correlates with air temperature at large scales, but there are differences at intra-urban scales \cite{Good2016-yk}.
+While LST is not a conventional meteorological variable, it is directly related to the urban heat budget and has the advantage of being available with extensive spatial coverage in gridded satellite products \cite{Hung2006-qy}.
 We analyze LST in this study.
@@
 
-Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
-Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relax their linear assumption, 4) rigorously testing and validing their models.
+Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
+Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relaxing their linear assumption, 4) rigorously testing and validating their models.
 The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
 1km resolution makes it difficult to attribute LST to urban characteristics.
-However, 30-m resolution LandSat8 (L8) night scenes have recently become available. 
-Due to the wide spatial availability of L8 imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
+However, 100m resolution LandSat8 (L8) night scenes have recently become available. 
+Due to the wide spatial availability of L8 imagery, comparative studies between cities that are necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
 
@@
 Beyond the 2D or 3D debate, important variables include green space, water bodies, albedo, and socio-economic factors \cite{Peng2018-cp}. 
-However, many studies do not capture these categories and, worse, many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
+However, many studies do not capture these categories and many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
 Considering a variable in isolation, without accounting for potential conflating by other variables, limits the understanding of the interdependent effects that exist.
@@
 Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
-Using linear models incur a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
+Using linear models incurs a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
 The first is that many of the urban characteristics exhibit high multicollinearity \cite{Zhou2014-wc}.
@@
 
-Finally, and perhaps most importantly, the existing studies fail to appropriately validate their models.
+Finally, most existing studies do not rigorously validate their models.
 The approach they use to assess their accuracy is in-sample validation.
@@
 
-The question we address is: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?''
+The question we address is: \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
 To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
 In addressing our question, we will address the, at times fundamental, limitations of existing studies.
-Most importantly we look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
+We look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
 The results have implications for mitigating the severity of future heat waves.
@@
 
-Figure \ref{fig:map} shows the gridded LST data for each of the cities and Figure \ref{fig:joy} shows the distribution of each city's day and night land surface temperature.
+Figure \ref{fig:map} shows the 500m gridded nighttime LST data for each of the cities and Figure \ref{fig:joy} shows the distribution of each city's 100m resolution day and night land surface temperature.
 
@@
     \includegraphics[width=\textwidth]{fig/report/map_nocturnal_lst.pdf}
-    \caption{The gridded nocturnal land surface temperature in $^o$C of the cities studied. This is change from the mean (average) for each city.}
+    \caption[The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied]{The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied.
+This is change from the mean (average) for each city.}
     \label{fig:map}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/joyplot_lst_500.pdf}
-    \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied.}
+    \includegraphics[width=\linewidth]{fig/report/joyplot_lst_100.pdf}
+    \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied. 100m resolution.}
     \label{fig:joy}
@@
 We convert Band-10 digital-number data to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. 
-We correct for emissivity using land cover data \cite{Alipour2003-gb}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
-Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
+We correct for emissivity using land cover data \cite{Alipour2003-ym}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
+Finally, an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
 This follows the process is described in \cite{Scott2016-lc}.
 
-To ameliorate the effect of shadows and other ephemeral changes \cite{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
+To ameliorate the effect of ephemeral changes \cite{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
 
@@
     \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/scatter_500.png}
-    \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution.}
+    \includegraphics[width=\linewidth]{fig/report/scatter_100.png}
+    \caption{The land surface temperature in $^o$C of the cities studied at a 100m square resolution.}
     \label{fig:scatter_lst}
@@
 
-\textit{Impervious surface percentage}.
-This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}.
-To generate the impervious surface area (ISA) LandSat data, NLCD land cover, and nighttime light imagery is used.
-The stable nighttime light intensity is only used to estimate the boundary of of urban areas.
+\textit{Impervious surface percentage}. This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}. 
+To generate the impervious surface area (ISA), the consortium uses LandSat data, NLCD land cover, and nighttime light imagery. 
+The stable nighttime light intensity is only used to estimate the boundary of urban areas. 
 The Landsat images were converted to top-of-atmosphere reflectance. 
@@
 
-\textit{NBDI}. 
+\textit{NDBI}. 
 Normalized difference built-up index indicates the intensity of imperviousness \cite{Bhatti2014-ae}. 
-It is calculated from satellite images as $$NDVI=\frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}}$$ where $B_{NIR}$ and $B_{SWIR}$ are the reflectances in the near-infrared and short-wave infrared bands respectively \cite{Alhawiti2016-wv}.
-Using LandSat8 imagery, this is $$NDVI=\frac{B6-B5}{B6+B5}$$ \cite{barsi2014}.
+It is calculated from satellite images as $$NDBI=\frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}}$$ where $B_{NIR}$ and $B_{SWIR}$ are the reflectances in the near-infrared and short-wave infrared bands respectively \cite{Alhawiti2016-wv}.
+Using LandSat8 imagery, this is $$NDBI=\frac{B6-B5}{B6+B5}$$ \cite{barsi2014}.
 % *Kaplan 2018, Peng 2018
@@
 \textit{Population density}. 
-Density allegedly increases the LST \cite{Li2017-yl, Peng2018-cp}, although this result may be the result of confounding with other factors. 
-Population density is calculated from the U.S. census at the block level. 
+Some studies have found that population density has a positive influence on LST \cite{Li2017-yl, Peng2018-cp}, although this may be the result of confounding with other factors. 
+We calculate population density using the U.S. census at the block level. 
 The most recent census was 2010, so we use that data as an estimator of where people reside in the evening. 
-This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that it's centroid is contained within.
+This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that its centroid is contained within.
 
 \textit{Sky view factor}.
-Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
-We calculate SVF using \cite{Van_doninck2018-ib} and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$.
+Urban canyons have been found to have an effect on UHI because they prevent air circulation and hinder nighttime radiative cooling \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
+We calculate SVF using the R packages \textit{horizon} \cite{Van_doninck2018-ib} that is based off the algorithm presented in \cite{Dozier1990-kn}.
+We use the same parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. 
 We use a spatial resolution of 6 meters.
@@
 \textit{Tree canopy cover}. 
-The percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}.
+The 2011 edition of percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}.
 The data provided is at a 30m resolution. 
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
-    \caption[Holdout cross-validation results at 500 meter resolution]{
+    \includegraphics[width=\linewidth]{fig/report/holdout_100.png}
+    \caption[Holdout cross-validation results at 100-meter resolution]{
+    Holdout cross-validation results at 100-meter resolution.
     The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
@@
     }
-    \label{fig:holdout_500}
+    \label{fig:holdout_100}
 \end{figure*}
@@
 To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
-Given a dataset and for a chosen characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature.
+For a selected characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature. 
 This is repeated over a range for the feature of interest.
@@
 We choose to resample the data into a grid of square cells.
-We conduct the resampling twice to produce a grid of 500m and 100 meter cells. 
-Having two grid resolutions allowed us to ensure our conclusions are robust.
-For each of the cells the the mean, maximum, and minimum of all variables were calculated. 
+We conduct the resampling twice to produce a grid of 100 and 500-meter square cells.
+
+Having two resolutions allowed us to evaluate whether our conclusions are robust.
+For each of the cells the mean, maximum, and minimum of all variables were calculated. 
 To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable.
-To address multicollinearity between the variables, that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than approximately 5.
-The result is a data set which with we can train a variety of statistical models.
+To address multicollinearity between the variables that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than 5.
+The result is a data set with which we can train statistical models.
 
-Various statistical models (\S \ref{ss:models}) were trained and then, crucially, validated on unseen data using a technique known as holdout cross-validation. 
+Various statistical models (\S \ref{ss:models}) were trained and then validated on unseen data using a technique known as holdout cross-validation. 
 This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
-The model resulting from the training, is then tested on this unseen test set (known as out-of-bag), to get an estimate of the model accuracy.
+The model resulting from the training is then tested on this unseen test set (known as out-of-bag) to get an estimate of the model accuracy.
 This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
@@
 When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
-Avoiding over fitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies (e.g.
-\cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
+Avoiding overfitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies.
 Cross-validation helps to avoid this by evaluating the model on unseen data.
 
-A further way to avoid over fitting is to carefully consider how the test data is selected.
+A further way to avoid overfitting is to carefully consider how the test data is selected.
 This is especially important with spatial data.
 Therefore, when selecting the test and training data the grid cells were grouped into a larger 8x8 grid.
-This avoids over fitting of the model by training the model on a cell adjacent to a cell that is included in the test set.
+This avoids overfitting of the model by training the model on a cell adjacent to a cell that is included in the test set.
 
 Finally, we attempt to capture uncertainty in the data and the models. 
-To represent the data uncertainty we draw a random sample from our data, with replacement, to simulate a new dataset.
+To represent the data uncertainty, we draw a random sample from our data, with replacement, to simulate a new dataset.
 The models are trained on this new dataset and this is repeated.
@@
 
-\textit{Linear model}. Linear models are suitable when there is a linear relationship between the explanatory variables (urban characteristics) and the response variable (LST).
+\textit{Linear model}. Linear models are suitable when the relationship between the explanatory variables (urban characteristics) and the response variable (LST) is linear. 
+That is, linear models take the form $$ y = \beta_0 + \sum_i \beta_i x_i + \varepsilon$$ where $y$ is the response (LST), $x_i$ is each for the variables (urban characteristics), $\beta_0$ is the intersection, $\beta_i$ is how the LST changes linearly with each urban characteristic, and $\varepsilon$ is the normally distributed error.
 
-\textit{Multivariate Adaptive Regression Spline (MARS)}. The MARS model extends the linear model, by using piece-wise functions to fit the data. \cite{Friedman1991-of}
+\textit{Multivariate Adaptive Regression Spline (MARS)}. 
+MARS models are a type of non-parametric regression approach that are useful for high dimensional problems \cite{Hastie2009-ky}. 
+They extend the linear model by using piecewise functions to fit the data \cite{Friedman1991-of}. 
+Multiple linear approximations are summed across the variable space to construct an estimate of the  LST\cite{Friedman1991-of}.
 
 \textit{Generalized Additive Model (GAM)}. 
-The GAM is also an extension of the linear model, although it does not require a linear relationship.
-Instead, the response variable is estimated as the sum of smoothing functions that are applied to each covariate or set of covariates \cite{Hastie1990-cg}.
+The GAM is also an extension of the linear model but relaxes the assumption that the relationship is linear \cite{Hastie1990-cg}.
+Instead, the response variable is estimated as the sum of smoothing functions, splines, that are applied to each covariate or set of covariates: $$ y = \sum_i s_i(x_i)$$ where $s_i(x_i)$ is the smoothing function applied to each covariate \cite{Shortridge2015-ub}.
 
@@
 A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}. 
-It continues to divide the data on different covariates until the partitions are as similar as possible.
-The result is a tree-like structure, after which the model is named.
-In similar naming fashion, a random forest is a collection of regression trees.
-Each tree is \textit{grown} from a random subset of the data and the prediction from the random forest model is the average of the prediction from all of the trees \cite{Breiman2001-rt}.
+The tree continues to `grow' by recursively partitioning the data with the objective of maximizing the node impurity, so that the partitions are as similar as possible. 
+The result is a tree-like structure that uses thresholds on the covariates to estimate the response.
+
+We use a total of 500 regression trees.
+Each regression tree is trained on a randomly selected subset of 1/3 of the covariates. 
+The resulting prediction from the random forest is the unweighted mean of the prediction from all of the trees \cite{Breiman2001-rt}.
+A subset of the covariates is used to reduce the correlation between the predictions of each tree.
 Tree-based models do not assume linearity and so are generally very flexible and powerful models \cite{Breiman2001-rt, Geron2017-ek}.
 
-\textit{Gradient Boosted Regression Trees}. This is similar to the random forest, that it is a collective of regression trees. 
-However, the difference is that each tree is trained sequentially on the residuals of the previous tree.
-The result is the average of all of the regression trees \cite{Geron2017-ek}.
+\textit{Gradient Boosted Regression Trees}. 
+GBRT's are also a collective of regression trees. 
+In contrast to random forest models, each tree is trained sequentially on the residuals of the previous tree.
+That is, each regression tree has the objective of reducing the error of the previous tree.
+The prediction from the GBRT is the unweighted mean of all of the regression trees \cite{Geron2017-ek}.
+In this case, we again use 500 trees.
 
 \textit{Convolutional Neural Network}. 
-<add explanation>
+CNN's are a new technique in deep learning and have most commonly been used to analyze visual images.
+Their use in image processing makes them potentially suitable for geographic studies.
+As with other forms of neural networks, CNN's contain layers of mathematical functions (neurons) that operate on the independent variables \cite{Geron2017-ek}.
+There can be multiple layers of these neurons that pass the results as an input to subsequent layers.
+In this manner, neural networks can capture non-linearities.
+The advantage of convolutional neural networks above other types of neural networks is their ability to capture the spatial dependencies. 
+Because CNN's have seldom been used for geospatial data, we provide a more detailed explanation in (\ref{ss:cnn}).
 
@@
 In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
-To achieve this, the statistical models need to be accurate.
-The results in Figure \ref{fig:holdout_500} indicate that both day and night land surface temperature can be predicted within 1$^o$C at night and 2$^o$C during the day using urban characteristics.
-The R$^2$ results, also based on unseen data, suggest that more than 80\% of the data variance is captured by the models for the mean LST during both night and day, while the mean \% variance explained is generally between 50 and 80\% for the maximum temperatures.
-Compared to models presented in similar studies, these are strong results, especially given that the existing studies failed to validate their models on unseen data.
-We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees.
-The weakest model is the linear regression, incidentally the one that is used in the majority of existing studies into land surface temperature.
-This strong result allows us to assess the influence and relative importance that these characteristics.
+To achieve this, we first select variables to remove based on collinearity. 
+The remaining variables for both 100 and 500-meter resolution analysis are shown in \ref{tab:vif}. 
+The next phase is to assess the accuracy of the statistical models.
+The results in Figure \ref{fig:holdout_100} show that the best models can predict both day and night land surface temperature to within 1$^o$C using urban characteristics.
+The R$^2$ results, also based on unseen data, suggest that more than 90\% of the data variance is captured by the best models.
+These are strong results.
+We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees and convolutional neural network.
+The weakest model, consistently with R$^2$ less than 80\%, is the linear regression - incidentally the model that is used in the majority of existing studies into land surface temperature.
+These results allow us to assess the influence and relative importance that these characteristics.
 
@@
     \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
-    \caption[Variable influence on LST at 500 meter resolution]{
+    \includegraphics[width=\linewidth]{fig/report/importance_100.png}
+    \caption[Variable influence on LST at 100-meter resolution]{
+    Variable influence on LST at 100-meter resolution.
     The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
-    \label{fig:importance_500}
+    \label{fig:importance_100}
     \end{center}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_500.png}
-    \caption[Partial dependence plots for LST at 500 meter resolution]{
+    \includegraphics[width=\linewidth]{fig/report/pdp_100.png}
+    \caption[Partial dependence plots for LST at 100-meter resolution]{
+    Partial dependence plots for LST at 100-meter resolution.
     Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each urban characteristic as the other variables are held at their average (mean) value. 
@@
     The histograms on the $x$-axis shown the distribution of the observed data.
-    This is for the 500 meter resolution.
+    This is for the 500-meter resolution.
     }
-    \label{fig:pdp_500}
+    \label{fig:pdp_100}
 \end{figure*}
 
-Relative variable importance for all of the urban characteristics, calculated with \textit{swing}, is shown in Figure \ref{fig:importance_500}. 
-The characteristics, shown on the $y$ axis are ordered by the mean swing across all of the models, and the models are ordered left-to-right by their cross-validated accuracy.
-The most important characteristics during the night are tree canopy cover, greeness (NDVI), water, the sky view factor, and the elevation (digital surface model).
-This is consistent for both the mean and maximum temperature during the night.
-It is also generally consistent between the models. 
-In contrast, during the day, albedo (surface whiteness) is more important and the sky view factor is less influential.
-Albedo's small association with nighttime temperatures is surprising given that darker surfaces store heat during the day that is released during the night \cite{Voogt2003-mm, Zhou2014-wc}.
+Relative variable importance for all the urban characteristics, calculated with \textit{swing}, is shown in Figure \ref{fig:importance_100}. 
+The characteristics, shown on the $y$ axis, are ordered by the mean swing across all of the models, and the models are ordered left-to-right by their cross-validated accuracy.
+The most important characteristics during the night are tree canopy cover, greenness (NDVI), the sky view factor, and the elevation (digital surface model).
+The albedo is found to be important when assessing the 100-meter data, while the \% area covered in water is important with the 500-meter data.
+This is likely due to 98\% of the cells at 100m resolution having 0\% water, so the effect on the dataset is negligible.
+The important urban characteristics are consistent for both the mean and maximum temperature during the night.
+The characteristics are also relatively consistent between night and day and between the models. 
+Although albedo is among the top factors, its relative importance is low compared to tree canopy and greenness, especially during the night. 
+This low association with nighttime temperatures is surprising given that darker surfaces store heat during the day that is released during the night \cite{Voogt2003-mm, Zhou2014-wc}.
 This, and the other influences can be examined using partial dependence.
@@
 During the day, increasing impervious surfaces and decreasing vegetation causes increased sensible heat flux and lowered latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}.
-This is thought to be less important during the night because latent and sensible heat dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
-Additionally, vegetation is expected to reduce daytime temperature due to transpiration which increases the latent heat flux \cite{Zhou2014-wc}. 
+This is thought to be less important during the night because latent and sensible heat are dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
 However, because transpiration does not occur at night, vegetation's effect on night temperatures is debated. 
 In our study, percentage tree canopy and impervious surface cover must be discussed together because they are 100\% correlated (Figure \ref{fig:imp_tree}).
-To distinguish between vegetated surfaces and other pervious surfaces, we also include NDVI, a measure of a surface's greeness.
-At night, we find that \% tree canopy is the most influencial characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:pdp_500}).
-These partial dependence results indicate a decrease in up to 7$^oC$ as the percentage area of trees or other pervious surface increases.
+To distinguish between vegetated surfaces and other pervious surfaces, we also include NDVI, a measure of a surface's greenness.
+At night, we find that \% tree canopy is the most influential characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:pdp_100}).
+These partial dependence results indicate a decrease in up to 10$^oC$ as the percentage area of trees or other pervious surface increases.
 This reduction in temperature could be due to a reduction in the impervious surfaces that store heat.
 However, we also observe that as the NDVI increases, the land surface temperature decreases.
-The two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_500}) evaluates how LST varies with both characteristics.
-Figure \ref{fig:pdp_2dnight_500}d shows how the LST changes as both tree canopy/impervious surface and the NDVI change.
+The two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_100}) evaluates how LST varies with both characteristics.
+Figure \ref{fig:pdp_2dnight_100}d shows how the LST changes as both tree canopy/impervious surface and the NDVI change.
 The greener and more pervious a surface is, the cooler it is during the night. 
-The total change here is also approximately 7$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less variable of the five models shown in Figure \ref{fig:pdp_500}.
-These results contradict standing conclusions that vegetation has no effect on night time LST \cite{Peng2012-iy, Zhou2014-wc};
+The total change here is also approximately 7.5$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less variable of the five models shown in Figure \ref{fig:pdp_100}.
+These results contradict standing conclusions that vegetation has no effect on nighttime LST \cite{Peng2012-iy, Zhou2014-wc};
 rather, vegetation appears to be strongly associated with lower land surface temperatures during the night.
 
-During the day, we again see that vegetation and impervious surfaces are associated with the greatest reduction in land surface temperature, with approximately 10$^oC$ change (Figure \ref{fig:pdp_2dday_500}d).
+During the day, we again see that vegetation and impervious surfaces are associated with the greatest reduction in land surface temperature, with approximately 15$^oC$ change (Figure \ref{fig:pdp_2dday_100}: ndvi vs.
+\% tree canopy).
 This daytime result is consistent with expectation \cite{Chun2017-mm, Peng2018-cp, Wang2019-tree,Zhou2014-wc, Chun2018-so}.
@@
 The rationale is that water releases heat during the night, resulting in elevated nighttime temperature \cite{Chun2017-mm}.
-However, while this is intuitive for air temperature, the affect on LST is not as apparent.
+However, while this is intuitive for air temperature, the effect on LST is not as apparent.
 Our findings show that water reduces the LST during both the day and night (Figure \ref{fig:pdp_500}).
-Although these results are not supported at the 100 meter resolution (\S \ref{ss:100_meter}), this is due to 98\% of the data at the 100 meter level having zero water.
-At the 500 meter resolution, we see that the presence of water can decrease LST between 1.5 and 8 $^oC$ during the night and by substantially more during the day (Figure \ref{fig:pdp_500}).
+Although these results are not supported at the 100-meter resolution because 98\% of the data at the 100-meter level has zero percentage water.
+At the 500-meter resolution, we see that the presence of water can decrease LST between 1.5 and 8$^oC$ during the night and by substantially more during the day (Figure \ref{fig:pdp_500}).
 
@@
 Urbanization can lead to heat storage in roads and buildings \cite{Zhou2014-wc, Voogt2003-mm}. 
-We discussed the role of impervious surfaces, alongside the effect of vegetation, but we also considered the influence of albedo (the whiteness of a surface), the percentage area of building, the NBDI (built-up index), and the sky view factor (a measure of the urban canyon effect).
-The results for albedo are surpringly inconsistent.
-However, this is potentially confounded as both vegetation and impervious surfaces can be dark surfaces (with low albedo).
+We discussed the role of impervious surfaces, alongside the effect of vegetation, but we also considered the influence of albedo (the whiteness of a surface), the percentage area of building, the NDBI (built-up index), and the sky view factor (a measure of the urban canyon effect).
+The results for albedo are surprisingly low.
+During both the night and day, albedo has a negative effect on temperature.
+During the day, this is most pronounced, and whiter surfaces appear to be more than 5$^o$C cooler than darker ones (Figure \ref{fig:pdp_100}).
+The partial dependence plot (Figure \ref{fig:pdp_100}) also indicates that albedo's influence on temperature is potentially confounded as both vegetation and impervious surfaces can be dark surfaces (with low albedo).
 This stresses the importance of statistical techniques for evaluating the effect of multiple variables.
-Considering 2D partial dependence, we see that albedo has little affect at night, although the highest temperature does occur when there is high acreage of building footprint with low albedo (Figure \ref{fig:pdp_2dnight_500}a).
-During the day, albedo has a greater influence: increasing the whiteness, decreases the temperature (Figure \ref{fig:pdp_500}).
-The higest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_500}a), as well as high impervious area and low albedo (Figure \ref{fig:pdp_2dday_500}b).
+We can investigate this further using the 2D partial dependence; the highest temperature occurs when there are dark impervious surfaces, but dark vegetated surfaces are cool (Figure \ref{fig:pdp_2dnight_100}).
+Additionally, we see that albedo has little affect at night, although the highest temperature does occur when there is high acreage of building footprint with low albedo (Figure \ref{fig:pdp_2dnight_100}a).
+During the day, albedo has a greater influence: increasing the whiteness decreases the temperature (Figure \ref{fig:pdp_100}).
+The highest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_100}a), as well as high impervious area and low albedo (Figure \ref{fig:pdp_2dday_100}b).
 This supports existing findings that albedo decreases LST during the day, however, that we find no strong effect during the night contrasts existing reports \cite{Peng2012-iy, Zhou2014-wc}.
@@
 The built-up index (NDBI), \% building area, and digital surface elevation model also had relatively minor associations with LST.
-The building area had no affect during the night (Figure \ref{fig:pdp_500}a) and increased temperature during the day (Figure \ref{fig:pdp_500}c).
-The largest affect observed is that maximum NDBI is associated with maximum daytime LST (Figure \ref{fig:pdp_500}).
+The building area had no effect during the night (Figure \ref{fig:pdp_100}a) and increased temperature during the day (Figure \ref{fig:pdp_100}c).
+The largest effect observed is that maximum NDBI is associated with maximum daytime LST (Figure \ref{fig:pdp_100}).
 This is surprising given that NDBI has been reported as among the most important urban characteristics during the day \cite{Peng2018-cp}.
@@
 The canyon effect is also often attributed with causing warmer temperatures \cite{Chun2017-mm,Oke1988-re}.
-While there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_500}a,b).
+While there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_100}a,b).
 This is potentially due to heat being stored within the canyons (areas with low, sky view factors).
-This is supported in the two-dimensional partial dependence plot (Figure \ref{fig:pdp_2dnight_500}) where the higher temperatures are observed when there is high \% building area and low sky view factor.
-It follows that heat is being captured in the canyons, however, compared to the other urban characteristics this has a lesser affect, changing the temperature by approximately 1.5$^o$C.
-The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_500} and \ref{fig:pdp_2dday_500}).
+This is supported in the two-dimensional partial dependence plot (Figure \ref{fig:pdp_2dnight_100}) where the higher temperatures are observed when there is high \% building area and low sky view factor.
+It follows that heat is being captured in the canyons. 
+However, compared to the other urban characteristics this has a lesser affect, changing the temperature by approximately 1.5$^o$C.
+The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_100} and \ref{fig:pdp_2dday_100}), but suggests that there is higher temperatures when the sky view factor is high.
 
 \textit{Population density}.
-Contrary to expectation (e.g. \cite{Peng2012-iy}), we found that population density had no discernible affect during night or day (Figure \ref{fig:importance_500}).
+We found that population density had no discernible effect during night or day (Figure \ref{fig:importance_100}).
 
 \textit{Result sensitivity}.
-To assess the robustness of the results, we conduct the analysis again at the 100 meter resolution (\S \ref{ss:100_meter}).
+To assess the robustness of the results, we conduct the analysis again at the 500-meter resolution (\ref{ss:500_meter}).
 The results are consistent.
-Additionally, to ensure that the affects are consistent between cities, we construct the partial dependence plots for each city (\S \ref{ss:city}).
-The partial dependence is similar to Figure \ref{fig:pdp_500}.
-Therefore these results are robust.
+Additionally, to ensure that the effects are consistent between cities, we construct the partial dependence plots for each city (\ref{ss:city}).
+The partial dependence is similar to Figure \ref{fig:pdp_100}.
+Therefore, the consistency between 100 and 500-meter lends confidence to these conclusions.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_2d_night_500.pdf}
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_night_100.pdf}
     \caption{
-    \textbf{Night time, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
+    \textbf{Nighttime, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
     }
-    \label{fig:pdp_2dnight_500}
+    \label{fig:pdp_2dnight_100}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_2d_day_500.pdf}
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_day_100.pdf}
     \caption{
-    \textbf{Day time, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
+    \textbf{Daytime, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
     }
-    \label{fig:pdp_2dday_500}
+    \label{fig:pdp_2dday_100}
 \end{figure*}
@@
 \section{Conclusion}
-To assist planners tackling heat waves, climate change and high urban temperatures generally, we seek to determine how different urban characteristics are associated with land surface temperatures. 
+To assist planners tackling climate change, heat waves, and general high urban temperatures, we seek to determine how different urban characteristics are associated with land surface temperatures. 
 The results strongly support initiatives for increasing green infrastructure in cities \cite{Larsen2015-da, Meerow2017-xv}. 
 We found that increasing vegetation and reducing impervious surface area had the greatest effect on temperature during both night and day, with the potential to reduce the temperature up to 10$^o$C.
-We also see strong evidence for blue space and increasing the area of water, although there various limitations with doing so in cities such as Phoenix.
+We also see strong evidence for blue space and increasing the area of water, although there are various limitations with doing so in cities such as Phoenix.
+We do note that this analysis is for land surface temperature, and not air temperature. 
+The correlation between LST and air temperature is not perfect at intra-urban scales, and the magnitude of LST variability tends to be larger  than that of air temperature.
+Nevertheless, LST is a highly relevant variable for analyses of the urban energy balance, and thus for studies of the UHI process.
 
@@
 This result opens opportunities for further detailed analysis into potential interventions.
-Such interventions for mitigating high temperatures are naturally place-specific, and while these results have proven general to four different US cities, work is needed to understand questions such as what types of greenery are better than others \cite{Gober2009-im}.
-This study also demonstrates the importance of improved statistical analysis in these studies, given that the results observed constrasted with some existing conclusions.
-At the very least, studies need to appropriately validate their models using cross-validation in future.
-Another future step is conducting a thorough nonlinear statistical analysis that accounts for seasonality to ensuring that interventions are effective year-round.
+Such interventions for mitigating high temperatures are naturally place-specific, and while these results have proven general to four different US cities, work is needed to understand questions such as which types of greenery are better than others \cite{Gober2009-im}.
+This study also demonstrates the opportunities of modern statistical techniques and the ability to assess potentially non-linear interactions and the interactions between multiple variables. 
+Another future step is conducting a thorough nonlinear statistical analysis that accounts for seasonality to ensure that interventions are effective year-round.
 
@@
 If this type of action is explored, a further necessary study is to compare sprawling and dense cities, prior to taking steps to dissuade density. 
-Now that 30 meter resolution nighttime data is available, as well as the increasing availability of lidar data, studies advocating creative interventions such as increasing the vertical and horizontal randomness of buildings, increasing the prevalence of green roofs \cite{Gago2013-ta}, or other urban planning strategies can quantitatively be explored.
+Now that 100-meter resolution nighttime data is available, as well as the increasing availability of lidar data, studies advocating creative interventions such as increasing the vertical and horizontal randomness of buildings, increasing the prevalence of green roofs \cite{Gago2013-ta}, or other urban planning strategies can quantitatively be explored.
 
@@
 It has also demonstrated the utility of leveraging advanced statistical analysis to study land surface temperature.
-The results are robust to both data and model uncertainty, and are general across the cities studied.
+The results are robust to both data and model uncertainty and are general across the cities studied.
 They suggest that vegetation and impervious surfaces are the most important urban characteristics associated with land surface temperature.
@@
 \newpage
-\section{Statistical Models}
+\section{Covariates included in models}
 
-\lbrack Here I'll provide in-depth explanation of the statistical models used \rbrack
+We select variables to include in the model based on the variance inflation factor (VIF). 
+This approach removes variables that exhibit high multicollinearity, that is the variable can be predicted from a combination of other variables. 
+It is important to remove variables with high multicollinearity in inferential studies because otherwise the variables may confound the effect of one another on the variables of interest.
+
+\begin{tabu}to \textwidth{ lr }
+ \hline
+ Resolution & Included variable \\
+ \hline
+100-meter  & Albedo mean \\
+& NDVI mean \\
+& Sky view factor mean \\
+& \% tree canopy stand. dev. spatial lag \\
+& \% tree canopy stand. dev. \\
+& NDBI stand. dev. spatial lag \\
+& \% building area\\
+& Sky view factor max \\
+& \% tree canopy mean \\
+& NDVI stand. dev. \\
+& \% water area\\
+& Digital surface model mean \\
+& Population density mean\\
+\hline
+500-meter & NDVI mean \\ 
+& Albedo mean \\
+& Sky view factor mean \\
+& Digital surface model stand. dev. \\
+& \% building area\\
+& \% tree canopy mean \\
+& \% water area\\
+& Sky view factor max \\
+& NDBI max \\
+& \% tree canopy max \\
+& Population density mean\\
+& Digital surface model mean \\
+& \% tree canopy min \\
+\hline
+\label{tab:vif}
+\end{tabu}
+
+\newpage
+\section{Technical appendix: convolutional neural network}
+\label{ss:cnn}
+\subsection{Overview}
+
+Our CNN model was adapted from a U-Net architecture \cite{unet}.
+This architecture is commonly used for image segmentation: the input to a U-Net model is a 2D greyscale image, and the output is another 2D image with values representing the classification of each pixel.
+The novel property of a U-Net CNN is that the internal layers learn at various resampled resolutions of the input data, which makes the model good at learning phenomena that are driven at multiple scales.
+
+Several modifications were made to adapt U-Net to a geospatial use-case, and are outlined below.
+
+\subsection{Data preparation}
+
+The gridded data for each city was treated as an image, with each pixel representing a 100m or 500m cell.
+Instead of having three channels (red, green, and blue) like a colour image, these images had one channel for each of the independent variables in the dataset.
+The target was a 2D single channel image of the same shape.
+
+CNNs require all inputs to be the same shape, which isn't the case for our city domains.
+Each city image was therefore split into images of $32 \times 32$ pixels ($24 \times 24$ for the 500m resolution).
+Splitting the images up also gives more training samples to work with: each of the small square images is a training sample.
+
+Because the city boundaries aren't square and contain holes, missing data is introduced when placing the data into square images.
+The missing data needs to be filled because neural networks have no built-in way to handle non-real numbers.
+A simple approach like replacing with the median for each variable would result in unrealistic abrupt spatial jumps near the city boundaries: these discontinuities would affect the convolution operations in the CNN which can rely on features such as edges and spatial variance.
+Instead, holes and concave boundaries were filled using linear interpolation, then edges were extended by setting missing-data corner pixels to the median value of each variable and performing linear interpolation.
+The progression of the missing data filling algorithm is shown in Figure~\ref{fig:cnn_missing_data}.
+The result is all-real images with smooth changes at the missing data boundaries.
+
+\begin{figure*}[h]
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/cnn_data_prep.pdf}
+    \caption{Handling of missing data for the CNN. The image shown is for a $32 \times 32$ cell section of mean day temperature for Portland at the 100m resolution.}
+    \label{fig:cnn_missing_data}
+\end{figure*}
+
+\subsection{Model}
+
+The model architecture is shown in Figure~\ref{fig:cnn_architecture}.
+It consists of multiple convolutional layers at both the input resolution and a 50\% downsampled resolution.
+Additionally, a skip connection concatenates the raw input data with one of final layers to reduce the depth between the input and output.
+
+\begin{figure*}[h]
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/cnn_architecture_cropped.pdf}
+    \caption{CNN model architecture. Each layer shows the spatial dimensions (e.g., $32^2$) and the number of channels (e.g., 12). In the final {$32^2 \times 1$} layer, the ReLU layer is omitted.}
+    \label{fig:cnn_architecture}
+\end{figure*}
+
+The size of the network is much smaller compared to the original U-Net in order to reduce overfitting due to our small dataset.
+
+The final layer uses a linear activation function to enable regression. All other layers use ReLU activation, and dropout is applied after the contraction.
+
+To prevent the CNN overfitting to the naively-imputed missing data, a masked loss function was used. The mean square error was greatly reduced for cells $i$ with missing data according to a mask $M$
+
+\begin{align*}
+    \mathrm{loss}_i &= M_i(y_i - \hat y _i) ^ 2\\
+    M_i &= \begin{cases}
+      0.01, & \text{if}\ y_i \mathrm{undefined} \\
+      1, & \text{otherwise}
+    \end{cases}
+\end{align*}
+
+This reduces the impact missing data has on the model weights, while leaving a small amount of gradient to avoid numerical issues with lack of convergence.
+The cells with missing data were excluded from any results presented.
+
+The mask was also added to the input data as an additional channel.
 
@@
     \includegraphics[width=\linewidth]{fig/report/pdp_cities_500.png}
-    \caption[City specific partial dependence plots]{
+    \caption[City specific partial dependence plots at the 500-meter resolution]{
+    City specific partial dependence plots at the 500-meter resolution.
     The partial dependence plots for a random forest model trained on each city.
@@
     \includegraphics[width=\linewidth]{fig/report/pdp_cities_100.png}
-    \caption[City specific partial dependence plots]{
-    The partial dependence plots for a random forest model trained on each city.
+    \caption[City specific partial dependence plots at the 100-meter resolution]{
+    City specific partial dependence plots at the 100-meter resolution.
+    These are partial dependence plots for a random forest model trained on each city.
     This is to evaluate whether the influence of urban characteristics on LST is consistent between the cities.
@@
 \clearpage
-\section{100 meter resolution results}
-\label{ss:100_meter}
-The figures presented in the main text, are replicated here based on data at a 100 meter resolution.
-\\
-\lbrack Discuss the VIF results  \rbrack\\
-\lbrack Discuss the results of the variable influence - why is water not recognized  \rbrack
+\section{500-meter resolution results}
+\label{ss:500_meter}
+The figures presented in the main text, are replicated here based on data at a 500-meter resolution.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/holdout_100.png}
-    \caption[Holdout cross-validation results at 100 meter resolution]{
-    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
+    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
+    \caption[Holdout cross-validation results at 500-meter resolution]{
+    Holdout cross-validation results at 500-meter resolution. 
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 500-fold holdout cross-validation. 
     The models were trained on 80\% of the data and tested on the unseen 20\%.
@@
     Good models have MAE near 0.
-    This is for the 100 meter resolution.
     }
-    \label{fig:holdout_100}
+    \label{fig:holdout_500}
 \end{figure*}
@@
     \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/importance_100.png}
-    \caption[Variable influence on LST at 100 meter resolution]{
-    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature at 100 meter resolution.}
-    \label{fig:importance_100}
+    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
+    \caption[Variable influence on LST at 500-meter resolution]{
+    Variable influence on LST at 500-meter resolution.
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
+    \label{fig:importance_500}
     \end{center}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_100.png}
-    \caption[Partial dependence plots for LST at 100 meter resolution]{
+    \includegraphics[width=\linewidth]{fig/report/pdp_500.png}
+    \caption[Partial dependence plots for LST at 500-meter resolution]{
+    Partial dependence plots for LST at 500-meter resolution.
     Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each urban characteristic as the other variables are held at their average (mean) value. 
@@
     The histograms on the $x$-axis shown the distribution of the observed data.
-    This is for the 100 meter resolution.
     }
-    \label{fig:pdp_100}
+    \label{fig:pdp_500}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_2d_night_100.pdf}
-    \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_night_500.pdf}
+    \caption[Partial dependence contour plots for LST at 500-meter resolution during the night]{
+    Partial dependence contour plots for LST at 500-meter resolution during the night.
+    These partial dependence contours show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
 The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
@@
     }
-    \label{fig:pdp_2dnight}
+    \label{fig:pdp_2dnight_500}
 \end{figure}
@@
     \centering
-    \includegraphics[width=\linewidth]{fig/report/pdp_2d_day_100.pdf}
-    \caption{
-    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
+    \includegraphics[width=\linewidth]{fig/report/pdp_2d_day_500.pdf}
+    \caption[Partial dependence contour plots for LST at 500-meter resolution during the day]{
+    Partial dependence contour plots for LST at 500-meter resolution during the day.
+    These partial dependence contours show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
 The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
@@
     }
-    \label{fig:pdp_2dday}
+    \label{fig:pdp_2dday_500}
 \end{figure}

### 2020-01-16 code/processing/L8_processing.py
formattign changes

@@ -103,7 +103,7 @@ def read_metadata(info_satellite):
     logger.info('Reading metadata')
     # metadata file and location
     fn_metadata = 'data/raw/{}/{}_MTL.txt'.format(info_satellite['city'],info_satellite['landsat_product_id'])
-
+    #
     # list of variables needed from metadata
     meta_variables = set(['K1_CONSTANT_BAND_10','K2_CONSTANT_BAND_10'])
     bands = [1,2,3,4,5,6,10]
@@ -111,7 +111,7 @@ def read_metadata(info_satellite):
     for rad in ['MULT', 'ADD']:
         for b in bands:
             meta_variables.add('RADIANCE_{}_BAND_{}'.format(rad, b))
-
+    #
     # init dictionary
     meta_dict = dict.fromkeys(meta_variables)
     # open the metadata file
@@ -127,7 +127,7 @@ def read_metadata(info_satellite):
                 var = list(var)[0]
                 # add to dictionary
                 meta_dict[var] = float(elements[-1])
-
+    #
     return meta_dict
 
 
@@ -146,7 +146,7 @@ def clip_geographic_data(info_satellite, source_city):
     # Define command
     command = 'Rscript'
     path2script = 'code/processing/clip_geographic_data.R'
-
+    #
     # Define arguments
     city = info_satellite['city']
     city_idx = source_city.loc[source_city['city']==city].index
@@ -160,10 +160,10 @@ def clip_geographic_data(info_satellite, source_city):
     bands = '1,2,3,4,5,6,10'
     # args into list - order does not matter, it is sorted at the bottom of the R script
     args_clip = [city, landsat_product_id, fn_land_cover, fn_boundary, bands, fn_tree_canopy, fn_impervious_surface, fn_elevation]
-
+    #
     # Build subprocess command
     cmd = [command, path2script] + args_clip
-
+    #
     # run the R function to clip and project the data as required
     x = subprocess.check_output(cmd, universal_newlines=True)
 
@@ -296,7 +296,7 @@ def atmos_correction(temp_satellite, info_satellite, emissivity):
     b_6 = 0.458606
     w = 1.6
     t_6 = 0.974290 - 0.08007*w
-    # variables dependent on land cover (emissivity) and temperature
+        # variables dependent on land cover (emissivity) and temperature
     c_6 = emissivity * t_6
     d_6 = (1 - t_6)*(1 + (1 - emissivity)*t_6)
     t_a = 16.0110 + 0.92621*temp_max

### 2020-01-16 report/article.tex

@@
 % \documentclass[final,3p,times,onecolumn,sort&compress]{elsarticle}
-\documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
-%\documentclass[review]{elsarticle}
+% \documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
+\documentclass[review]{elsarticle}
 
-\usepackage{lineno,hyperref}
+\usepackage{lineno}
 \modulolinenumbers[2]
@@
 %% Packages
+\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
 \usepackage{tabu}
+\usepackage{booktabs}
 \usepackage{breakurl}
@@
 \usepackage{amsmath}
+\usepackage{setspace}
 
@@
 %% or include affiliations in footnotes:
-\author[1]{T.M. Logan\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
+% \author[1]{T.M. Logan\corref{mycorrespondingauthor}}
+% \cortext[mycorrespondingauthor]{Corresponding author}
 % \ead[url]{www.tomlogan.co.nz}
-\ead{tom.logan@canterbury.ac.nz}
+% \ead{tom.logan@canterbury.ac.nz}
 
-\author[2]{B. Zaitchik}
-\author[1]{S. Guikema}
-\author[]{A. Nisbet}
-% \footnote{Beginning September 2019, T Logan will be a faculty member at the University of Canterbury, NZ}
+% \author[2]{B. Zaitchik}
+% \author[3]{S. Guikema}
+% \author[]{A. Nisbet}
 
-\address[1]{Industrial and Operations Engineering, University of Michigan, Ann Arbor, MI}
-\address[2]{Earth and Planetary Sciences, Johns Hopkins University, Baltimore, MD}
+% \address[1]{Civil and Natural Resources Engineering, University of Canterbury, New Zealand}
+% \address[2]{Earth and Planetary Sciences, Johns Hopkins University, Baltimore, MD}
+% \address[3]{Industrial and Operations Engineering, University of Michigan, Ann Arbor, MI}
 
@@
 Our results also demonstrate the potential for using nonlinear statistical analysis to investigate land surface temperature and its relationships with urban characteristics. 
+For example, we can evaluate the compounding influence of two urban characteristics on the temperature.
 Improved understanding of these relationships influencing both night and day land surface temperature will assist planners undertaking climate change adaptation and heat wave mitigation.
@@
 Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy}.
-However, the mechanisms and urban characteristics driving LST allegedly differs between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
-Given that UHI is a nocturnal phenomena, the lack of study on nighttime LST leaves a critical gap in our understanding. 
+However, the mechanisms and urban characteristics driving LST allegedly differ between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
+Given that UHI is primarily a nocturnal phenomena, the lack of study on nighttime LST leaves a critical gap in our understanding. 
 This can now be addressed as high resolution nighttime satellite images have become available.
 
-Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
+Most existing studies that have analyzed LST, particularly nighttime LST, have used low-resolution imagery, making it challenging to deduce a clear understanding of the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
 Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relaxing their linear assumption, 4) rigorously testing and validating their models.
@@
 The question we address is: \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
-To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
+To achieve this we conduct a comparative study of four cities in the United States using nonlinear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
 In addressing our question, we will address the, at times fundamental, limitations of existing studies.
 We look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
-The results have implications for mitigating the severity of future heat waves.
+While we do not claim causality, understanding the associations between urban characteristics on temperatures can complement and further inform our understanding of the processes that lead to high urban land surface temperature.
+The results therefore have implications for mitigating the severity of future heat waves.
 
@@
 We correct for emissivity using land cover data \cite{Alipour2003-ym}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
-Finally, an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
+Finally, an atmospheric correction is made as per the mono-window algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
 This follows the process is described in \cite{Scott2016-lc}.
@@
 \subsection{Data preparation and robustness}
-\begin{figure*}
-    \centering
-    \includegraphics[width=\linewidth]{fig/report/holdout_100.png}
-    \caption[Holdout cross-validation results at 100-meter resolution]{
-    Holdout cross-validation results at 100-meter resolution.
-    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
-    The models were trained on 80\% of the data and tested on the unseen 20\%.
-    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
-    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
-    Good models have MAE near 0.
-    }
-    \label{fig:holdout_100}
-\end{figure*}
 
@@
 
+It is important to note that we are assessing the association between the urban characteristics and the land surface temperature.
+These models are not explicitly evaluating causality. 
+Instead, we use these models to control for urban characteristics and evaluate how land surface temperature changes with changes in each of the other urban characteristic in turn.
+
 \textit{Null model: average}. The first model, to compare other models against, is the null model. 
@@
 MARS models are a type of non-parametric regression approach that are useful for high dimensional problems \cite{Hastie2009-ky}. 
-They extend the linear model by using piecewise functions to fit the data \cite{Friedman1991-of}. 
-Multiple linear approximations are summed across the variable space to construct an estimate of the  LST\cite{Friedman1991-of}.
+They extend the linear model by using piecewise functions to fit the data \cite{Friedman1991-of}: $$ y = \sum_i c_i B_i(x)$$
+$B_i(x)$ is the basis function that includes multiple linear functions and indicator variables that set these functions to zero for certain ranges of $x$.
+That is, multiple linear approximations are summed across the variable range and provide a nonlinear estimate of the LST \cite{Friedman1991-of}.
 
@@
 There can be multiple layers of these neurons that pass the results as an input to subsequent layers.
-In this manner, neural networks can capture non-linearities.
+In this manner, neural networks can capture nonlinearities.
 The advantage of convolutional neural networks above other types of neural networks is their ability to capture the spatial dependencies. 
@@
 \section{Results and Discussion}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\linewidth]{fig/report/holdout_100.png}
+    \caption[Holdout cross-validation results at 100-meter resolution]{
+    Holdout cross-validation results at 100-meter resolution.
+    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
+    The models were trained on 80\% of the data and tested on the unseen 20\%.
+    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
+    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
+    Good models have MAE near 0.
+    }
+    \label{fig:holdout_100}
+\end{figure*}
+
 In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
@@
 The R$^2$ results, also based on unseen data, suggest that more than 90\% of the data variance is captured by the best models.
-These are strong results.
 We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees and convolutional neural network.
@@
 This is thought to be less important during the night because latent and sensible heat are dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
-However, because transpiration does not occur at night, vegetation's effect on night temperatures is debated. 
-In our study, percentage tree canopy and impervious surface cover must be discussed together because they are 100\% correlated (Figure \ref{fig:imp_tree}).
-To distinguish between vegetated surfaces and other pervious surfaces, we also include NDVI, a measure of a surface's greenness.
-At night, we find that \% tree canopy is the most influential characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:pdp_100}).
-These partial dependence results indicate a decrease in up to 10$^oC$ as the percentage area of trees or other pervious surface increases.
+Indeed, because transpiration does not occur at night, vegetation's effect on night temperatures is debated. 
+In our study, percentage tree canopy and impervious surface cover must be discussed together because they are 100\% correlated in the data (Figure \ref{fig:imp_tree}).
+To distinguish between vegetated surfaces and other pervious surfaces, we include NDVI, a measure of a surface's greenness.
+At night, we find that \% tree canopy (and \% impervious surface) is the most influential characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:importance_100}).
+The partial dependence results (Figure \ref{fig:pdp_100}) indicate a decrease in up to 10$^oC$ as the percentage area of trees or other pervious surface increases.
 This reduction in temperature could be due to a reduction in the impervious surfaces that store heat.
-However, we also observe that as the NDVI increases, the land surface temperature decreases.
-The two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_100}) evaluates how LST varies with both characteristics.
-Figure \ref{fig:pdp_2dnight_100}d shows how the LST changes as both tree canopy/impervious surface and the NDVI change.
-The greener and more pervious a surface is, the cooler it is during the night. 
-The total change here is also approximately 7.5$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less variable of the five models shown in Figure \ref{fig:pdp_100}.
+We also observe that as the NDVI increases, the land surface temperature decreases.
+
+To further evaluate whether the influence with temperature is due to vegetation or pervious surfaces, consider the the two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_100}d) that evaluates how LST varies with \% tree canopy/impervious surface and the NDVI.
+The greener (higher NDVI) and more pervious (higher \% tree canopy cover) a surface is, the cooler it is during the night. 
+The total change here is also approximately 7.5$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less sensitive of the five models (as seen in Figure \ref{fig:pdp_100}).
 These results contradict standing conclusions that vegetation has no effect on nighttime LST \cite{Peng2012-iy, Zhou2014-wc};
-rather, vegetation appears to be strongly associated with lower land surface temperatures during the night.
+although much of the reduction in temperature could be due to the perviousness of the surface, lowered temperatures appear to be associated with the greenness (NDVI) of a surface during the night.
 
@@
 The rationale is that water releases heat during the night, resulting in elevated nighttime temperature \cite{Chun2017-mm}.
-However, while this is intuitive for air temperature, the effect on LST is not as apparent.
-Our findings show that water reduces the LST during both the day and night (Figure \ref{fig:pdp_500}).
+Our findings, however, show that water reduces the LST during both the day and night (Figure \ref{fig:pdp_500}).
 Although these results are not supported at the 100-meter resolution because 98\% of the data at the 100-meter level has zero percentage water.
@@
 The built-up index (NDBI), \% building area, and digital surface elevation model also had relatively minor associations with LST.
-The building area had no effect during the night (Figure \ref{fig:pdp_100}a) and increased temperature during the day (Figure \ref{fig:pdp_100}c).
-The largest effect observed is that maximum NDBI is associated with maximum daytime LST (Figure \ref{fig:pdp_100}).
-This is surprising given that NDBI has been reported as among the most important urban characteristics during the day \cite{Peng2018-cp}.
-However, the discrepancy may be due to \% impervious surface area being incorporated already.
+The elevation (non-built-up) variable was excluded because it was found to be collinear with other variables; the digital surface elevation model was included and represents the elevation including buildings and vegetation.
+However, this elevation only appears to have minor associations with land surface temperature (Figure \ref{fig:pdp_100} \& \ref{fig:pdp_500}).
+The building area had no effect during the night and minorly increased temperature during the day (Figure \ref{fig:importance_100}).
+Maximum NDBI appears to have a slight positive associated with maximum daytime LST (Figure \ref{fig:pdp_500}), but this effect is not observed at the 100m resolution.
+This minor-to-negligible relationship is surprising given that NDBI has been reported as among the most important urban characteristics during the day \cite{Peng2018-cp}.
+However, the discrepancy may be due to \% impervious surface area being incorporated already with the \% tree canopy data.
 
@@
 While there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_100}a,b).
-This is potentially due to heat being stored within the canyons (areas with low, sky view factors).
+This is potentially due to heat being stored within the canyons (areas with low sky view factors).
 This is supported in the two-dimensional partial dependence plot (Figure \ref{fig:pdp_2dnight_100}) where the higher temperatures are observed when there is high \% building area and low sky view factor.
@@
 However, compared to the other urban characteristics this has a lesser affect, changing the temperature by approximately 1.5$^o$C.
-The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_100} and \ref{fig:pdp_2dday_100}), but suggests that there is higher temperatures when the sky view factor is high.
+The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_100} and \ref{fig:pdp_2dday_100}), but suggests that there are higher temperatures when the sky view factor is high.
 
@@
 Therefore, the consistency between 100 and 500-meter lends confidence to these conclusions.
+
+\textit{Urban strategies}.
 
@@
 Such interventions for mitigating high temperatures are naturally place-specific, and while these results have proven general to four different US cities, work is needed to understand questions such as which types of greenery are better than others \cite{Gober2009-im}.
-This study also demonstrates the opportunities of modern statistical techniques and the ability to assess potentially non-linear interactions and the interactions between multiple variables. 
-Another future step is conducting a thorough nonlinear statistical analysis that accounts for seasonality to ensure that interventions are effective year-round.
+This study also demonstrates the opportunities of modern statistical techniques and the ability to assess potentially nonlinear interactions and the interactions between multiple variables.
 
 Rigorous statistical analysis can continue to answer on-going questions central to land surface temperature.
-For example, our results support the suggestion that 3D metrics do not outperform 2D ones \cite{Berger2017-lx}, however, iteratively removing these variables would address the potential for conflating the importance of different characteristics \cite{Chun2017-mm}.
+For example, our results support the suggestion that 3D variables (e.g. sky view factor) do not outperform 2D ones (e.g. \% tree canopy cover) \cite{Berger2017-lx}. 
+However, iteratively removing these variables from the analysis can ameliorate the potential for conflating the importance of different characteristics \cite{Chun2017-mm}.
+Our analysis is not causal inference; such an analysis would require controlled variables and urban spaces.
+We instead aim to complement and inform existing understanding of the relationships by demonstrating how land surface temperature and the urban characteristics are associated.
+This can be used to inform potential strategies for urban heat reduction, that is, if an association exists between two variables then we can evaluate the potential for causality based on understanding the physical mechanisms.
 
-Additionally, our results do not support claims advocating to constrain floor area ratios \cite{Chun2017-mm}.
+However, if no association exists or is in the opposite direction of proposed strategies for heat reduction, then such strategies need to be re-evaluated.
+For example, our results assessing the relative effect of sky view factor (a measure of building density), population, built-up index, and \% building area do not support claims advocating to constrain floor area ratios \cite{Chun2017-mm}.
 Such policy could lead to increased sprawl.
-If this type of action is explored, a further necessary study is to compare sprawling and dense cities, prior to taking steps to dissuade density. 
+Instead, our results show that reducing impervious area and increasing vegetation and greenness have a stronger association with lowered temperatures.
+Such action could be achieved without constraining floor area ratios.
+If a constraint is explored, an additional necessary study is to compare sprawling and dense cities, prior to taking steps to dissuade density. 
 Now that 100-meter resolution nighttime data is available, as well as the increasing availability of lidar data, studies advocating creative interventions such as increasing the vertical and horizontal randomness of buildings, increasing the prevalence of green roofs \cite{Gago2013-ta}, or other urban planning strategies can quantitatively be explored.
@@
 \section*{Acknowledgements}
-This research was conducted with the support of the University of Michigan's Rackham PreDoctoral Fellowship, and the National Science Foundation's Grant SEES-1631409.
-This support is gratefully acknowledged.
+% TML conducted this research as a graduate student at the University of Michigan with the support of a Rackham PreDoctoral Fellowship.
+% The work was also supported by the US National Science Foundation's Grant SEES-1631409. 
+% This support is gratefully acknowledged.
 
 \section*{References}
+\singlespacing
 
@@
 \section{Data sources}
+The data sources for the analysis are listed in Table \ref{tab:data}.
 
+\begin{table}[H]
+\caption{The data sources for the LST analysis.}
+\label{tab:data}
+\fontsize{8}{11}\selectfont
 \begin{tabu}to \textwidth{ X[l]  X[c]  X[c] X[l] X[l] }
- \hline
- Data Provider & Data Type & Data Date & Description & URL \\
+\toprule
+ Data Provider & Data Type & Data Date & Description & Source \\
  \hline
 U.S. Geological Survey  & Raster  & 2013-2017 &
-    Landsat 8 day and night satellite imagery & \burl{https://earthexplorer.usgs.gov/} \\
+    Landsat 8 day and night satellite imagery & \url{https://earthexplorer.usgs.gov/} \\
 Microsoft  & Polygon  & 2018 & Building footprint polygons for the US &
-    \burl{https://github.com/Microsoft/USBuildingFootprints} \\
+    \url{https://github.com/Microsoft/USBuildingFootprints} \\
 Defense Meteorological Satellite Program  & Raster  & 2013 &
-    Stable nighttime light intensity & \burl{https://www.ngdc.noaa.gov/eog/dmsp/downloadV4composites.html} \\
+    Stable nighttime light intensity & \url{https://www.ngdc.noaa.gov/eog/dmsp/downloadV4composites.html} \\
 Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
-    Land cover & \burl{https://www.mrlc.gov} \\
+    Land cover & \url{https://www.mrlc.gov} \\
 Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
-    Percent developed imperviousness & \burl{https://viewer.nationalmap.gov} \\
+    Percent developed imperviousness & \url{https://viewer.nationalmap.gov} \\
 Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
-    Percent tree canopy cover & \burl{https://viewer.nationalmap.gov} \\
+    Percent tree canopy cover & \url{https://viewer.nationalmap.gov} \\
 U.S. Geological Survey  & Raster  & 2015 & 1/3 arc-second elevation &
-    \burl{https://nationalmap.gov/3DEP/3dep_prodserv.html} \\
+    \url{https://nationalmap.gov/3DEP/3dep_prodserv.html} \\
 U.S. National Oceanic and Atmospheric Administration  & Lidar  & 2014 & Point cloud of surface elevation &
-    \burl{https://coast.noaa.gov/htdata/lidar2_z/geoid12b/data/6377/} \\
+    \url{https://coast.noaa.gov/htdata/lidar2_z/geoid12b/data/6377/} \\
 IPUMS NHGIS  & Area Level  & 2010 &
     Block-level population from the US census & \cite{nhgis}\\
-\hline
-\label{tab:data}
+\bottomrule
 \end{tabu}
+
+\end{table}
 
@@
 
-\begin{tabu}to \textwidth{ lr }
- \hline
- Resolution & Included variable \\
- \hline
+\begin{table}[h]
+% \small
+% \fontsize{8}{11}\selectfont
+\centering
+\caption{Covariates included after accounting for multicollinearity.}
+\label{tab:vif}
+\begin{tabular}{p{0.23\linewidth} p{0.7\linewidth}}
+\toprule
+       \textbf{Resolution} & \textbf{Included variable} \\
+\midrule
 100-meter  & Albedo mean \\
@@
 & \% tree canopy min \\
-\hline
-\label{tab:vif}
-\end{tabu}
+\bottomrule
+\end{tabular}
+\end{table}
 

### 2020-01-17 code/analysis/analysis.py
creating stacked bar plot

@@ -41,19 +41,19 @@ RANDOM_SEED = 3201
 
 # other constants
 city_names = {'bal':'baltimore','por':'portland','phx':'phoenix','det':'detroit'}
-model_names = {'rf':'random forest','mlr':'multivariate linear',
-                'gam':'generalized additive\n(gam)',
-                'gbrt':'gradient boosted\ntrees',
-                'mars':'multivariate adaptive\nspline (mars)',
-                'cnn':'convolutional\nneural network'}
-feature_names = {'lcov_11' : '% water','tree_mean':'% tree canopy','ndvi_mean':'ndvi',
-                'svf_mean':'sky view factor','dsm_mean':'digital surface model',
-                'alb_mean':'albedo','dsm_sd':'dsm stand. dev.','nbdi_max':'max nbdi',
-                'tree_max':'max % tree can.','bldg':'% building area','pdens_mean':'pop. density',
-                'tree_min':'min % tree can.','svf_max':'max sky view factor',
-                'tree_sd': '% tree can. stand. dev.', 'nbdi_sd_sl':'nbdi surrounding stand. dev.',
-                'tree_sd_sl':'% tree can. surrounding stand. dev.',
-                'ndvi_sd': 'ndvi stand. dev'
+model_names = {'rf':'random forest','mlr':'linear',
+                'gam':'generalized additive (gam)',
+                'gbrt':'g. boosted trees',
+                'mars':'multiv. adaptive spline (mars)',
+                'cnn':'c. neural network'}
+feature_names = {'lcov_11' : 'water %','tree_mean':'tree can. % mean','ndvi_mean':'ndvi',
+                'svf_mean':'sky view factor mean','dsm_mean':'d. surface model mean',
+                'alb_mean':'albedo mean','dsm_sd':'d. surface model sdev','nbdi_max':'nbdi max',
+                'tree_max':'tree can. % max','bldg':'building % area','pdens_mean':'pop. density',
+                'tree_min':'tree can. % min','svf_max':'sky view factor max',
+                'tree_sd': 'tree can. % sdev.', 'nbdi_sd_sl':'nbdi surrounding sdev.',
+                'tree_sd_sl':'tree can. % surrounding sdev.',
+                'ndvi_sd': 'ndvi sdev'
                 }
 
 CORES_NUM = min(50,int(os.cpu_count()*3/4))
@@ -1039,6 +1039,68 @@ def plot_importance(results_swing, grid_size):
         # fig.set_size_inches(15,20)
 
[…]

### 2020-01-17 code/analysis/analysis.py
variable importance stacked bar

@@ -1094,8 +1094,8 @@ def plot_importance_stacked(results_swing, grid_size):
 
         g.set_xlabel("")
         g.set_ylabel("variable\ninfluence")
-        g.set_ylim(0,2)
-        g.axes.xaxis.set_ticklabels([])
+        g.set_ylim(0,2.5)
+        # g.axes.xaxis.set_ticklabels([])
         g.legend().remove()
         # g.legend(loc='lower center', ncol=3, frameon=False)
 

### 2020-01-17 code/analysis/analysis.py
final changes to the stacked bar varimpo figures

@@ -1095,7 +1095,7 @@ def plot_importance_stacked(results_swing, grid_size):
         g.set_xlabel("")
         g.set_ylabel("variable\ninfluence")
         g.set_ylim(0,2.5)
-        # g.axes.xaxis.set_ticklabels([])
+        g.axes.xaxis.set_ticklabels([])
         g.legend().remove()
         # g.legend(loc='lower center', ncol=3, frameon=False)
 

### 2020-05-03 report/article.tex
accepted into RSE

@@
 % \documentclass[sort&compress]{elsarticle}
-% \documentclass[final,3p,times,onecolumn,sort&compress]{elsarticle}
+\documentclass[final,3p,times,onecolumn,sort&compress]{elsarticle}
 % \documentclass[final,3p,times,twocolumn,sort&compress]{elsarticle}
-\documentclass[review]{elsarticle}
+% \documentclass[review]{elsarticle}
 
@@
 \usepackage{setspace}
+\usepackage{array}
+\newenvironment{conditions}
+  {\par\vspace{\abovedisplayskip}\noindent\begin{tabular}{>{$}l<{$} @{${}={}$} l}}
+  {\end{tabular}\par\vspace{\belowdisplayskip}}
+\usepackage{listings}
+\usepackage{xcolor}
+\lstset{
+  basicstyle=\ttfamily,
+  columns=fullflexible,
+  frame=single,
+  breaklines=true,
+  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
+}
 
@@
 %% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
+\bibliographystyle{model5-names}\biboptions{authoryear}
 
@@
 %% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
+% \bibliographystyle{elsarticle-num}
 
@@
 
-\title{Night and day: What is the influence and relative importance of urban characteristics on land surface temperature?}
+\title{Night and day: The influence and relative importance of urban characteristics on remotely sensed land surface temperature}
 
 %% or include affiliations in footnotes:
-% \author[1]{T.M. Logan\corref{mycorrespondingauthor}}
-% \cortext[mycorrespondingauthor]{Corresponding author}
-% \ead[url]{www.tomlogan.co.nz}
-% \ead{tom.logan@canterbury.ac.nz}
+\author[1]{T.M. Logan\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead[url]{www.tomlogan.co.nz}
+\ead{tom.logan@canterbury.ac.nz}
 
-% \author[2]{B. Zaitchik}
-% \author[3]{S. Guikema}
-% \author[]{A. Nisbet}
+\author[2]{B. Zaitchik}
+\author[3]{S. Guikema}
+\author[4]{A. Nisbet}
 
-% \address[1]{Civil and Natural Resources Engineering, University of Canterbury, New Zealand}
-% \address[2]{Earth and Planetary Sciences, Johns Hopkins University, Baltimore, MD}
-% \address[3]{Industrial and Operations Engineering, University of Michigan, Ann Arbor, MI}
+\address[1]{Civil and Natural Resources Engineering, University of Canterbury, New Zealand}
+\address[2]{Earth and Planetary Sciences, Johns Hopkins University, Baltimore, MD}
+\address[3]{Industrial and Operations Engineering, University of Michigan, Ann Arbor, MI}
+\address[4]{www.ajnisbet.com}
 
 \begin{abstract}
-Heat waves are among the deadliest natural hazards and are expected to increase in frequency and severity under climate change. 
-Their impacts in cities can be exacerbated by the urban heat island (UHI), but the mechanisms that underlie severity and timing of the UHI and its interactions with heat waves are poorly understood.
-Understanding these mechanisms is necessary to design strategies that reduce land surface temperature and mitigate the effect of heat waves.
-Making use of recently available high-resolution day and night thermal satellite imagery and employing advanced nonlinear statistical models, we seek to answer the question \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
+The characteristics of urban land surfaces contribute to the urban heat island, and, in turn, can exacerbate the severity of heat wave impacts.
+However, the mechanisms and complex interactions in urban areas underlying land surface temperature are still being understood.
+Understanding these mechanisms is necessary to design strategies that mitigate land temperatures in our cities.
+Using the recently available night-time moderate-resolution thermal satellite imagery and employing advanced nonlinear statistical models, we seek to answer the question \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
 To answer this question, we analyze urban land surface temperature in four cities across the United States.
-In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
-We model the effects of these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
-The effects from the daytime and nighttime analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
+We devise techniques for training and validating nonlinear statistical models on geostatistical data and use these models to assess the interdependent effects of urban characteristics on urban surface temperature.
 Our results suggest that vegetation and impervious surfaces are the most important urban characteristics associated with land surface temperature.
-Increasing and decreasing these, respectively, is necessary for reducing high urban temperatures during both night and day.
+While this may be expected, this is the first study to quantify this relationship for Landsat-resolution nighttime temperature estimates.
 Our results also demonstrate the potential for using nonlinear statistical analysis to investigate land surface temperature and its relationships with urban characteristics. 
-For example, we can evaluate the compounding influence of two urban characteristics on the temperature.
 Improved understanding of these relationships influencing both night and day land surface temperature will assist planners undertaking climate change adaptation and heat wave mitigation.
@@
 \begin{keyword}
-LandSat; Land surface temperature; Urban canyon; Machine learning; Convolutional Neural Network
+LandSat; Land surface temperature; Urban canyon; Geospatial machine learning; Convolutional neural network; Spatial random forest
 \end{keyword}
@@
 
-\linenumbers
+% \linenumbers
 
 \section{Introduction}
-In a warming world, understanding the factors contributing to high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and wellbeing of their communities.
-And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}.
-Heat waves' effect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
-The UHI is understood to be a product of multiple factors including: enhanced absorption of solar radiation, geometric effects that limit radiative cooling and ventilation, anthropogenic heat from vehicles and buildings, air pollution that traps outgoing radiation, high ratios between sensible and latent heat flux due to low vegetation cover and high impervious land cover, and the radiative characteristics of building materials.
-% \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
-The result is increased diurnal and nocturnal temperatures that reduce people's ability to cool off, especially during the night, which drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.
+In a warming world, understanding the factors contributing to high land surface temperature (LST) will aid in adapting cities in mitigating urban heat for the health and wellbeing of their communities.
+And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \citep{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \citep{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \citep{klinenberg2015heat}.
+Heat waves' effect on people is exacerbated by the urban heat island (UHI) \citep{Wicki2017-fv, Echevarria_Icaza2016-fr}.
+Both the air and surface UHIs are understood to be a product of multiple factors including: enhanced absorption of solar radiation, geometric effects that limit radiative cooling and ventilation, anthropogenic heat from vehicles and buildings, air pollution that traps outgoing radiation, high ratios between sensible and latent heat flux due to low vegetation cover and high impervious land cover, and the radiative characteristics of building materials.
+% \citep{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
+The result is increased diurnal and nocturnal temperatures that reduce people's ability to cool off, especially during the night, which drives an increase in mortality \citep{Echevarria_Icaza2016-fr, Murage2017-wj}.
 
-Studies of urban heat include analyses of air temperature, which is a conventional meteorological variable that is often associated with health outcomes(e.g.
-\cite{Scott2016-lc}), and studies that use satellite-derived land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
-LST correlates with air temperature at large scales, but there are differences at intra-urban scales \cite{Good2016-yk}.
-While LST is not a conventional meteorological variable, it is directly related to the urban heat budget and has the advantage of being available with extensive spatial coverage in gridded satellite products \cite{Hung2006-qy}.
-We analyze LST in this study.
+Studies of urban heat include analyses of air temperature (e.g.
+\cite{Scott2016-lc}) and satellite-derived land surface temperature (LST) \citep{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
+LST correlates with air temperature at large scales, but there are differences at intra-urban scales \citep{Good2016-yk}.
+While LST is not a conventional meteorological variable, it is directly related to the urban heat budget and has the advantage of being available with extensive spatial coverage in gridded satellite products \citep{Hung2006-qy}.
+Most existing LST studies focus on the daytime \citep{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy} and use the LST to infer the surface UHI (SUHI) \cite{Zhou2018-iy}.
+However, the mechanisms and urban characteristics driving LST allegedly differ between night and day \citep{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
+Given that night temperatures have a major effect on health (as high night temperatures prevent the human body from recovering from daytime heat exposure \cite{Laaidi2012-ir}), the lack of study on nighttime LST leaves a critical gap in our understanding.
+Ideally, to study this phenomenon satellite imagery that has both higher spatial resolution and captures an area at regular intervals over a 24-hour period would be available; this would enable analysis to understand the heating and cooling processes that lead to the UHI.
+While the time-series images are currently not available at scale, higher resolution nighttime satellite images have recently become available and these enable us to study the nighttime LST.
 
-Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy}.
-However, the mechanisms and urban characteristics driving LST allegedly differ between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}.
-Given that UHI is primarily a nocturnal phenomena, the lack of study on nighttime LST leaves a critical gap in our understanding. 
-This can now be addressed as high resolution nighttime satellite images have become available.
+There are an increasing number of studies that use satellite remote sensing to analyze land surface temperature (see \cite{Zhou2018-iy} for a review).
+We enhance these studies in the following ways:
+\begin{enumerate}
+    \item We use surface temperature data at night and day at 100-meter resolution.
+    Data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \citep{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
+    1km resolution makes it difficult to attribute LST to urban characteristics \citep{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}. 
+    However, 100m resolution LandSat8 (L8) night scenes have recently become available.
 
-Most existing studies that have analyzed LST, particularly nighttime LST, have used low-resolution imagery, making it challenging to deduce a clear understanding of the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}.
-Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) diversifying their statistical models and relaxing their linear assumption, 4) rigorously testing and validating their models.
-The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
-1km resolution makes it difficult to attribute LST to urban characteristics.
-However, 100m resolution LandSat8 (L8) night scenes have recently become available. 
-Due to the wide spatial availability of L8 imagery, comparative studies between cities that are necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
+    \item We study four cities in the USA.
+    Due to the wide spatial availability of L8 imagery, comparative studies between cities that are necessary to understand the generalizability of findings \citep{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.
 
-Existing studies have also been criticized for the explanatory variables they've used \cite{Chun2017-mm,Peng2018-cp}.
-There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
-Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
-Beyond the 2D or 3D debate, important variables include green space, water bodies, albedo, and socio-economic factors \cite{Peng2018-cp}. 
-However, many studies do not capture these categories and many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
+    \item We include both 2D and 3D urban characteristics in our analysis.
+    Existing studies have been criticized for the explanatory variables they've used \citep{Chun2017-mm,Peng2018-cp}.
+    One point of disagreement is regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.
+    Competing studies suggest that 3D factors are not important \citep{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \citep{Chun2017-mm}.
+    Beyond the 2D or 3D debate, important variables include green space, water bodies, albedo, and socio-economic factors \citep{Peng2018-cp}. 
+    However, many studies do not capture all of these categories and many analyze only the univariate effect of each variable: e.g., from \cite{Peng2018-cp} (p.256) ``many studies used only a single influencing factor to establish the regression model with LST, and compare the individual effects of different factors on LST based on the coefficient of regression.'' (see \cite{Peng2018-cp, Chun2017-mm} for further discussion).
     Considering a variable in isolation, without accounting for potential conflating by other variables, limits the understanding of the interdependent effects that exist.
 
-The third potential enhancement is in the statistical models.
-Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
-Using linear models incurs a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
-The first is that many of the urban characteristics exhibit high multicollinearity \cite{Zhou2014-wc}.
-The second is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \cite{Peng2018-cp, Zhou2014-wc}.
+    \item We use a variety of statistical models that have the ability to capture nonlinear effects and the compounding effects of variables.
+    Almost all studies use linear techniques (e.g.
+\cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water, Galletti2019-wa}.
+    Using linear models can be much simpler than the alternatives, but it incurs limitations on the models' ability to explain the interdependencies between variables. 
+    One limitation is that many of the urban characteristics exhibit high multicollinearity \citep{Zhou2014-wc}.
+    A second issue is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \citep{Peng2018-cp, Zhou2014-wc}.
+    In achieving this we devise a method for training a random forest on spatially auto-correlated data and use a convolution neural network in a geospatial context for the first time to our knowledge.
+    We also use these different statistical models to present the model uncertainty in the effect of each factor.
 
-Finally, most existing studies do not rigorously validate their models.
-The approach they use to assess their accuracy is in-sample validation.
-In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data. 
+    \item We rigorously validate our statistical models.
+    Many existing studies do not rigorously validate their models.
+    The most common approach used to assess statistical accuracy is in-sample validation.
+    In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data \cite{Shmueli2011-wd}.
     This risks overestimating the accuracy of their models.
+    Instead, we test our models on data unused during the model training and subset this data to avoid overfitting.
+\end{enumerate}
 
-The question we address is: \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
+With these enhancements, we seek to answer the question: \textit{``What is the influence and relative importance of urban characteristics on land surface temperature, during both the day and night?''}
 To achieve this we conduct a comparative study of four cities in the United States using nonlinear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
-In addressing our question, we will address the, at times fundamental, limitations of existing studies.
 We look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent, nonlinear, effects of urban characteristics on urban heat.
 While we do not claim causality, understanding the associations between urban characteristics on temperatures can complement and further inform our understanding of the processes that lead to high urban land surface temperature.
-The results therefore have implications for mitigating the severity of future heat waves.
+The results therefore have implications for mitigating the severity of future heat waves and preparing our urban regions for a warmer climate.
 
@@
 \subsection{Cities studied}
-
 We study four cities in the contiguous United States (Figure \ref{fig:map}): Baltimore, MD; Detroit, MI; Phoenix, AZ; and Portland, OR. 
@@
 
-Figure \ref{fig:map} shows the 500m gridded nighttime LST data for each of the cities and Figure \ref{fig:joy} shows the distribution of each city's 100m resolution day and night land surface temperature.
+\subsection{Land surface temperature (dependent variable)}
+We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}).
+Our code is available on our Github\footnote{URL redacted for review} repository, and follows the process described in \cite{Scott2016-lc}:
+\begin{itemize}
+    \item 
+        Convert Band-10 digital-number data to top-of-atmosphere radiance \citep{Jimenez-Munoz2003-wc}: 
+        \begin{equation}
+        \label{eqn:convert_TOA}
+        L_\lambda = M_L Q_{cal} + A_L
+        \end{equation}
+        where:
+        \begin{conditions}
+         L_\lambda & Top-of-atmosphere spectral radiance \\
+         M_L &  Band-specific multiplicative rescaling factor from the satellite metadata \\   
+         A_L &  Band-specific additive rescaling factor from the satellite metadata \\   
+         Q_{cal} &  Quantized and calibrated standard product pixel values
+        \end{conditions}
+    \item Determine the emissivity using land cover data \citep{Alipour2003-ym}, and then calculate the at-satellite brightness temperature \citep{Jimenez-Munoz2003-wc}: 
+        \begin{equation}
+        \label{eqn:at_sat}
+        T_s = \frac{K_2}{ln(1 + \frac{K_1}{L})}
+        \end{equation}
+        where:
+        \begin{conditions}
+         T_s & At-satellite brightness temperature \\
+         L &  Spectral radiance ($L_\lambda$) divided by the emissivity \\   
+         K_1, K_2 & Band-specific thermal conversion constants from the satellite metadata
+        \end{conditions}
+    \item Correct the brightness temperature for atmospheric water vapor using the mono-window algorithm \citep{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station (see the code in \ref{appendix:code}).
+\end{itemize}
+The result is a geospatial dataset of land surface temperature at a 100-meter resolution.
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{fig/report/map_nocturnal_lst.pdf}
-    \caption[The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied]{The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied.
-This is change from the mean (average) for each city.}
-    \label{fig:map}
-\end{figure*}
+To ameliorate the effect of ephemeral changes \citep{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
+The time and dates for the images used are shown in Table \ref{tab:satellite_meta}.
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{fig/report/joyplot_lst_100.pdf}
-    \caption{The distribution of nocturnal and diurnal land surface temperature of the cities studied. 100m resolution.}
-    \label{fig:joy}
-\end{figure}
-
-\subsection{Land surface temperature}
-We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}).
-Our code is available on our Github\footnote{URL redacted for review} repository, and is as follows:
-We convert Band-10 digital-number data to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. 
-We correct for emissivity using land cover data \cite{Alipour2003-ym}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
-Finally, an atmospheric correction is made as per the mono-window algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
-This follows the process is described in \cite{Scott2016-lc}.
-
-To ameliorate the effect of ephemeral changes \cite{Zhou2018-iy} we use at least three minimal-cloud images for each city and night/day period and calculate the mean of LST.
-
-So we can conduct the comparative study between the cities, the land surface temperature for each city is calculated as the difference from the city's mean.
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/scatter_100.png}
-    \caption{The land surface temperature in $^o$C of the cities studied at a 100m square resolution.}
-    \label{fig:scatter_lst}
-    \end{center}
-\end{figure}
+So we can conduct the comparative study between the cities, the land surface temperature is normalized by subtracting the mean (average) value of that city, i.e., we analyze the `difference from the mean' as the response variable for this study.
 
 \subsection{Independent variables}
+The independent variables (i.e., the urban characteristics included to analyze their relationship with land surface temperature) were selected to capture the important factors including green space, water bodies, albedo, and socio-economic factors \citep{Peng2018-cp}.
+The data sources are listed in Table \ref{tab:data}.
+Note that these variables are remotely sensed and therefore our study relies on these remote sensing techniques.
+Because these techniques are established and from well-regarded data sources, we do not include additional ground truth assessments for these parameters.
 
@@
 Albedo is a measure of the reflectivity or brightness of a surface. 
-It is normalized between 0-100 where the darker surfaces are lower values.
+It is normalized between 0-100 where the darker surfaces have lower values. 
 Albedo is calculated using the LandSat8 images using the algorithm described in \cite{Smith2010-nw, Liang2001-jd}. 
+These data are at a 100-meter resolution and are calculated using the daytime satellite images.
 
-\textit{Building floor area}. The building floor area within the cell. This uses data released in 2018 by Microsoft where building footprints are estimated using areal images.
+\textit{Building floor area}. 
+The building floor area within the cell. 
+This uses data released in 2018 by Microsoft where building footprints are estimated using areal images.
 
-% \textit{Building height}.
-Building height is estimated using the building footprints and available lidar data following \cite{Chun2017-mm}.
-The mean lidar elevation within each building's footprint is subtracted by the mean digital elevation (topography) within the footprint to estimate the building's height.
+\textit{Elevation}. 
+1/3 arc-second (~10m) bare-earth elevation (topography) data are available courtesy the U.S. Geological Survey.
 
-\textit{Elevation}. 1/3 arc-second (~10m) bare-earth elevation (topography) data is available courtesy the U.S. Geological Survey.
+\textit{Surface elevation}. 
+The surface elevation is determined from the lidar data. 
+Surface elevation captures the natural and built features. 
+The lidar data was aggregated to calculate the surface elevation at a 6-meter resolution.
 
-\textit{Surface elevation}. The surface elevation is determined from the lidar data. Surface elevation captures the natural and built features.
-
-\textit{Impervious surface percentage}. This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}. 
+\textit{Impervious surface percentage}. 
+This is provided in the National Land Cover Database from the Multi-Resolution Land Characteristic Consortium \citep{Xian2011-aa}. 
 To generate the impervious surface area (ISA), the consortium uses LandSat data, NLCD land cover, and nighttime light imagery. 
@@
 The Landsat images were converted to top-of-atmosphere reflectance. 
-The data is provided at a 30m resolution.
+These data are provided at a 30m resolution.
 
 \textit{Land cover}. 
-Classification of land cover is described in \cite{Homer2015-ce}. 
-However, as it is used in the calculation of emissivity when calculating the LST it cannot be used as an independent variable. 
-The only landcover that we use is category 11, water.
+Classification of land cover is described in \citep{Homer2015-ce}. 
+However, as land cover is used when calculating the LST it cannot be used as an independent variable.
+That is, a variable used to construct a dependent variable should not be used as a predictor in a statistical model; 
+if it were, the statistical model will likely find the mathematical relationship used to create the response, rather than finding the biophysical association we are looking for. 
+Additionally, using the same variable in both construction and prediction will likely artificially inflate the stated accuracy of the statistical model due to overfitting.
+However, due to the significant biophysical difference between water and other land classifications, and given that the potential to overfit when only one of the categories is included is low, we do include water as an independent variable for the statistical analysis.
+These data are provided at a 30m resolution.
 
 \textit{NDBI}. 
-Normalized difference built-up index indicates the intensity of imperviousness \cite{Bhatti2014-ae}. 
-It is calculated from satellite images as $$NDBI=\frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}}$$ where $B_{NIR}$ and $B_{SWIR}$ are the reflectances in the near-infrared and short-wave infrared bands respectively \cite{Alhawiti2016-wv}.
-Using LandSat8 imagery, this is $$NDBI=\frac{B6-B5}{B6+B5}$$ \cite{barsi2014}.
-% *Kaplan 2018, Peng 2018
+Normalized difference built-up index indicates the intensity of imperviousness \citep{Bhatti2014-ae}. 
+It is calculated from satellite images as 
+\begin{equation}
+    \label{eqn:ndbi}
+    NDBI = \frac{B_{SWIR}-B_{NIR}}{B_{SWIR}+B_{NIR}} = \frac{B6-B5}{B6+B5}
+\end{equation}
+where $B_{NIR}=B_5$ and $B_{SWIR}=B_6$ are the reflectances in the near-infrared and short-wave infrared bands respectively \citep{Alhawiti2016-wv, barsi2014}.
+These data are at a 100-meter resolution and are calculated using the daytime satellite images.
 
@@
 The normalized difference vegetation index (NDVI) measures green vegetation. 
-It is calculated from satellite images as $$NDVI=\frac{B_{NIR}-B_{red}}{B_{NIR}+B_{red}}$$ where $B_{NIR}$ and $B_{red}$ are the reflectances in the near-infrared and red bands respectively \cite{Alhawiti2016-wv}.
-Using LandSat8 imagery, this is $$NDVI=\frac{B5-B4}{B5+B4}$$ \cite{barsi2014}.
-
-% \textit{Nighttime light intensity}.
-Stable nighttime light intensity is available from the Defense Meteorological Satellite Program (DMSP) Operational Linescan System (OLS).
-They prepare the data to remove clouds and ephemeral light sources.
-We use 2013 (the most recently available) data of version 4 at a spatial resolution of 30 arc-seconds (1km at the equator) \cite{ngdc2013version}.
-The nighttime light intensity was found to be only in moderate agreement with density \cite{bagan2018}, meaning lights may indicate anthropogenic energy use during the night.
-However, it's lower resolution means that can not capture minor differences.
-We resampled the raster to a 30m resolution, as per the NLDC \cite{Homer2015-ce}, and for each grid calculated the mean, min, and max.
+It is calculated from satellite images as 
+\begin{equation}
+    \label{eqn:ndvi}
+    NDVI = \frac{B_{NIR}-B_{red}}{B_{NIR}+B_{red}} = \frac{B5-B4}{B5+B4}
+\end{equation}
+where $B_{NIR} = B_5$ and $B_{red} = B_4$ are the reflectances in the near-infrared and red bands respectively \citep{Alhawiti2016-wv, barsi2014}.
+These data are at a 100-meter resolution and are calculated using the daytime satellite images.
 
 \textit{Population density}. 
-Some studies have found that population density has a positive influence on LST \cite{Li2017-yl, Peng2018-cp}, although this may be the result of confounding with other factors. 
+Some studies have found that population density has a positive influence on LST \citep{Li2017-yl, Peng2018-cp}, although this may be the result of confounding with other factors. 
 We calculate population density using the U.S. census at the block level. 
 The most recent census was 2010, so we use that data as an estimator of where people reside in the evening. 
-This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that its centroid is contained within.
+These data are also at a lower resolution than the grid cells used, so the grid cell assumes the density of the block that its centroid is contained within.
+These data are at the neighborhood block level.
 
 \textit{Sky view factor}. 
-Urban canyons have been found to have an effect on UHI because they prevent air circulation and hinder nighttime radiative cooling \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}.
-We calculate SVF using the R packages \textit{horizon} \cite{Van_doninck2018-ib} that is based off the algorithm presented in \cite{Dozier1990-kn}.
-We use the same parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. 
+Urban canyons, which consist of the air volume and surfaces bounded by walls of buildings separated by a street \citep{Oke2017-or}, control the airflow and energy exchange between the ground and roof \citep{Oke1976-zn} so affect the air temperature within the `urban canopy layer.'
+This can intensify surface temperatures \citep{Landsberg1981-mq, Chun2017-mm, Matzarakis2007-xy}. 
+We calculate SVF using the R package \textit{horizon} \citep{Van_doninck2018-ib} that is based off the algorithm presented in \citep{Dozier1990-kn}. 
+We use the same parameters used by \citep{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. 
 We use a spatial resolution of 6 meters.
@@
 \textit{Tree canopy cover}. 
-The 2011 edition of percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}.
-The data provided is at a 30m resolution. 
-The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance).
+The 2011 edition of percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \citep{Coulston2012-uu, Homer2015-ce}.
+The data provided are at a 30m resolution. 
+The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \citep{Coulston2012-uu} so the data do not contain information used in the LST calculation (which requires radiance).
 % * Rogan 2013, Elmes 2017
+These data are available at the 30m resolution.
 
-\subsection{Data preparation and robustness}
+\subsection{Data processing and variable selection}
+To analyze the data from these disparate sources, it needs to be transformed into a suitable format that is consistent between the variables.
+The raw data are a variety of spatial data types from ``area level" (the population density data) and ``geostatistical raster" (e.g. the land surface temperature).
+We chose to resample the data into a grid of square cells.
+We conduct the resampling twice to produce two datasets. The first has 100-meter square cells, the other has 500-meter cells. 
+Having two resolutions allows us to evaluate whether our conclusions are sensitive to the spatial resolution used.
+For each of the cells the mean, maximum, minimum, and standard deviation of all variables were calculated. 
+To address the spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting `spatial lag' variable was included as an additional independent variable.
 
-The objective of our study is to determine the most influential urban characteristics and understand how they relate to land surface temperature.
-To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
-For a selected characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature. 
-This is repeated over a range for the feature of interest.
-To evaluate how multiple urban characteristics interact, partial dependence can also be conducted in multiple dimensions.
-The swing \cite{Shortridge2015-ub} measures the relative importance of variables by calculating the magnitude of change in temperature associated with each variable.
-To calculate partial dependence, we need a data set and a statistical model.
-
-To analyze the data, it needs to be transformed into a suitable format that is consistent between the variables.
-The raw data is a variety of spatial data types from ``area level" (the population density data), to ``geostatistical raster" (e.g. the land surface temperature).
-We choose to resample the data into a grid of square cells.
-We conduct the resampling twice to produce a grid of 100 and 500-meter square cells.
-
-Having two resolutions allowed us to evaluate whether our conclusions are robust.
-For each of the cells the mean, maximum, and minimum of all variables were calculated. 
-To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable.
-To address multicollinearity between the variables that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than 5.
-The result is a data set with which we can train statistical models.
-
-Various statistical models (\S \ref{ss:models}) were trained and then validated on unseen data using a technique known as holdout cross-validation. 
-This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
-The model resulting from the training is then tested on this unseen test set (known as out-of-bag) to get an estimate of the model accuracy.
-This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
-Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
-Over fitting occurs when a model fits to the randomness in a dataset, causing it to be unsuitable for generalizations. 
-When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
-Avoiding overfitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies.
-Cross-validation helps to avoid this by evaluating the model on unseen data.
-
-A further way to avoid overfitting is to carefully consider how the test data is selected.
-This is especially important with spatial data.
-Therefore, when selecting the test and training data the grid cells were grouped into a larger 8x8 grid.
-This avoids overfitting of the model by training the model on a cell adjacent to a cell that is included in the test set.
-
-Finally, we attempt to capture uncertainty in the data and the models. 
-To represent the data uncertainty, we draw a random sample from our data, with replacement, to simulate a new dataset.
-The models are trained on this new dataset and this is repeated.
-This approach, known as bootstrapping, means that the conclusions' sensitivity to the data can be assessed.
-If the results are vastly different for each bootstrap sample, we assume that the result is dependent on the data and is unlikely to be generalizable to other cities.
-Additionally, we capture model uncertainty by using different models and assessing how each model shows the urban characteristics influencing the land surface temperature.
-If all bootstrapped models are generally in agreement, it suggests that the conclusions are robust.
+Once the dataset was created, we needed to select which variables would be used in the statistical models.
+With so many related variables there is a high chance of multicollinearity (where one variable can in turn be created from a combination of other variables).
+In practice this means that the information provided by one variable can also be provided by a combination of others.
+This leads to confounding in the analysis and the importance of each variable is reduced.
+To address multicollinearity between the variables that would confound our analysis of influence, we remove variables iteratively that have a `Variance Inflation Factor' greater than 5.
+The remaining variables for both 100 and 500-meter resolution analysis are shown in Table \ref{tab:vif}. 
+The result is a dataset without the potential for confounding variables, and one with which we can train the statistical models.
 
@@
 \textit{Null model: average}. The first model, to compare other models against, is the null model. 
-This is a benchmark model to ensure the models we fit are not doing worse than no model.
+This is a benchmark model to ensure the models we fit are not doing worse than no model \cite{Shortridge2015-ub, Logan2016-if}.
 The mean of the observations in the training set is calculated and is used as the prediction for the test observations.
@@
 \textit{Linear model}. Linear models are suitable when the relationship between the explanatory variables (urban characteristics) and the response variable (LST) is linear. 
-That is, linear models take the form $$ y = \beta_0 + \sum_i \beta_i x_i + \varepsilon$$ where $y$ is the response (LST), $x_i$ is each for the variables (urban characteristics), $\beta_0$ is the intersection, $\beta_i$ is how the LST changes linearly with each urban characteristic, and $\varepsilon$ is the normally distributed error.
+That is, linear models take the form
+\begin{equation}
+    \label{eqn:linear}
+    y = \beta_0 + \sum_i \beta_i x_i + \varepsilon
+\end{equation}
+where:
+\begin{conditions}
+    y & is the response (LST) \\
+    x_i &  is each of the variables (urban characteristics) \\   
+    \beta_i &  is how the LST changes linearly with each $x_i$, the urban characteristics \\   
+    \varepsilon & is the normally distributed error
+\end{conditions}
+\cite{Nelder1972-tl}.
 
 \textit{Multivariate Adaptive Regression Spline (MARS)}. 
-MARS models are a type of non-parametric regression approach that are useful for high dimensional problems \cite{Hastie2009-ky}. 
-They extend the linear model by using piecewise functions to fit the data \cite{Friedman1991-of}: $$ y = \sum_i c_i B_i(x)$$
+MARS models are a type of non-parametric regression approach that are useful for high dimensional problems \citep{Hastie2009-ky}. 
+They extend the linear model by using piecewise functions to fit the data \citep{Friedman1991-of}: 
+\begin{equation}
+    \label{eqn:mars}
+    y = \sum_i c_i B_i(x)
+\end{equation}
 $B_i(x)$ is the basis function that includes multiple linear functions and indicator variables that set these functions to zero for certain ranges of $x$.
-That is, multiple linear approximations are summed across the variable range and provide a nonlinear estimate of the LST \cite{Friedman1991-of}.
+That is, multiple linear approximations are summed across the variable range and provide a nonlinear estimate of the LST \citep{Friedman1991-of}.
 
 \textit{Generalized Additive Model (GAM)}. 
-The GAM is also an extension of the linear model but relaxes the assumption that the relationship is linear \cite{Hastie1990-cg}.
-Instead, the response variable is estimated as the sum of smoothing functions, splines, that are applied to each covariate or set of covariates: $$ y = \sum_i s_i(x_i)$$ where $s_i(x_i)$ is the smoothing function applied to each covariate \cite{Shortridge2015-ub}.
+The GAM is also an extension of the linear model but relaxes the assumption that the relationship is linear \citep{Hastie1990-cg}.
+Instead, the response variable is estimated as the sum of smoothing functions, splines, that are applied to each covariate or set of covariates:
+\begin{equation}
+    \label{eqn:gam}
+    y = \sum_i s_i(x_i)
+\end{equation}
+where $s_i(x_i)$ is the smoothing function applied to each covariate.
 
 \textit{Random Forest}. A random forest is an ensemble model of \textit{Regression Trees}. 
-A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}. 
+A regression tree partitions the data based on thresholds for the covariates \citep{Breiman1984-hw}. 
 The tree continues to `grow' by recursively partitioning the data with the objective of maximizing the node impurity, so that the partitions are as similar as possible. 
@@
 Each regression tree is trained on a randomly selected subset of 1/3 of the covariates. 
-The resulting prediction from the random forest is the unweighted mean of the prediction from all of the trees \cite{Breiman2001-rt}.
+The resulting prediction from the random forest is the unweighted mean of the prediction from all of the trees \citep{Breiman2001-rt}.
 A subset of the covariates is used to reduce the correlation between the predictions of each tree.
-Tree-based models do not assume linearity and so are generally very flexible and powerful models \cite{Breiman2001-rt, Geron2017-ek}.
+Tree-based models do not assume linearity and so are generally very flexible and powerful models \citep{Breiman2001-rt, Geron2017-ek}.
 
@@
 That is, each regression tree has the objective of reducing the error of the previous tree.
-The prediction from the GBRT is the unweighted mean of all of the regression trees \cite{Geron2017-ek}.
+The prediction from the GBRT is the unweighted mean of all of the regression trees \citep{Geron2017-ek}.
 In this case, we again use 500 trees.
@@
 Their use in image processing makes them potentially suitable for geographic studies.
-As with other forms of neural networks, CNN's contain layers of mathematical functions (neurons) that operate on the independent variables \cite{Geron2017-ek}.
+As with other forms of neural networks, CNN's contain layers of mathematical functions (neurons) that operate on the independent variables \citep{Geron2017-ek}.
 There can be multiple layers of these neurons that pass the results as an input to subsequent layers.
@@
 
+\subsection{Model validation}
+These statistical models were then trained and validated using a technique known as holdout cross-validation.
+This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
+The model resulting from the training is then tested on this unseen test set (known as out-of-bag) to get an estimate of the model accuracy.
+This is repeated 100 times and the distribution of the accuracy metrics are reported (here we use the mean absolute error (MAE) and variance explained (out-of-bag R$^2$)) \citep{Shmueli2011-wd}.
+Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \citep{Geron2017-ek}.
+Over fitting occurs when a model fits to the randomness in a data set, causing it to be unsuitable for generalizations. 
+When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
+Avoiding overfitting is therefore essential when working with statistical models and is overlooked by the majority of existing studies.
+Cross-validation helps to avoid this by evaluating the model on unseen data.
+
+Given we are dealing with spatial data, we must take additional steps to ensure that we avoid overfitting.
+Therefore, when selecting the test and training data the grid cells were grouped into a larger 8x8 grid.
+The training and test sets were selected from this 8x8 grid so that if the larger cell is included in the test set, so are all of the cells within.
+This avoids training and then testing the model on neighboring cells, which could lead to overfitting.
+
+\subsection{Variable influence and importance}
+The objective of our study is to determine the most influential urban characteristics and understand how they relate to land surface temperature.
+To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
+For a selected characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the data set and predicting the land surface temperature. 
+This is repeated over a range for the feature of interest.
+To evaluate how multiple urban characteristics interact, partial dependence can also be conducted in multiple dimensions.
+The swing \citep{Shortridge2015-ub} measures the relative importance of variables by calculating the magnitude of change in temperature associated with each variable.
+
+\subsection{Uncertainty quantification}
+Finally, we attempt to capture uncertainty in the data and the statistical models. 
+To represent the data uncertainty, we draw a random sample from our data, with replacement, to simulate a new data set.
+The models are trained on this new data set and this is repeated.
+This approach, known as bootstrapping, means that the conclusions' sensitivity to the data can be assessed.
+If the results are vastly different for each bootstrap sample, we assume that the result is dependent on the data and is unlikely to be generalizable to other cities.
+Additionally, we capture model uncertainty by using different models and assessing how each model shows the urban characteristics influencing the land surface temperature.
+If all bootstrapped models are generally in agreement, it suggests that the conclusions are robust.
+
 \section{Results and Discussion}
+\subsection{Model accuracy}
+In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
+To do that we trained various statistical models on the data where land surface temperature was the response ($y$ variable).
+The error metrics from the cross-validation are shown in Figures \ref{fig:holdout_100} and \ref{fig:holdout_500} for the 100 and 500-meter spatial resolutions respectively.
+Two error metrics were used; the left hand side shows the out-of-bag R$^2$ results, and the right hand side shows the mean absolute error.
+The results in Figure \ref{fig:holdout_100} show that the best models can predict both day and night land surface temperature to within 1$^o$C using the included urban characteristics.
+The R$^2$ results, also based on unseen data, suggest that more than 90\% of the data variance is captured by the best models.
+In all cases the random forest performs the best or is the close second.
+On the other hand, the linear model consistently under performs compared to the more complex models in the 100-meter resolution and is only better than the convolution neural network in the 500-meter models.
+The CNN struggled with the smaller dataset of the 500-meter resolution.
+However, the linear model is still explaining more than ~70\% of the variance in all cases.
+Given that this model remains the most commonly used in SUHI studies, it is important that future studies consider the importance of high accuracy relative to the objective of their analysis when justifying the use of the linear model.
 
@@
 
-In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
-To achieve this, we first select variables to remove based on collinearity. 
-The remaining variables for both 100 and 500-meter resolution analysis are shown in \ref{tab:vif}. 
-The next phase is to assess the accuracy of the statistical models.
-The results in Figure \ref{fig:holdout_100} show that the best models can predict both day and night land surface temperature to within 1$^o$C using urban characteristics.
-The R$^2$ results, also based on unseen data, suggest that more than 90\% of the data variance is captured by the best models.
-We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees and convolutional neural network.
-The weakest model, consistently with R$^2$ less than 80\%, is the linear regression - incidentally the model that is used in the majority of existing studies into land surface temperature.
-These results allow us to assess the influence and relative importance that these characteristics.
+\subsection{Variable importance and influence}
+Our objective with this study is to determine the influence and importance of various urban surface characteristics on land surface temperatures.
+In this section we present results showing the importance and influence of the variables used in the statistical models.
+However, in statistical modelling, one challenge to understanding influence and importance is decoupling variables that are often interrelated.
+Failing to do this can result in confounding variables.
+In Section 2.4, we describe how we select variables to reduce this by removing those that are correlated and multicollinear.
+In doing so, we found that in this dataset tree canopy is 100\% negatively correlated with impervious surfaces (Figure \ref{fig:imp_tree}).
+Therefore, we do not include \% impervious surface as a variable in the models, but it means that the terms tree canopy and pervious surface are equivalent when used in relation to the variables in the statistical models.
 
@@
     \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/importance_100.png}
-    \caption[Variable influence on LST at 100-meter resolution]{
-    Variable influence on LST at 100-meter resolution.
-    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
-    \label{fig:importance_100}
+    \includegraphics[width=\linewidth]{fig/report/importance_stacked.png}
+    \caption[The relative importance of variables on land surface temperature]{
+    The relative importance of variables on land surface temperature. 
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature for each model.
+These results are then weighted by the each model's R$^2$ value and stacked to show which variables, across all of the models, have the largest influence on land surface temperature.
+
+    Both 100 and 500 meter resolutions are shown to demonstrate the agreement and disparity across the spatial scales.
+    }
+    \label{fig:importance_stack}
     \end{center}
@@
     There are multiple lines for each model based on bootstrap samples of the data, which indicates the data uncertainty.
-    The histograms on the $x$-axis shown the distribution of the observed data.
+    The histograms on the $x$-axis show the distribution of the observed data.
     This is for the 500-meter resolution.
@@
 
-Relative variable importance for all the urban characteristics, calculated with \textit{swing}, is shown in Figure \ref{fig:importance_100}. 
-The characteristics, shown on the $y$ axis, are ordered by the mean swing across all of the models, and the models are ordered left-to-right by their cross-validated accuracy.
-The most important characteristics during the night are tree canopy cover, greenness (NDVI), the sky view factor, and the elevation (digital surface model).
-The albedo is found to be important when assessing the 100-meter data, while the \% area covered in water is important with the 500-meter data.
-This is likely due to 98\% of the cells at 100m resolution having 0\% water, so the effect on the dataset is negligible.
+Figure \ref{fig:importance_stack} presents the relative importance for each of the urban characteristics included in the models.
+This stacked bar graph shows the cumulative importance of each variable to each statistical model, and is weighted by the R$^2$ of each statistical model.
+The unweighted results are shown in Figures \ref{fig:importance_100} and \ref{fig:importance_500}.
+The most important characteristics during the night are tree canopy cover and greenness (NDVI).
+The sky view factor and the elevation (digital surface model) also appear important. 
+The albedo is found to be important when assessing the 100-meter data, while the \% area covered in water is important with the 500-meter data; 
+This is likely due to 98\% of the cells at 100m resolution having 0\% water, so the effect on the data set is negligible.
 The important urban characteristics are consistent for both the mean and maximum temperature during the night.
+
 The characteristics are also relatively consistent between night and day and between the models. 
-Although albedo is among the top factors, its relative importance is low compared to tree canopy and greenness, especially during the night. 
-This low association with nighttime temperatures is surprising given that darker surfaces store heat during the day that is released during the night \cite{Voogt2003-mm, Zhou2014-wc}.
-This, and the other influences can be examined using partial dependence.
+Tree canopy and NDVI remain among the most important factors, however in the 500-meter dataset water is found to be the most important.
+The importance of albedo is much higher for these daytime models than those using nighttime LST.
+
+The minor association between albedo and nighttime temperatures is surprising given that darker surfaces store heat during the day that is released during the night \citep{Voogt2003-mm, Zhou2014-wc}.
+However, this may be due to the nature of the remotely sensed data which captures the `seen' surface of the urban environment \citep{Roth1989-jj}; 
+this `seen' surface is what is visible from a bird's eye view of the city, including roof-tops, tree-tops, roads, and open flat areas \citep{Roth1989-jj}.
+These areas are likely more exposed to air flow and may store less heat than other surfaces within the city.
+The relatively small reported importance of the sky view factor may also be due to the data sampled by satellites.
+These data will likely under-sample vertical surfaces and those under canopies \citep{Roth1989-jj, Arnfield2003-gn} and so this must be considered when analyzing the results \citep{Roth1989-jj}.
+
+Variable influence is shown in Figures \ref{fig:pdp_100}-\ref{fig:pdp_2dday_100} and \ref{fig:pdp_500}-\ref{fig:pdp_2dday_500}. 
+These are the partial dependence plots and show how the surface temperature changes, on average, with the change to that variable.
+Figures \ref{fig:pdp_100} and \ref{fig:pdp_500} show how LST changes with changes to one covariate.
+Figures \ref{fig:pdp_2dnight_100}, \ref{fig:pdp_2dday_100}, \ref{fig:pdp_2dnight_500}, and \ref{fig:pdp_2dday_500} show the changes in LST as two variables change.
 
 \textit{Vegetation and impervious surfaces}. 
-During the day, increasing impervious surfaces and decreasing vegetation causes increased sensible heat flux and lowered latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}.
-This is thought to be less important during the night because latent and sensible heat are dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
+During the day, increasing impervious surfaces and decreasing vegetation will increase sensible heat flux and lower latent heat flux \citep{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}.
+During the night, this process is thought to be less important because while latent and sensible heat are dominant during the day, ground heat flux dominates at night \citep{Zhou2014-wc, Voogt2003-mm}.
 Indeed, because transpiration does not occur at night, vegetation's effect on night temperatures is debated. 
-In our study, percentage tree canopy and impervious surface cover must be discussed together because they are 100\% correlated in the data (Figure \ref{fig:imp_tree}).
-To distinguish between vegetated surfaces and other pervious surfaces, we include NDVI, a measure of a surface's greenness.
-At night, we find that \% tree canopy (and \% impervious surface) is the most influential characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:importance_100}).
-The partial dependence results (Figure \ref{fig:pdp_100}) indicate a decrease in up to 10$^oC$ as the percentage area of trees or other pervious surface increases.
-This reduction in temperature could be due to a reduction in the impervious surfaces that store heat.
-We also observe that as the NDVI increases, the land surface temperature decreases.
+Our variable importance plots show that \% tree canopy (recall that in this study this is equivalent to \% pervious surface) is the most important factor on surface temperature during the night.
+This is corroborated by the partial dependence plots.
+Figure \ref{fig:pdp_100} shows that as tree canopy increases, the surface temperature decreases (a decrease in up to 10$^oC$ as the percentage area of trees or other pervious surface increases).
+There is strong agreement between the statistical models, although the nonlinear models indicate a sharp decline in the average (mean) temperature with an initial tree canopy cover and suggest that this effect (the lines' gradients) weakens with a higher tree canopy cover.
+This weakening may also be related to the sparser data at the upper tail of the dataset, and this is indicated by the wider uncertainty.
 
-To further evaluate whether the influence with temperature is due to vegetation or pervious surfaces, consider the the two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_100}d) that evaluates how LST varies with \% tree canopy/impervious surface and the NDVI.
+NDVI, a measure of the surface's greenness, was reported to be the next most important factor on land surface temperature. 
+Given the conflation between tree canopy cover and impervious surfaces, NDVI can help to distinguish between vegetated surfaces and other pervious surfaces.
+The relationship between NDVI and LST is surprisingly linear (Figure \ref{fig:pdp_100}), which clearly indicates that the greener a surface, the cooler it is.
+
+The two-dimensional partial dependence plots (Figure \ref{fig:pdp_2dnight_100}) enable us to further evaluate this relationship.
+By considering the joint effect of NDVI and tree canopy we can infer whether the influence on temperature is due to vegetation or simply pervious surfaces. 
 The greener (higher NDVI) and more pervious (higher \% tree canopy cover) a surface is, the cooler it is during the night. 
-The total change here is also approximately 7.5$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less sensitive of the five models (as seen in Figure \ref{fig:pdp_100}).
-These results contradict standing conclusions that vegetation has no effect on nighttime LST \cite{Peng2012-iy, Zhou2014-wc};
+The total change here is also approximately 7.5$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the least sensitive of the five models (as seen in Figure \ref{fig:pdp_100}).
+These results do not support existing conclusions that vegetation has no effect on nighttime LST \citep{Peng2012-iy, Zhou2014-wc};
 although much of the reduction in temperature could be due to the perviousness of the surface, lowered temperatures appear to be associated with the greenness (NDVI) of a surface during the night.
@@
 \% tree canopy).
-This daytime result is consistent with expectation \cite{Chun2017-mm, Peng2018-cp, Wang2019-tree,Zhou2014-wc, Chun2018-so}.
-Because not all pervious surface types have the same cooling potential, and because of trade-offs with irrigation of vegetated surfaces \cite{Gober2009-im}, understanding the ideal vegetated surface type for day and night temperatures is a future step.
+This daytime result is consistent with expectation \citep{Chun2017-mm, Peng2018-cp, Wang2019-tree,Zhou2014-wc, Chun2018-so}.
+
+An ongoing research step is determining the ideal vegetated and pervious surface.
+This is because there is significant diversity in the cooling effectiveness of different types of green coverage \citep{Saaroni2018-ct}. 
+For example, the effect of street trees can be highly localized \citep{Coutts2016-og}, while some types of trees cope with heat better than others \citep{Leuzinger2010-si}. 
+Additionally, there are trade-offs with irrigation on vegetated surfaces \citep{Gober2009-im}.
+There is also likely significant differences between tree canopy vs ground-level vegetation that cannot easily be discerned from remotely sensed studies given the data limitations.
+However, these results show the importance of vegetated and pervious surfaces at a larger scale, which should motivate further investigation into the finer urban scales.
+Understanding these differences will guide planning measures in urban spaces.
 
 \textit{Water}.
-Water is widely expected to decrease the LST during the day \cite{Wicki2017-fv, Zhou2018-iy, Wang2019-water}, but some claim it increases LST during the night \cite{Chun2017-mm}.
-The rationale is that water releases heat during the night, resulting in elevated nighttime temperature \cite{Chun2017-mm}.
-Our findings, however, show that water reduces the LST during both the day and night (Figure \ref{fig:pdp_500}).
+Water is widely expected to decrease the LST during the day \citep{Wicki2017-fv, Zhou2018-iy, Wang2019-water}, but some claim that it may also increase LST during the night \citep{Chun2017-mm}.
+The rationale is that water releases heat during the night, resulting in elevated nighttime temperature \citep{Chun2017-mm}.
+This behavior may change depending on the water body’s characteristics, the season, and location.
+Our findings, however, show that, during the summer months, water reduces the LST during both the day and night (Figure \ref{fig:pdp_500}).
 Although these results are not supported at the 100-meter resolution because 98\% of the data at the 100-meter level has zero percentage water.
 At the 500-meter resolution, we see that the presence of water can decrease LST between 1.5 and 8$^oC$ during the night and by substantially more during the day (Figure \ref{fig:pdp_500}).
+At the 500m scale this is likely referring to larger water bodies, however the depth and other water characteristics are not evaluated.
 
 \textit{Urbanization}.
-Urbanization can lead to heat storage in roads and buildings \cite{Zhou2014-wc, Voogt2003-mm}. 
-We discussed the role of impervious surfaces, alongside the effect of vegetation, but we also considered the influence of albedo (the whiteness of a surface), the percentage area of building, the NDBI (built-up index), and the sky view factor (a measure of the urban canyon effect).
-The results for albedo are surprisingly low.
-During both the night and day, albedo has a negative effect on temperature.
-During the day, this is most pronounced, and whiter surfaces appear to be more than 5$^o$C cooler than darker ones (Figure \ref{fig:pdp_100}).
-The partial dependence plot (Figure \ref{fig:pdp_100}) also indicates that albedo's influence on temperature is potentially confounded as both vegetation and impervious surfaces can be dark surfaces (with low albedo).
-This stresses the importance of statistical techniques for evaluating the effect of multiple variables.
-We can investigate this further using the 2D partial dependence; the highest temperature occurs when there are dark impervious surfaces, but dark vegetated surfaces are cool (Figure \ref{fig:pdp_2dnight_100}).
+Urbanization can lead to heat storage in roads and buildings \citep{Zhou2014-wc, Voogt2003-mm}. 
+We discussed the role of impervious surfaces, alongside the effect of vegetation, but the statistical models also analyzed the influence of albedo, the percentage area of building, the NDBI (built-up index), and the sky view factor (a measure of the urban canyon effect).
+The results for albedo are surprisingly low (Figure \ref{fig:importance_stack}): 
+At night it is vastly less important relative to the vegetation and pervious factors, during the day it has more of an influence.
+At both the night and day times sampled, albedo has a negative effect on temperature (Figure \ref{fig:pdp_100}).
+During the day, this is most pronounced, and more reflective surfaces appear to be more than 5$^o$C cooler than others.
+There is the potential that the effects of albedo on temperature are potentially confounded as both vegetation and impervious surfaces can have low albedo.
+This highlights the utility of statistical models that can evaluate the effect of multiple variables.
+We can investigate this further using the 2D partial dependence (Figures \ref{fig:pdp_2dnight_100} and \ref{fig:pdp_2dday_100}). 
+Consider the plot showing the \% tree canopy vs albedo: the highest temperature occurs when there are dark impervious surfaces (bottom left).
+While increasing the tree canopy (or, in this case, perviousness of the surface) will reduce temperature, we also see that the lowest temperature is with high ``\% tree canopy'' and high albedo.
 Additionally, we see that albedo has little affect at night, although the highest temperature does occur when there is high acreage of building footprint with low albedo (Figure \ref{fig:pdp_2dnight_100}a).
-During the day, albedo has a greater influence: increasing the whiteness decreases the temperature (Figure \ref{fig:pdp_100}).
-The highest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_100}a), as well as high impervious area and low albedo (Figure \ref{fig:pdp_2dday_100}b).
-This supports existing findings that albedo decreases LST during the day, however, that we find no strong effect during the night contrasts existing reports \cite{Peng2012-iy, Zhou2014-wc}.
+%  increasing the reflectance decreases the temperature (Figure \ref{fig:pdp_100}).
+During the day, albedo has a greater influence: the highest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_100}a), as well as high impervious area and low albedo (Figure \ref{fig:pdp_2dday_100}b).
+Our results show that albedo decreases LST during the day, as expected \cite{Oke2017-or}.
+However, in contrast to existing studies \citep{Peng2012-iy, Zhou2014-wc}, we find no strong association between albedo and surface temperature during the night.
 
@@
 The building area had no effect during the night and minorly increased temperature during the day (Figure \ref{fig:importance_100}).
-Maximum NDBI appears to have a slight positive associated with maximum daytime LST (Figure \ref{fig:pdp_500}), but this effect is not observed at the 100m resolution.
-This minor-to-negligible relationship is surprising given that NDBI has been reported as among the most important urban characteristics during the day \cite{Peng2018-cp}.
+Maximum NDBI appears to have a slight positive association with maximum daytime LST (Figure \ref{fig:pdp_500}), but this effect is not observed at the 100m resolution.
+This minor-to-negligible relationship is surprising given that NDBI has been reported as among the most important urban characteristics during the day \citep{Peng2018-cp}.
 However, the discrepancy may be due to \% impervious surface area being incorporated already with the \% tree canopy data.
 
-The canyon effect is also often attributed with causing warmer temperatures \cite{Chun2017-mm,Oke1988-re}.
+The canyon effect is also often attributed with causing warmer temperatures \citep{Chun2017-mm}, in particular, warmer air temperatures due to the urban canopy layer \cite{Oke1988-re}.
 While there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_100}a,b).
-This is potentially due to heat being stored within the canyons (areas with low sky view factors).
-This is supported in the two-dimensional partial dependence plot (Figure \ref{fig:pdp_2dnight_100}) where the higher temperatures are observed when there is high \% building area and low sky view factor.
+This is due to heat being captured within the canyons (areas with low sky view factors) due to changes in the air flow and energy exchange \cite{Oke1982-px} and, conversely, more exposed areas with high sky view factors will release their heat.
+This is shown in the two-dimensional partial dependence plot (Figure \ref{fig:pdp_2dnight_100}e) where the higher temperatures are observed when there is high \% building area and low sky view factor.
 It follows that heat is being captured in the canyons. 
 However, compared to the other urban characteristics this has a lesser affect, changing the temperature by approximately 1.5$^o$C.
-The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_100} and \ref{fig:pdp_2dday_100}), but suggests that there are higher temperatures when the sky view factor is high.
+The effect of the sky view factor during the day is low-to-indiscernible (Figures \ref{fig:importance_100} and \ref{fig:pdp_2dday_100}f), but suggests that there are higher temperatures when the sky view factor is high.
+Given that the relationship between urban canyons and urban temperatures is well-studied and known to exist \cite{Arnfield2003-gn}, the indiscernible relationship found here may reflect the limitations of remotely sensed data to investigate this effect.
+This is due to the satellite imagery over sampling surfaces that are `seen' and therefore will include a lower proportion of areas that are within canyons that are obstructed by the structures that create the canyon \cite{Roth1989-jj}.
 
@@
 
-\textit{Result sensitivity}.
+\subsection{Result sensitivity}.
 To assess the robustness of the results, we conduct the analysis again at the 500-meter resolution (\ref{ss:500_meter}).
@@
 
-\textit{Urban strategies}.
-
 \begin{figure*}
@@
     \caption{
-    \textbf{Nighttime, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
+    \textbf{Nighttime, mean}: A two-dimension partial dependence contour plot showing how the land surface temperature ($^oC$) at 100-meter resolution changes with the two variables shown on the $x$ and $y$ axes, while the remaining variables are held at their average value.
+    This figure is the two-dimensional equivalent of Figure \ref{fig:pdp_100}, except with only the random forest model and without the uncertainty shown.
     }
@@
     \caption{
-    \textbf{Daytime, mean}: A two-dimension partial dependence plot showing how the land surface temperature ($^oC$, contours) changes the variables on the $x$ and $y$ axes, while the remaining variables are unchanged.
+    \textbf{Daytime, mean}: A two-dimension partial dependence contour plot showing how the land surface temperature ($^oC$) at 100-meter resolution changes with the pair of variables shown on the $x$ and $y$ axes, while the remaining variables are held at their average value.
     }
@@
 To assist planners tackling climate change, heat waves, and general high urban temperatures, we seek to determine how different urban characteristics are associated with land surface temperatures. 
-The results strongly support initiatives for increasing green infrastructure in cities \cite{Larsen2015-da, Meerow2017-xv}. 
-We found that increasing vegetation and reducing impervious surface area had the greatest effect on temperature during both night and day, with the potential to reduce the temperature up to 10$^o$C.
+The results strongly support initiatives for increasing green infrastructure in cities (e.g., \citep{Larsen2015-da, Meerow2017-xv}). 
+We found that increasing vegetation and reducing impervious surface area had the greatest effect on surface temperature during both night and day, with the potential to reduce the temperature up to 10$^o$C.
+While it is not surprising that vegetation and imperviousness are related to surface temperatures, this study is the first to quantify them for Landsat-resolution nighttime temperature estimates. 
+These results are also useful for corroborating this relationship, because in complex urban environments it is far from clear that the anticipated behaviour would eventuate; for example, in the case of urban vegetation at night, much of the urban vegetation exists in water stressed environments that limit evaporative cooling, and so the mechanics and magnitude of the vegetation's effect on nighttime temperature is far from clear.
 We also see strong evidence for blue space and increasing the area of water, although there are various limitations with doing so in cities such as Phoenix.
 We do note that this analysis is for land surface temperature, and not air temperature. 
-The correlation between LST and air temperature is not perfect at intra-urban scales, and the magnitude of LST variability tends to be larger  than that of air temperature.
-Nevertheless, LST is a highly relevant variable for analyses of the urban energy balance, and thus for studies of the UHI process.
+The correlation between LST and air temperature is varied at intra-urban scales, and the magnitude of LST variability tends to be larger than that of air temperature.
+Also, remotely sensed data has limitations in that it cannot capture a representative sample of the urban environment, instead it favors surfaces seen from a bird's eye view.
+Nevertheless, remotely sensed LST is a highly relevant variable for analyses of the urban energy balance, and thus for studies of the urban heat island and ways to ameliorate it.
 
@@
 This result opens opportunities for further detailed analysis into potential interventions.
-Such interventions for mitigating high temperatures are naturally place-specific, and while these results have proven general to four different US cities, work is needed to understand questions such as which types of greenery are better than others \cite{Gober2009-im}.
+Such interventions for mitigating high temperatures are naturally place-specific, and while these results have proven general to four different US cities, work is needed to understand questions such as which types of greenery are better than others \citep{Gober2009-im}.
+
 This study also demonstrates the opportunities of modern statistical techniques and the ability to assess potentially nonlinear interactions and the interactions between multiple variables. 
+While the linear model did not perform as well as the more complex models, it was largely in agreement with the other models regarding the influence of the urban characteristics on land surface temperature.
+The major limitation with linear models in this analysis is therefore the assumption that the relationship between surface temperature and each of the variables is constant (i.e., it doesn't change with the size of the covariate or dependent on any thresholds or other factors).
+Another advantage of the advanced statistical models is their ability to analyze the interaction between factors, which is far less nuanced when using a linear model.
+For example, in this analysis we were able to examine and disentangle the influence of two, often covarying, variables such as vegetation and albedo, and vegetation and NDVI.
+These limitations may be perfectly suitable for many analyses, but are necessary to be aware of and considered when more detailed analysis is required.
 
 Rigorous statistical analysis can continue to answer on-going questions central to land surface temperature.
-For example, our results support the suggestion that 3D variables (e.g. sky view factor) do not outperform 2D ones (e.g. \% tree canopy cover) \cite{Berger2017-lx}. 
-However, iteratively removing these variables from the analysis can ameliorate the potential for conflating the importance of different characteristics \cite{Chun2017-mm}.
-Our analysis is not causal inference; such an analysis would require controlled variables and urban spaces.
+For example, our results support the suggestion that 3D variables (e.g. sky view factor) do not outperform 2D ones (e.g. \% tree canopy cover) \citep{Berger2017-lx}. 
+However, due to sampling limitations of the remotely sensed data, we do not believe the 3D variables are unimportant.
+Additionally, we find some discrepancies between the analyses at the two spatial scales (notably in the effect and importance of water); this sensitivity is something that further studies should be aware of.
+Our analysis is also not causal inference; such an analysis would require controlled variables and urban spaces.
 We instead aim to complement and inform existing understanding of the relationships by demonstrating how land surface temperature and the urban characteristics are associated.
@@
 However, if no association exists or is in the opposite direction of proposed strategies for heat reduction, then such strategies need to be re-evaluated.
-For example, our results assessing the relative effect of sky view factor (a measure of building density), population, built-up index, and \% building area do not support claims advocating to constrain floor area ratios \cite{Chun2017-mm}.
-Such policy could lead to increased sprawl.
+For example, our results assessing the relative effect of sky view factor (a measure of building density), population, built-up index, and \% building area do not support claims advocating to constrain floor area ratios \citep{Chun2017-mm}.
+Such policy could lead to increased urban sprawl (low-density or single-use urban development characterized by poor accessibility and a lack of functional open-space \citep{Ewing2015-xj}).
 Instead, our results show that reducing impervious area and increasing vegetation and greenness have a stronger association with lowered temperatures.
@@
 If a constraint is explored, an additional necessary study is to compare sprawling and dense cities, prior to taking steps to dissuade density. 
-Now that 100-meter resolution nighttime data is available, as well as the increasing availability of lidar data, studies advocating creative interventions such as increasing the vertical and horizontal randomness of buildings, increasing the prevalence of green roofs \cite{Gago2013-ta}, or other urban planning strategies can quantitatively be explored.
+Now that 100-meter resolution nighttime data are available, as well as the increasing availability of lidar data, studies advocating creative interventions such as increasing the vertical and horizontal randomness of buildings, increasing the prevalence of green roofs \citep{Gago2013-ta, Kelbaugh2019-th}, or other urban planning strategies can quantitatively be explored.
 
@@
 
-% Here are two sample references: \cite{Feynman1963118,Dirac1953888}.
+% Here are two sample references: \citep{Feynman1963118,Dirac1953888}.
 \section*{Acknowledgements}
-% TML conducted this research as a graduate student at the University of Michigan with the support of a Rackham PreDoctoral Fellowship.
+% TML conducted the majority of this research as a graduate student at the University of Michigan with the support of a Rackham PreDoctoral Fellowship.
 % The work was also supported by the US National Science Foundation's Grant SEES-1631409. 
@@
 The data sources for the analysis are listed in Table \ref{tab:data}.
+The satellite imagery used is described in Table \ref{tab:satellite_meta}.
+All of the data was gridded as shown in Figure \ref{fig:map}.
 
@@
 \fontsize{8}{11}\selectfont
-\begin{tabu}to \textwidth{ X[l]  X[c]  X[c] X[l] X[l] }
+\begin{tabu}to \textwidth{ X[l]  X[c]  X[c] X[c] X[l] X[l] }
 \toprule
- Data Provider & Data Type & Data Date & Description & Source \\
+ Data Provider & Data Type & Data Date & Resolution & Description & Source \\
  \hline
-U.S. Geological Survey  & Raster  & 2013-2017 &
+U.S. Geological Survey  & Raster  & 2013-2017 & 100m &
     Landsat 8 day and night satellite imagery & \url{https://earthexplorer.usgs.gov/} \\
-Microsoft  & Polygon  & 2018 & Building footprint polygons for the US &
+Microsoft  & Polygon  & 2018 & n/a & Building footprint polygons for the US &
     \url{https://github.com/Microsoft/USBuildingFootprints} \\
-Defense Meteorological Satellite Program  & Raster  & 2013 &
-    Stable nighttime light intensity & \url{https://www.ngdc.noaa.gov/eog/dmsp/downloadV4composites.html} \\
-Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
+% Defense Meteorological Satellite Program  & Raster  & 2013 &
+    % Stable nighttime light intensity & \url{https://www.ngdc.noaa.gov/eog/dmsp/downloadV4composites.html} \\
+Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 & 30m &
     Land cover & \url{https://www.mrlc.gov} \\
-Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
+Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 & 30m &
     Percent developed imperviousness & \url{https://viewer.nationalmap.gov} \\
-Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 &
+Multi-Resolution Land Characteristics Consortium  & Raster  & 2011 & 30m &
     Percent tree canopy cover & \url{https://viewer.nationalmap.gov} \\
-U.S. Geological Survey  & Raster  & 2015 & 1/3 arc-second elevation &
+U.S. Geological Survey  & Raster  & 2015 & 10m & 1/3 arc-second elevation & 
     \url{https://nationalmap.gov/3DEP/3dep_prodserv.html} \\
-U.S. National Oceanic and Atmospheric Administration  & Lidar  & 2014 & Point cloud of surface elevation &
+U.S. National Oceanic and Atmospheric Administration  & Lidar  & 2014 & 1m (Detroit) - resolved to 6m & Point cloud of surface elevation &
     \url{https://coast.noaa.gov/htdata/lidar2_z/geoid12b/data/6377/} \\
-IPUMS NHGIS  & Area Level  & 2010 &
-    Block-level population from the US census & \cite{nhgis}\\
+IPUMS NHGIS  & Area Level  & 2010 & city block &
+    Block-level population from the US census & \citep{nhgis}\\
 \bottomrule
@@
 \end{table}
+
+\begin{table}[H]
+\caption{The date, time, and cloud cover information of the satellite images used for the remote sensing data. The night time images do not have percent cloud cover calculated in their metadata.}
+\label{tab:satellite_meta}
+\begin{tabular}{ll|lll}
+          &       & date      & time  & \% cloud cover \\
+          \hline
+Baltimore & Day   & 07-Sep-17 & 11:46 & 11.6\%         \\
+          &       & 22-Aug-17 & 11:46 & 0.7\%          \\
+          &       & 18-May-17 & 11:45 & 20.3\%         \\
+          &       & 18-Jul-16 & 11:46 & 11.5\%         \\
+          &       & 02-Jul-16 & 11:46 & 1.4\%          \\
+          & Night & 08-Sep-17 & 22:50 & -              \\
+          &       & 05-Sep-16 & 22:50 & -              \\
+          &       & 20-Aug-16 & 22:50 & -              \\
+          &       & 04-Aug-16 & 22:50 & -              \\
+Detroit   & Day   & 26-Sep-17 & 12:16 & 0.4\%          \\
+          &       & 03-Jun-16 & 12:15 & 6.9\%          \\
+          &       & 18-May-16 & 12:15 & 1.9\%          \\
+          & Night & 04-Sep-17 & 23:16 & -              \\
+          &       & 19-Aug-17 & 23:16 & -              \\
+          &       & 18-Jul-17 & 23:15 & -              \\
+          &       & 31-May-17 & 23:15 & -              \\
+          &       & 15-May-17 & 23:15 & -              \\
+Phoenix   & Day   & 01-Sep-17 & 11:03 & 0.0\%          \\
+          &       & 16-Aug-17 & 11:03 & 0.0\%          \\
+          &       & 15-Jul-17 & 11:03 & 2.2\%          \\
+          &       & 29-Jun-17 & 11:03 & 0.0\%          \\
+          &       & 13-Jun-17 & 11:03 & 0.0\%          \\
+          & Night & 16-Sep-17 & 22:17 & -              \\
+          &       & 31-Aug-17 & 22:17 & -              \\
+          &       & 15-Aug-17 & 22:17 & -              \\
+          &       & 28-Jun-17 & 22:17 & -              \\
+          &       & 12-Jun-17 & 22:16 & -              \\
+Portland  & Day   & 30-Jul-17 & 11:55 & 2.9\%          \\
+          &       & 14-Jul-17 & 11:55 & 0.2\%          \\
+          &       & 27-May-17 & 11:55 & 1.8\%          \\
+          &       & 13-Sep-16 & 11:56 & 0.0\%          \\
+          &       & 12-Aug-16 &       & 0.1\%          \\
+          & Night & 28-Sep-17 & 22:45 & -              \\
+          &       & 12-Sep-17 & 22:45 & -              \\
+          &       & 27-Aug-17 & 22:45 & -              \\
+          &       & 24-Jun-17 & 22:45 & -              \\
+          &       & 23-May-17 & 22:45 & -             
+\end{tabular}
+\end{table}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{fig/report/map_nocturnal_lst.pdf}
+    \caption[The nighttime land surface temperature in $^o$C, gridded into 500-meter cells]{The nighttime land surface temperature in $^o$C, gridded into 500-meter cells, for the cities studied.
+This is change from the mean (average) for each city.}
+    \label{fig:map}
+\end{figure*}
+
+\newpage
+\section{Code for the land surface temperature calculation}
+\label{appendix:code}
+
+This is an extract of the code provided in the GitHub repository: []
+
+\begin{lstlisting}[language=Python]
+
+def calc_LST(info_satellite, meta_dict, source_city):
+    '''
+        Returns the land surface temperature (LST) based on satellite imagery
+    '''
+    # read in band 10 data
+    image_b10 = gdal.Open(fn_b10)
+    dn = image_b10.ReadAsArray()
+
+    # conversion to TOA radiance
+    TOA = calc_TOA(dn, meta_dict, 10)
+
+    # emissivity correction
+    emissivity = determine_emissivity(info_satellite, dn, source_city)
+
+    # calculate the at-satellite brightness temperature
+    temp_satellite = calc_satellite_temperature(TOA, meta_dict, emissivity)
+
+    # atmospheric correction
+    temp_surface = atmos_correction(temp_satellite, info_satellite, emissivity)
+
+def calc_TOA(dn, meta_dict, band_number):
+    '''
+        Calculate the Top Atmosphere Spectral Radiance (TOAr) from Band 10 digital
+        number (DN) data (also refered to as the Q_cal - quantized and
+        calibrated standard product pixel value)
+    '''
+
+    TOAr = meta_dict['RADIANCE_MULT_BAND_{}'.format(band_number)] * dn + meta_dict['RADIANCE_ADD_BAND_{}'.format(band_number)]
+
+    return(TOAr)
+
+def calc_satellite_temperature(TOA, meta_dict, emissivity):
+    '''
+        Calculate the At-Satellite Brightness temperature
+        First, need to calculate the emissivity from the land use
+        Then convert
+        Returns temperature in Kelvin
+    '''
+    # spectral radiance
+    L_lam = TOA/emissivity
+
+    # calculate the satellite brightness
+    temp_satellite = meta_dict['K2_CONSTANT_BAND_10']/(np.log(1 + (meta_dict['K1_CONSTANT_BAND_10']/L_lam)))
+
+    return temp_satellite
+
+def atmos_correction(temp_satellite, info_satellite, emissivity):
+    '''
+        Using the mono-window algorithm (Qin et al., 2001, International Journal of Remote Sensing)
+        make the atmospheric correction
+        Returns land surface temperature in celsius
+    '''
+
+    # temparature from csv file
+    temp_max = info_satellite['max_temp_celsius']
+    # convert to Kelvin
+    temp_max += 273.15
+
+    # constants for the algorithm
+    a_6 = -67.355351
+    b_6 = 0.458606
+    w = 1.6
+    t_6 = 0.974290 - 0.08007*w
+    # variables dependent on land cover (emissivity) and temperature
+    c_6 = emissivity * t_6
+    d_6 = (1 - t_6)*(1 + (1 - emissivity)*t_6)
+    t_a = 16.0110 + 0.92621*temp_max
+
+    # mono-window algorithm (Qin et al., 2001, page 3726)
+    T = a_6*(1 - c_6 - d_6) + (b_6*(1 - c_6 - d_6) + c_6 + d_6)*temp_satellite - d_6*t_a
+    temp_landsurface = T/c_6
+
+    # converting to celsius
+    temp_landsurface -= 273.15
+
+    return temp_landsurface
+\end{lstlisting}
 
@@
 \newpage
-\section{Technical appendix: convolutional neural network}
+\section{Technical appendix: Convolutional neural network}
 \label{ss:cnn}
@@
 
-Our CNN model was adapted from a U-Net architecture \cite{unet}.
+Our CNN model was adapted from a U-Net architecture \citep{unet}.
 This architecture is commonly used for image segmentation: the input to a U-Net model is a 2D greyscale image, and the output is another 2D image with values representing the classification of each pixel.
@@
 
-Because the city boundaries aren't square and contain holes, missing data is introduced when placing the data into square images.
+Because the city boundaries aren't square and contain holes, missing data are introduced when placing the data into square images.
 The missing data needs to be filled because neural networks have no built-in way to handle non-real numbers.
@@
 \newpage
+\section{Detailed variable importance results }
+\label{ss:varimp}
+
+The results shown in Figure \ref{fig:importance_stack} show the stacked variable importance where the relative variable importance for each model is weighted by the model's R$^2$ value.
+In these two figures the raw relative variable influence (calculated using swing) is presented.
+This allows for the differences between the models to be compared.
+
+\begin{figure*}
+    \begin{center}
+    \includegraphics[width=\linewidth]{fig/report/importance_100.png}
+    \caption[Variable influence on LST at 100-meter resolution]{
+    Variable influence on LST at 100-meter resolution.
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
+    \label{fig:importance_100}
+    \end{center}
+\end{figure*}
+
+\begin{figure*}[h]
+    \begin{center}
+    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
+    \caption[Variable influence on LST at 500-meter resolution]{
+    Variable influence on LST at 500-meter resolution.
+    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
+    \label{fig:importance_500}
+    \end{center}
+\end{figure*}
+
+\clearpage
 \section{City specific results}
@@
 \begin{figure*}[h]
-    \begin{center}
-    \includegraphics[width=\linewidth]{fig/report/importance_500.png}
-    \caption[Variable influence on LST at 500-meter resolution]{
-    Variable influence on LST at 500-meter resolution.
-    The variable influence, measured by swing, shows the relative importance of each urban characteristic on land surface temperature.}
-    \label{fig:importance_500}
-    \end{center}
-\end{figure*}
-
-\begin{figure*}[h]
     \centering
@@
     \caption[Partial dependence contour plots for LST at 500-meter resolution during the night]{
-    Partial dependence contour plots for LST at 500-meter resolution during the night.
+    \textbf{Night.} Partial dependence contour plots for LST at 500-meter resolution during the night.
     These partial dependence contours show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value.
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
     }
@@
     \caption[Partial dependence contour plots for LST at 500-meter resolution during the day]{
-    Partial dependence contour plots for LST at 500-meter resolution during the day.
+    \textbf{Day.} Partial dependence contour plots for LST at 500-meter resolution during the day.
     These partial dependence contours show how the land surface temperature ($^oC$, y axis) changes with each variable as the other variables are held at their average value. 
-The left hand side shows the effect each variable has on LST during the night, while the right hand side shows the effect during the day.
-This shows that trees coverage in the cell has the greatest influence on the temperature, and the greenness (NDVI) of that coverage matters during the day.
     }
@@
     \caption{
-    Percentage tree canopy cover and impervious surface are 100\% correlated.
+    In the MRLC data provided by the USA's National Land Cover Database, the percentage tree canopy cover and impervious surface are 100\% correlated.
+Therefore we can only use one of these two variables in our statistical models.
+In the results we discuss the two factors together.
     }

### 2020-05-05 report/article.tex

@@
 We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}).
-Our code is available on our Github\footnote{URL redacted for review} repository, and follows the process described in \cite{Scott2016-lc}:
+Our code is available on our Github\footnote{\url{https://github.com/tommlogan/land_surface_temperature}} repository, and follows the process described in \cite{Scott2016-lc}:
 \begin{itemize}
@@
 \section*{Acknowledgements}
-% TML conducted the majority of this research as a graduate student at the University of Michigan with the support of a Rackham PreDoctoral Fellowship.
-% The work was also supported by the US National Science Foundation's Grant SEES-1631409. 
-% This support is gratefully acknowledged.
+The work was funded by the US National Science Foundation's Grant SEES-1631409 and a University of Michigan Rackham PreDoctoral Fellowship. 
+This support is gratefully acknowledged.
 
@@
 
-This is an extract of the code provided in the GitHub repository: []
+This is an extract of the code provided in the GitHub repository: \url{https://github.com/tommlogan/land_surface_temperature}
 
